---
title: "Parsing_QC"
author: "Ruben Schep"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  pdf_document:
    toc: yes
  editor_options:
    chunk_output_type: console
  html_document:
    theme: journal
    highlight: monochrome
    toc: yes
    toc_float: yes
    code_folding: show
---

knitr document van Steensel lab

# Parsing and starcode quality control
# Introduction
After demultiplexing all the samples from the sequencing data, the data has been processed by the CRISPR-TRIP snakemake script. This script maps the barcodes from the iPCR (iPCR from 5 feb 2018). It calls the true barcodes with the starcode script for both the mutation and mapping reads. It call the mutations on all the mutation samples and spits them in mapped, unmapped and non genuine.


Here we want to do a QC of the parsing of the mapping, the barcodes and mutations. We will load the statistics data of all the files. We will also look at barcode counts from the table files (those list the barcodes from the starcode stript). I obtained the counts though shell with 

Pasted the result in a text editor and changes all `^ ` and changed the spaces to `\t`, removed the last total row, removed `.table` at the end and saved as text file. 
I will export a bed file also containing the broad integrations for the generation of mean chip values over the integration sites. 


# Data loading and processing 

## Path, Libraries, Parameters and Useful Functions  

```{r message=FALSE}
StartTime <-Sys.time()

# 6-digit Date tag:
Date <- substr(gsub("-","",Sys.time()),1,8) 

# libraries:
library(plyr)
library(dplyr)
library(GenomicRanges)
library(RColorBrewer)
library(ggplot2)
library(magrittr)
library(platetools)
library(grid)
library(gridExtra)
```
 
## Custom functions  

Functions used in this script:   

```{r}
MakeGranges <- function(x) {
  x$end_pos <- as.numeric(as.character(x$start_pos+3))
  colnames(x) <- c("name",
                   "seqname", 
                   "strand", 
                   "start",
                   "total_mapped", 
                   "mapq_sum1", 
                   "reads1", 
                   "mapq_sum2", 
                   "reads2", 
                   "seq", 
                   "end")
  gr <- makeGRangesFromDataFrame(x, keep.extra.columns = T)
  return(gr)
}

BarcodeOverlap <- function(read1, read2) {
  counts <- c(nrow(read1), 
              nrow(read2), 
              sum(read2$barcode %in% read1$barcode), 
              sum(!read1$barcode %in% read2$barcode),
              sum(!read2$barcode %in% read1$barcode))
}

SetFileName <- function(filename, initials) {
  # Set filename with extension and initials to make filename with date integrated.
  filename <- substitute(filename)
  initials <- substitute(initials)
  filename <- paste0(initials, substr(gsub("-","",Sys.time()),1,8), filename)
  filename
}
```

## Data loading  

These are data from the crispr-trip.snake script, and a text file that has has been obtained as explained above.  

```{r}
# First I load the statistics data from the mapping.
setwd("/DATA/projects/DSBrepair/data/rs20210526_EpiScreen/parsed/")

file.list <- list.files(pattern='.*statistics.txt')
indel.statistics.list <- lapply(file.list, 
                                read.table, 
                                header = TRUE)
names(indel.statistics.list)<- gsub('indelPCR.(.*?).statistics.txt', 
                                    '\\1', 
                                    file.list)
# we need to remove the index column for the time being
indel.statistics.list <- lapply(indel.statistics.list, function(x) { x["index"] <- NULL; x })
indel.statistics.df <- do.call(rbind.data.frame, 
                               indel.statistics.list)
indel.statistics.df$sample.well <- rownames(indel.statistics.df)
indel.statistics.tib = indel.statistics.df %>%
  as_tibble() %>% 
  mutate(concentration = gsub("^.*_.._(1|10|100)_.*", "\\1", sample.well), 
        replicate = gsub("^(E1504|E177)_.*", "\\1", sample.well),
        drug_plate = as.integer(gsub("^E.*_(.*)_.*", "\\1", sample.well)),
        sample = gsub("^[E1504|E177]_.._(.*)$", "\\2", sample.well),
        plate_ID = gsub("(.*)_[A-Z].*", "\\1", sample),
        plate = gsub("^[E1504|E177]_(..).*", "\\1", sample),
        well = gsub(".*_._._(.*)$*", "\\1", sample.well),
        row = gsub("[0-999]", "\\1", well),
        column = gsub("[A-Z]", "\\1", well)) 

plates_numbers = read.table("/DATA/projects/DSBrepair/git/EpiScreen/plates_episcreen.txt",
                            sep = "\t", col.names = c("plate_ID", "plate"))


indel.statistics.tib$plate %<>% 
  as.numeric %>% sprintf("%02d", .)
indel.statistics.tib$column %<>% 
  as.numeric %>% sprintf("%02d", .)

indel.statistics.tib %<>% mutate(well = paste0(row, column))
  
```

### Indel parsing  


```{r}
low.reads.indel.stat.df <- indel.statistics.tib[order(indel.statistics.tib[, "reads"]), ][1:100, ]
mypalette5<-brewer.pal(5,"Greens")

barplot(as.matrix(t(low.reads.indel.stat.df[c(1:20),c(1,3,6:8)])), 
        col = mypalette5,
        beside = T, 
        main = "Read parsing statistics", 
        ylab = "reads", 
        legend = colnames(indel.statistics.tib[,c(1,3,6:8)]), 
        las = 2,
        cex.names = 0.4,
        cex.axis = 0.6,
        args.legend = list(x = "topright", cex = 0.6))

barplot(as.matrix(t(low.reads.indel.stat.df[c(1:50),c(1,3)])), 
        col = c("Red", "Orange"),
        beside = T, 
        main = "Read parsing statistics", 
        ylab = "reads", 
        legend = colnames(indel.statistics.tib[,c(1,3)]), 
        las = 2,
        cex.names = 0.4,
        cex.axis = 0.6,
        args.legend = list(x = "topright", cex = 0.6))

barplot(as.matrix(t(low.reads.indel.stat.df[c(51:100),c(1,3)])), 
        col = c("Red", "Orange"),
        beside = T, 
        main = "Read parsing statistics ", 
        ylab = "reads", 
        legend = colnames(indel.statistics.tib[,c(1,3)]), 
        las = 2,
        cex.names = 0.4,
        cex.axis = 0.6,
        args.legend = list(x = "topright", cex = 0.6))


b <- ggplot(indel.statistics.tib, aes(sample, log10(reads))) + 
  geom_bar(stat = "identity", aes(fill = concentration))
b
b <- ggplot(indel.statistics.tib, aes(sample, reads)) + 
  geom_bar(stat = "identity", aes(fill = plate))
b
b <- ggplot(indel.statistics.tib[indel.statistics.tib$replicate=="rep2" & indel.statistics.tib$concentration=="1um", ], aes(sample, reads)) + 
  geom_bar(stat = "identity", aes(fill = replicate))
b
# How many reads per sample?
indel.statistics.tib %>% group_by(sample) %>% summarise(reads = sum(reads))
# How many reads per row?
indel.statistics.tib %>% group_by(row) %>% summarise(reads = sum(reads))
# How many reads per column?
indel.statistics.tib %>% group_by(column) %>% summarise(reads = sum(reads))
# How many reads per well (for the 96 index)?
well.statistics <- indel.statistics.tib %>% group_by(replicate, well) %>% dplyr::summarise(reads = sum(reads))
well.statistics

platesE177 = well.statistics %>% 
  filter(replicate == "E177")

raw_map(platesE177$reads, platesE177$well, plate = 96) + 
  scale_fill_distiller(type = "div", name = "reads")

platesE1504 = well.statistics %>% 
  filter(replicate == "E1504")

raw_map(platesE1504$reads, platesE1504$well, plate = 96) + 
  scale_fill_distiller(type = "div", name = "reads")

# % of reads written from the total reads :
indelreadswritten.vec <- indel.statistics.tib$reads_written/indel.statistics.tib$reads*100
names(indelreadswritten.vec) <- rownames(indel.statistics.tib)
barplot(indelreadswritten.vec, 
        las=2, 
        cex.names = 0.4)

```

```{r, fig.width = 12, fig.height = 4}
for (i in 1:18) {
  
  # Select the flowjo data set
  data <- indel.statistics.tib[indel.statistics.tib$plate == i, ]
  
  # Find reasonable limits 
  limits <- c(2000, 25000)

  # Plot this, first the PI staining with a custom color-coded density
  y <- density(data$reads)
  plt1 <- ggplot(data.frame(x = y$x, y = y$y),
                 aes(x = x, y = y)) +
    geom_line() +
    geom_segment(aes(xend = x, yend = 0, colour = x)) +
    xlim(limits) +
    ggtitle(paste("Cell cycle -",
                unique(data$sample))) +
    scale_color_distiller(type = "div", limits = limits, name = "PI_area") +
    theme_bw() +
    theme(legend.position="none")
  
  # Then the plate with color-coded wells
  plt2 <- raw_map(data$reads, 
                 data$well) +
    ggtitle(unique(data$sample)) +
    scale_fill_distiller(type = "div", limits = limits, name = "reads")
  plt2
  grid.arrange(plt1, plt2, ncol = 2, widths = c(1, 1.3))
}
```

# Conclusion
- In general, the data looks very good
- Some plates have wells with 0 reads - other wells take over the reads in this plate - we have to investigate further whether this can be a problem.

# Session Info  
```{r}
sessionInfo()
getwd()
date()
paste("Run time: ",format(Sys.time()-StartTime))

```