---
title: "FACS analysis"
author: "Max Trauernicht"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output: 
  html_document:
    theme: journal #cerulean
    highlight: monochrome
    toc: true
    toc_float: true
    code_folding: show
  editor_options:
    chunk_output_type: console
---

# knitr document van Steensel lab

# TRIP Clone identification

# Introduction
This script generates plots of FACS data. This way the effect of drugs on cell cycle will be visualized.
In this particular experiment, all effective drugs that did lead to a change of +1/-7 ratio were subjected to FACS. 


## Description of Data
How to make a good rendering table: 
```{r table1, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
tabl <- "
| column1	|	column2	|	column3	|	
|----|----|----|
|1	|	2	|	3	|	
|a	|	b	|	c	|	
"
cat(tabl) # output the table in a format good for HTML/PDF/docx conversion
```

# Data processing
## Path, Libraries, Parameters and Useful Functions
```{r setup}
knitr::opts_chunk$set(echo = TRUE)
StartTime <-Sys.time()

# 8-digit Date tag:
Date <- substr(gsub("-","",Sys.time()),1,8) 
# libraries:
library(tidyverse)
library(maditr)
library(ggplot2)
```

## Data import: data from 2 experiments is sampled in this analysis
```{r}
setwd("/DATA/usr/m.trauernicht/projects/EpiScreen/FACS/Data/")
gfp.samples <- read.csv2("mt20190419_FlowJo_GFP.csv")
lbr.samples <- read.csv2("mt20190419_LBR_FlowJo.csv")

# Import data from experiment where I investigated cell cycle arrest in 5 candidate drugs
five.drugs <- get(load("mt20190605_five.drugs"))
```

# Analysis

## Bring dataframe in right format
```{r}
# Remove not needed rows
gfp.samples <- gfp.samples[gfp.samples$Statistic != "",]
lbr.samples <- lbr.samples[lbr.samples$Statistic != "",]

# Remove not needed column
gfp.samples <- gfp.samples[,-1]
lbr.samples <- lbr.samples[,-1]

# Rename column
gfp.samples$Name <- gsub("[A-Z][1-9].fcs/Cells(.*?)", "\\1", gfp.samples$Name)
lbr.samples$Name <- gsub("[A-Z][1-9].fcs/Cells(.*?)", "\\1", lbr.samples$Name)
gfp.samples$Name <- gsub("[A-Z][1-9][0-9].fcs/Cells(.*?)", "\\1", gfp.samples$Name)
lbr.samples$Name <- gsub("[A-Z][1-9][0-9].fcs/Cells(.*?)", "\\1", lbr.samples$Name)
gfp.samples$Name <- gsub("/Single Cells", "\\1", gfp.samples$Name)
lbr.samples$Name <- gsub("/Single Cells", "\\1", lbr.samples$Name)
gfp.samples$Name <- gsub("/", "\\1", gfp.samples$Name)
lbr.samples$Name <- gsub("/", "\\1", lbr.samples$Name)


# Remove not needed rows
gfp.samples <- gfp.samples[gfp.samples$Name != "",]
lbr.samples <- lbr.samples[lbr.samples$Name != "",]

# Make apopototic cells df
gfp.samples$Statistic <- as.numeric(as.character(gfp.samples$Statistic))
apoptotic.samples.gfp <- gfp.samples[gfp.samples$Name == "Apoptotic cells",]
apoptotic.samples.gfp$drug <- paste0(apoptotic.samples.gfp$drug, "_", apoptotic.samples.gfp$conc)
lbr.samples$Statistic <- as.numeric(as.character(lbr.samples$Statistic))
apoptotic.samples.lbr <- lbr.samples[lbr.samples$Name == "Apoptotic cells",]
apoptotic.samples.lbr$drug <- paste0(apoptotic.samples.lbr$drug, "_", apoptotic.samples.lbr$conc)

# Look at which samples have highest apoptotic cell fractions
ggplot(data = apoptotic.samples.gfp[apoptotic.samples.gfp$Statistic >= 10,], aes(x = reorder(drug, -Statistic), y = Statistic)) +
  geom_bar(stat = "identity")+
   theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 10))+ 
  ylab("Apoptotic cell population (%)") + xlab("Drug and concentration")+ labs(title = "Apoptotic cell population after drug treatment - GFP samples")

ggplot(data = apoptotic.samples.lbr[apoptotic.samples.lbr$Statistic >= 10,], aes(x = reorder(drug, -Statistic), y = Statistic)) +
  geom_bar(stat = "identity")+
   theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 10))+ 
  ylab("Apoptotic cell population (%)") + xlab("Drug and concentration")+ labs(title = "Apoptotic cell population after drug treatment - LBR2 samples")


# Remove data with low cell numbers 
gfp.samples$cell.count <- ave(gfp.samples$X.Cells, gfp.samples$drug, gfp.samples$conc, gfp.samples$rep, FUN = function(x) sum(x))
gfp.samples <- gfp.samples[gfp.samples$cell.count > 500,]
gfp.samples <- gfp.samples [,-7]

## Rearrange to wide format
gfp.samples <- dcast(gfp.samples, drug + conc + rep ~ Name, value.var = "Statistic", fun.aggregate = mean)
lbr.samples <- dcast(lbr.samples, drug + conc + rep ~ Name, value.var = "Statistic", fun.aggregate = mean)

# Compute relative G2, G1 and S fraction
for (i in gfp.samples$rep) {
gfp.samples$sum.apo[gfp.samples$rep == i] <- gfp.samples$G1[gfp.samples$rep == i] + 
  gfp.samples$S[gfp.samples$rep == i] + gfp.samples$G2[gfp.samples$rep == i] + 
  gfp.samples$`Apoptotic cells`[gfp.samples$rep == i]
}
for (i in gfp.samples$rep) {
gfp.samples$sum[gfp.samples$rep == i] <- gfp.samples$G1[gfp.samples$rep == i] + 
  gfp.samples$S[gfp.samples$rep == i] + gfp.samples$G2[gfp.samples$rep == i] 
}

gfp.samples$G2 <- gfp.samples$G2 / gfp.samples$sum
gfp.samples$G1 <- gfp.samples$G1 / gfp.samples$sum
gfp.samples$S <- gfp.samples$S / gfp.samples$sum
gfp.samples$`Apoptotic cells` <- gfp.samples$`Apoptotic cells` / gfp.samples$sum.apo
gfp.samples <- gfp.samples[-8]
gfp.samples$ID <- "GFP"

for (i in lbr.samples$rep) {
lbr.samples$sum.apo[lbr.samples$rep == i] <- lbr.samples$G1[lbr.samples$rep == i] + 
  lbr.samples$S[lbr.samples$rep == i] + lbr.samples$G2[lbr.samples$rep == i] + 
  lbr.samples$`Apoptotic cells`[lbr.samples$rep == i]
}
for (i in lbr.samples$rep) {
lbr.samples$sum[lbr.samples$rep == i] <- lbr.samples$G1[lbr.samples$rep == i] + 
  lbr.samples$S[lbr.samples$rep == i] + lbr.samples$G2[lbr.samples$rep == i] 
}

lbr.samples$G2 <- lbr.samples$G2 / lbr.samples$sum
lbr.samples$G1 <- lbr.samples$G1 / lbr.samples$sum
lbr.samples$S <- lbr.samples$S / lbr.samples$sum
lbr.samples$`Apoptotic cells` <- lbr.samples$`Apoptotic cells` / lbr.samples$sum.apo
lbr.samples <- lbr.samples[-8]
lbr.samples$ID <- "LBR2"

# Remove conditions with high apoptotic cell counts
# gfp.samples <- gfp.samples[gfp.samples$`Apoptotic cells` < 0.10,]
# lbr.samples <- lbr.samples[lbr.samples$`Apoptotic cells` < 0.10,]


# Merge the two dfs
all.samples.unfiltered <- rbind(lbr.samples, gfp.samples)
all.samples.unfiltered <- melt(all.samples.unfiltered, id.vars = c("drug", "conc", "rep", "ID"))

repl.filter = all.samples.unfiltered %>% group_by(drug, conc) %>% distinct(drug, conc, ID, rep) %>% count() %>% filter(n == 4) %>% distinct(drug, conc)
all.samples = repl.filter %>% left_join(all.samples.unfiltered) %>% ungroup()  %>% 
  filter(variable %in% c("G2", "S", "G1", "Apoptotic cells") & !is.na(value & drug != "PAO")) %>%
  mutate(conc_char = ifelse(conc == "100nm", "100 nM", 
                            ifelse(conc == "1um", "1 µM", "10 µM")),
    drug_conc = paste0(drug, " ", conc_char)) %>%
  group_by(drug_conc, drug, conc_char, variable, ID) %>%
  summarise(SD = sd(value),
    mean_value =  mean(value)) %>%
  ungroup()
```


# Conclusion: Aurora kinases induce G2 arrest

## Exporting potential data. 
```{r}
all.samples.export <- all.samples.unfiltered %>%
  mutate(exp = "effective_drugs")

write_csv(all.samples.export, "/DATA/usr/m.trauernicht/projects/EpiScreen/files_scripts/mt20220802_facs_all_drugs.csv")
```

# Session Info
```{r}
paste("Run time: ",format(Sys.time()-StartTime))
getwd()
date()
sessionInfo()
```

