---
title: "DSB epigenetic drug screen - Figure 3"
author:
  - name: "Max Trauernicht & Ruben Schep"
    email: "m.trauernicht@nki.nl"
    affiliation: "Netherlands Cancer Institute"
date: '`r format(Sys.time(), "%d/%m/%Y")`'
output:
  html_document:
    theme: united
    highlight: pygments
    fig_caption: yes
    code_folding: hide
    df_print: kable
    toc: yes
    toc_depth: 4
    toc_float:
      collapsed: no
  pdf_document:
    toc: yes
    toc_depth: '4'
---

*knitr document van Steensel lab*


# Introduction
Clone 5 with ~18 sgRNA target site IPRs was cut and treated with 160 epigenetic drugs. The repair outcomes of that experiment will be visualized here. I previously processed the raw sequencing data, and calculated the repair outcomes. In this script, the main results plots will be generated.


## Setup {.tabset}

<!-- little HTML script to do indentation of the table of contents -->
<script>
$(document).ready(function() {
$items = $('div#TOC li');
$items.each(function(idx) {
num_ul = $(this).parentsUntil('#TOC').length;
$(this).css({'text-indent': num_ul * 10, 'padding-left': 0});
});
});
</script>

```{css, echo = FALSE}
div.sourceCode {
overflow-x: hidden;
}
```
### Libraries

```{r load_packages, results = 'hide', warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
StartTime <-Sys.time()

# 8-digit Date tag:
Date <- substr(gsub("-","",Sys.time()),1,8) 
# libraries:
library(knitr)
library(tidyverse)
library(outliers)
library(ggplot2)
library(ggbeeswarm)
library(data.table)
library(ggrepel)
library(GGally)
# library(DESeq2)
library(gghighlight)
# library(platetools)
library(ggpubr)
# library(Laurae)
library(pheatmap)
library(RColorBrewer)
# # library(ggrastr)
library(broom)
# library(ggsignif)
library(cowplot)
library(rstatix)
library(colorspace)
library(Hmisc)
library(pls)
library(MASS)
library(corrr)
library(ggstatsplot)


## Select outdir
out.dir = paste0("./figures/", Date, "/")
dir.create(out.dir)
in.dir = list.dirs(path = "/DATA/projects/DSBrepair/data/R") %>% .[grepl("episcreen", .)] %>% tail(n = 1)
opts_chunk$set(dev=c('png', 'pdf'), fig.path = file.path(out.dir))
pdf.options(useDingbats = FALSE)
```

### Functions

```{r functions, message=FALSE, warnings=FALSE}
CorrelationDFMean <- function(data, feature_name, targets, p.value=0.001){
  hist.corr.list <- data %>% dplyr::select(all_of(feature_name), targets) %>% as.matrix() %>% rcorr(type = "spearman")
  
  histone.corr.df <- data.frame(rowname = names(hist.corr.list$r[1, -1]),
                                corr = hist.corr.list$r[1, -1],
                                pvalue = hist.corr.list$P[1, -1])
  
  return(histone.corr.df=histone.corr.df)
}


CorrelationPlotMean <- function(data, feature_name, targets, title = "with treatment", p.value=0.01, alfa = c(0.5, 1)){
  cor.df = data.frame()
  for (i in feature_name){
    corr.list <- data %>% 
      dplyr::select(all_of(i), as.character(targets)) %>% 
      as.matrix() %>% 
      rcorr(type = "spearman")
    
    df <- data.frame(rowname = names(corr.list$r[1, -1]),
                     corr = corr.list$r[1, -1],
                     pvalue = corr.list$P[1, -1],
                     feature = i)
    cor.df <- rbind(cor.df, df) %>%
      transform(rowname = factor(rowname, levels = levels(targets)),
                corr = corr,
                pvalue = pvalue) %>% 
      mutate(p.adj = p.adjust(pvalue)) %>%
      add_significance("p.adj")
  }
  p <-  ggplot(cor.df, aes(x = rowname, y = corr, fill = p.adj.signif)) +
    geom_bar(stat = "identity", position = "dodge") +
    ylab(paste("Spearman correlation ",title)) +
    xlab("Chromatin Feature") +
    # ggtitle(paste("Repair pathways correlation with chromatin features in", condition_name, "exps"))  +
    theme_classic2(base_size = 16) +
    scale_y_continuous(breaks=seq(-1,1, .1)) +
    scale_fill_manual(values = c("****" = "black", "***" = "gray20", "**" = "gray40", 
                                 "*" = "gray50", "ns" = "gray80")) +
    coord_flip()
  return(list(dataframe=cor.df, plot = p))
}

clustering = data.frame(row.names=c('LMNB1', 'late_replicating', 'H3K9me2',
                                    'H3K27me3', 'EZH2', 'CTCF', 'SMC3',
                                    'HDAC3', 'HDAC2', 'HDAC1', 'H3K4me1',
                                    'H3K4me2', 'H3K4me3', 'H3K27ac',
                                    'H4K5acK8ac', 'H2AFZ', 'DNAse', 'Dam', 
                                    'm5C', 'H3K79me2', 'TTseq', 'H3K36me3', 
                                    'POL2AS2', 'POL2'),
                        group=factor(c(rep('repressive', 5), rep('insulator', 2),
                                       rep('HDAC', 3), rep('euchromatin', 6),
                                       rep('accessibility', 2), 'methylation',
                                       rep('transcribing', 5)),
                                     levels=c('transcribing', 'accessibility',
                                              'methylation', 'euchromatin',
                                              'HDAC', 'insulator', 'repressive')))

cor_test <- function(y, X, left, right){
  cor_list = lapply(X, function(x){
    cor.test(x,y)
  })
  p_vec = unlist(lapply(cor_list, function(x){x$p.value}))
  cor_vec = unlist(lapply(cor_list, function(x){x$estimate}))
  p_adj = p.adjust(p_vec, method = "holm")
  color = ifelse(p_adj < 0.05, ifelse(cor_vec > 0, left, right),
                 'None')
  importance = abs(cor_vec)
  target=factor(colnames(X), levels=rownames(clustering))
  feature_group = clustering[colnames(X), ]
  return(list(correlation=cor_vec,importance = importance, p=p_vec,
              p_adj=p_adj, target=target, color=color,
              feature_group=feature_group))
}

get_combinations <- function(domain_vec){
  n_domain = length(domain_vec)
  get_ivec <-function (last_x, n_domain){
    if (last_x < n_domain){
      i_vec = c(0, c((last_x+1):n_domain))
    } else{
      i_vec = c(0)
    }
  }
  
  combine <- function(x_new, x_vec, n_domain){
    new_vec = c(x_vec, x_new)
    if (length(new_vec)==n_domain){
      return(new_vec)
    } else {
      i_vec = get_ivec(max(new_vec), n_domain)
      x_list = lapply(i_vec, combine, x_vec=new_vec,
                      n_domain=n_domain)
      return(do.call(rbind, x_list))
    }
  }
  
  comb_list = lapply(c(0:n_domain), combine, x_vec=c(), n_domain=n_domain)
  
  comb = do.call(rbind, comb_list)
  comb_df = unique(t(apply(comb, 1, function(x){x[order(x)]})))
  
  name_list = apply(comb_df, 1, function(x){
    name = paste(domain_vec[x], collapse='-')
    if (name == ''){
      name = 'iDomain'
    }
    return(name)
  })
  
  opt_list = lapply(domain_vec, function(d){
    apply(comb_df, 1, function(x){
      opt = domain_vec[x]
      return(d%in%opt)
    })
  })
  
  opt_df = do.call(cbind, opt_list)
  colnames(opt_df) = domain_vec
  rownames(opt_df) = name_list
  
  return(opt_df)
}

quantiles_95 <- function(x) {
  r <- quantile(x, probs=c(0.05, 0.25, 0.5, 0.75, 0.95))
  names(r) <- c("ymin", "lower", "middle", "upper", "ymax")
  r
}

plot_comb_grid_beeswarm <- function(dt, opt_df, domains, y_name, min_count=0,
                                    max_count=Inf, group_by=NULL, color=NULL, y_plabel=c(-1.5, 1.5)){
  
  opt_melt = reshape2::melt(opt_df)
  colnames(opt_melt) = c('group', 'domain', 'is_in')
  
  opt_melt$is_in = ifelse(opt_melt$is_in, 'in', 'out')
  opt_melt$domain = factor(opt_melt$domain, levels=domains)
  
  
  group_list = lapply(rownames(opt_df), function(opt_name){
    opt = opt_df[opt_name, ]
    in_vec = domains[opt]
    out_vec = domains[!opt]
    
    in_sum = rowSums(dt[,..in_vec])
    out_sum = rowSums(dt[,..out_vec])
    if (length(in_vec)==0){
      in_group = which(out_sum == 0)
    } else if (length(out_vec) == 0){
      in_group = which(in_sum==length(in_vec))
    } else  {
      in_group = which(in_sum == length(in_vec) & out_sum==0)
    }
    return(in_group)
  })
  
  names(group_list) = rownames(opt_df)
  for (opt_name in names(group_list)){
    i_vec = group_list[[opt_name]]
    dt[i_vec, group := opt_name]
  }
  if (is.null(group_by)){
    group_count_df = data.frame(dt[,table(group)])
  } else {
    group_count_df = data.frame(dt[,table(group)/length(unique(get(group_by)))])
  }
  group_count_df = group_count_df[order(match(group_count_df$group, rownames(opt_df))),]
  # print(group_count_df)
  group_count_sel = group_count_df[group_count_df$Freq>=min_count,]
  group_count_sel = group_count_sel[group_count_sel$Freq<=max_count,]
  group_levels =  group_count_sel$group
  opt_sel = opt_melt[opt_melt$group %in% group_levels, ]
  
  opt_sel$group = factor(opt_sel$group, levels=group_levels)
  # print(opt_sel)
  
  indel_selection = dt[group%in%group_levels, ]
  indel_selection[,group:=factor(group, levels=group_levels)]
  
  # wilcox_list = lapply(group_levels[group_levels!='iDomain'],
  #                      function(g){
  #                          x = indel_selection[group==g, y_name, with=F]
  #                          y = indel_selection[group=='iDomain', y_name,
  #                                              with=F]
  #                          wilcox.test(unlist(x),unlist(y))
  #                      })
  
  group_count_sel$group = factor(group_count_sel$group, levels=group_levels)
  
  count_plot = ggplot(group_count_sel, aes(x=group,y=1, label=Freq)) +
    geom_label() +
    theme_void()
  
  
  
  dots = ggplot(opt_sel, aes(x=group, y=domain, color=is_in)) +
    geom_point(size=2) +
    geom_line(data=opt_sel[opt_sel$is_in=='in',], aes(group=group)) +
    theme_bw() +
    scale_color_manual(values=c('in'='black', 'out'='grey70')) +
    guides(color="none")  +
    geom_vline(xintercept=seq(1.5, length(unique(dt$group))-0.5, 1),
               lwd=0.5, colour="grey90") +
    theme(axis.title=element_blank(),
          axis.text=element_blank(),
          axis.ticks=element_blank(),
          panel.grid.major = element_blank())
  
  symnum = list(cutpoints = c(0, 0.0001, 0.001, 0.01, 0.05, 1),
                symbols = c("****", "***", "**", "*", " "))
  
  
  if (is.null(group_by)){
    beeswarm = ggplot(indel_selection, aes_string(x='group', y=y_name)) +
      geom_quasirandom(width=0.3) +
      stat_summary(fun=median, fun.min = median,
                   fun.max = median,
                   geom="crossbar", width = 0.5,
                   color='red') +
      stat_compare_means(aes(label = ..p.signif..),
                         symnum.args = symnum,
                         method = 'wilcox.test',
                         ref.group='iDomain')
    
  } else {
    wilcox_dt = indel_selection[,test_wilcox(.SD, group_by, y_name),
                                by='group']
    wilcox_dt[,p_adj:=p.adjust(p_value, method = "holm")]
    wilcox_dt[,p_signif:=symnum(p_adj, cutpoints=symnum$cutpoints,
                                symbols=symnum$symbols)]
    
    
    beeswarm = ggplot(indel_selection,
                      aes_string(x='group', y=y_name, color=group_by)) +
      geom_quasirandom(dodge.width=1) +
      scale_color_manual(values=color) +
      stat_summary(aes_string(group=group_by), fun=median,
                   fun.min = median, fun.max = median,
                   geom="crossbar", width = 0.5,
                   color='red',
                   position= position_dodge(width =1)) +
      geom_text(data=wilcox_dt, aes(x=group, y=0.8, label=p_signif),
                inherit.aes=F)
    print(wilcox_dt)
  }
  beeswarm = beeswarm +
    theme_bw() +
    geom_vline(xintercept=seq(1.5, length(unique(dt$group))-0.5, 1),
               lwd=0.5, colour="grey90") +
    theme(axis.title=element_blank(),
          axis.text.x=element_blank(),
          panel.grid.major.x = element_blank(),
          axis.ticks.x=element_blank()) +
    ylim(y_plabel)
  
  count = data.frame(domain=gsub('_domain', '', domains),
                     count=colSums(dt[,..domains], na.rm=T))
  
  count$domain = factor(count$domain, levels=domain_levels)
  
  histogram = ggplot(count, aes(x=1, y=domain, label = count)) +
    geom_label() +
    theme_bw() +
    theme(plot.background = element_blank(),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.border = element_blank(),
          axis.title=element_blank(),
          axis.text.x=element_blank(),
          axis.ticks=element_blank())
  
  empty = ggplot(data=data.frame(1,1),aes(x=1,y=1)) +
    geom_point() +
    theme_void()
  
  count_beeswarm = plot_grid(beeswarm, count_plot, ncol=1, rel_heights=c(20,1),
                             align='v', axis='lr')
  return(plot_grid(empty, beeswarm, empty, count_plot,
                   histogram, dots, nrow=3,
                   rel_widths=c(1,10), rel_heights=c(15,1,5), align='vh',
                   axis='tlbr'))
}

p.adjust <- function(p, method = p.adjust.methods, n = length(p))
{
  ## Methods 'Hommel', 'holm', 'BY' and speed improvements
  ## contributed by Gordon Smyth
  method <- match.arg(method)
  if(method == "fdr") method <- "holm"	# back compatibility
  nm <- names(p)
  p <- as.numeric(p)
  p0 <- setNames(p, nm)
  if(all(nna <- !is.na(p))) nna <- TRUE
  p <- p[nna]
  lp <- length(p)
  # stopifnot(n >= lp)
  if (n <= 1) return(p0)
  if (n == 2 && method == "hommel") method <- "hochberg"
  
  p0[nna] <-
    switch(method,
           bonferroni = pmin(1, n * p),
           holm = {
             i <- seq_len(lp)
             o <- order(p)
             ro <- order(o)
             pmin(1, cummax( (n - i + 1L) * p[o] ))[ro]
           },
           hommel = { ## needs n-1 >= 2 in for() below
             if(n > lp) p <- c(p, rep.int(1, n-lp))
             i <- seq_len(n)
             o <- order(p)
             p <- p[o]
             ro <- order(o)
             q <- pa <- rep.int( min(n*p/i), n)
             for (j in (n-1):2) {
               ij <- seq_len(n-j+1)
               i2 <- (n-j+2):n
               q1 <- min(j*p[i2]/(2:j))
               q[ij] <- pmin(j*p[ij], q1)
               q[i2] <- q[n-j+1]
               pa <- pmax(pa,q)
             }
             pmax(pa,p)[if(lp < n) ro[1:lp] else ro]
           },
           hochberg = {
             i <- lp:1L
             o <- order(p, decreasing = TRUE)
             ro <- order(o)
             pmin(1, cummin( (n - i + 1L) * p[o] ))[ro]
           },
           BH = {
             i <- lp:1L
             o <- order(p, decreasing = TRUE)
             ro <- order(o)
             pmin(1, cummin( n / i * p[o] ))[ro]
           },
           BY = {
             i <- lp:1L
             o <- order(p, decreasing = TRUE)
             ro <- order(o)
             q <- sum(1L/(1L:n))
             pmin(1, cummin(q * n / i * p[o]))[ro]
           },
           none = p)
  p0
}

SetFileName <- function(filename, initials) {
  # Set filename with extension and initials to make filename with date integrated.
  filename <- substitute(filename)
  initials <- substitute(initials)
  filename <- paste0(initials, Date, filename)
  filename
}

# From Fede:
# ggpairs custom functions
corColor <- function(data, mapping, color = I("black"), sizeRange = c(1, 3), ...) {
  
  x   <- eval_data_col(data, mapping$x)
  y   <- eval_data_col(data, mapping$y)
  r   <- cor(x, y)
  rt  <- format(r, digits = 3)
  tt  <- as.character(rt)
  cex <- max(sizeRange)
  
  # helper function to calculate a useable size
  percent_of_range <- function(percent, range) {
    percent * diff(range) + min(range, na.rm = TRUE)
  }
  
  # plot correlation coefficient
  p <- ggally_text(label = tt, mapping = aes(), xP = 0.5, yP = 0.5,
                   size = I(percent_of_range(cex * abs(r), sizeRange)), color = color, ...) +
    theme(panel.grid.minor=element_blank(),
          panel.grid.major=element_blank())
  
  corColors <- RColorBrewer::brewer.pal(n = 7, name = "RdYlBu")[2:6]
  
  if (r <= boundaries[1]) {
    corCol <- corColors[1]
  } else if (r <= boundaries[2]) {
    corCol <- corColors[2]
  } else if (r < boundaries[3]) {
    corCol <- corColors[3]
  } else if (r < boundaries[4]) {
    corCol <- corColors[4]
  } else {
    corCol <- corColors[5]
  }
  
  p <- p +
    theme(panel.background = element_rect(fill = corCol))
  
  return(p)
}

theme_classic_lines <- function() {
  theme_pubr(border = T, legend = "top") +
    theme(panel.grid.major = element_line(colour = "#adb5bd", size = 0.25),
          strip.background = element_rect(fill = "white", color = "white")
    )
}
theme_classic_lines_45 <- function() {
  theme_pubr(border = T, legend = "top", x.text.angle = 45) +
    theme(panel.grid.major = element_line(colour = "#adb5bd", size = 0.25),
          strip.background = element_rect(fill = "white", color = "white")
    )
}

theme_classic_lines_90 <- function() {
  theme_pubr(border = T, legend = "top", x.text.angle = 90) +
    theme(panel.grid.major = element_line(colour = "#adb5bd", size = 0.25),
          strip.background = element_rect(fill = "white", color = "white")
    )
}

theme_classic_lines_90_noborder <- function() {
  theme_pubr(border = F, legend = "top", x.text.angle = 90) +
    theme(strip.background = element_rect(fill = "white", color = "white")
    )
}

theme_set(theme_classic_lines())

colors_diverse <- c("#264653", "#9AC1AE", "#5D987B", "#f2cc8f", "#e76f51")
# colors_drugs <- c("DMSO (neg control)" = "#65BFA4", "DNA-PKi (NHEJ control)" = "#8CA0C4", "drugs" = "#808184", "Mirin (MMEJ control)" = "#5D79AC")
colors_drugs <- c("#FA8D62", "#65BFA4", "#8CA0C4", "#808184", "#F6D289", "#E5E5E5", "#ACACAC", "#737373")
#colors_diverse <- c("#264653", "#2a9d8f", "#e9c46a", "#f4a261", "#e76f51") 
## Chromatin groups
group_colors = c("Euchromatin" = "#f7941d",
                 "other-heterochromatin" = "#838687",
                 "Triple Heterochromatin" = "#662d91",
                 "H3K27me3" = "#d21f8a")

colors_targets <-  c(HDAC="#60988D", HAT="#C6D0A8", Sirtuin="#8FCAAC",
                     HMT="#FDCDAC", DNMT="#D69C81", `Histone Demethylase`="#FFF2AE",
                     HIF="#E3DCD5", JAK="#C8D7E2", PIM="#9AB8C4", `Aurora Kinase`="#F4CAE4",
                     PARP="#CB546F", `Epigenetic Reader Domain`= "#476D61", `DNA-PK`="#E07A43")

conc_control_colors = c("CTRL" = colors_drugs[2], "100 nM" = colors_drugs[7], "1 µM" = colors_drugs[4], "10 µM" = "#000000")
conc_colors = c("100 nM" = colors_drugs[7], "1 µM" = colors_drugs[4], "10 µM" = "#000000")

ggplot_custom <- function(...) ggplot2::ggplot(...) + 
  scale_color_manual(values = colors_diverse) + 
  scale_fill_manual(values = colors_diverse)

```

---

# Data import

## Screen and validation data

#### Import
```{r data import, echo = FALSE, warning = FALSE}
# Import data from preprocessing script
in.dir = list.dirs(path = "/DATA/projects/DSBrepair/data/R") %>% .[grepl("episcreen", .)] %>% tail(n = 1)
file = list.files(in.dir, pattern = "episcreen_ratios", full.names = T) %>% tail(n = 1)
indel.data <- readRDS(file) # %>% filter(pathway_bal_filter == TRUE & viab_reproducibility == TRUE)
```

---

#### Heatmap annotation
```{r heatmap annotation}
# Order IPRs by their starting freqCut value
dmso_freq <- indel.data %>%
  filter(drug == "DMSO") %>%
  distinct(IPR, freqCut) %>%
  mutate(freqCut = ave(freqCut, IPR, FUN = function(x) mean(x, na.rm = TRUE))*100) %>%
  dplyr::select('ctrl_TIF' = freqCut, IPR) %>%
  distinct() %>%
  arrange(desc(ctrl_TIF))

iprs <- dmso_freq$IPR
iprs <- as.character(iprs)

# Set annotation for heatmap
# viability
viab_drug = indel.data %>% filter(drug != "PAO") %>% 
  mutate(drug = gsub(" [(].*[)]", "", drug),
         drug_conc = paste(drug, conc_char, sep = " ")) %>% 
  distinct(viab_mean, drug_conc, replicate) %>% 
  group_by(drug_conc) %>%
  summarise(viab = mean(viab_mean))
# Row annotation: annotate drugs with the target group
target_set_drugs <- indel.data %>% 
  filter(!target %in% c("Negative Control", "MRN", "DNA-PK")) %>%
  mutate(drug = gsub(" [(].*[)]", "", drug),
         drug_conc = paste(drug, conc_char, sep = " ")) %>%
  left_join(viab_drug) %>%
  distinct(drug, drug_conc, conc_char, target, viab) %>% 
  column_to_rownames(var="drug_conc")

# Set annotation for heatmap
# viability
viab_drug = indel.data %>% filter(drug != "PAO") %>% 
  mutate(drug = gsub(" [(].*[)]", "", drug),
         drug_conc = paste(drug, conc_char, sep = " ")) %>% 
  distinct(viab_mean, drug_conc, replicate) %>% 
  group_by(drug_conc) %>%
  summarise(viab = mean(viab_mean))
# Row annotation: annotate drugs with the target group
target_set_drugs <- indel.data %>% 
  filter(!target %in% c("Negative Control", "MRN", "DNA-PK")) %>%
  mutate(drug = gsub(" [(].*[)]", "", drug),
         drug_conc = paste(drug, conc_char, sep = " ")) %>%
  left_join(viab_drug) %>%
  distinct(drug, drug_conc, conc_char, target, viab) %>% 
  column_to_rownames(var="drug_conc")

ctrl.vals <- indel.data %>%
  filter(drug == "DMSO") %>%
  distinct(IPR, freqCut, MMEJratio) %>%
  mutate(ctrl_TIF = ave(freqCut, IPR, FUN = function(x) mean(x, na.rm = TRUE))*100,
         ctrl_balance = ave(MMEJratio, IPR, FUN = function(x) mean(x, na.rm = TRUE))*100) %>%
  distinct(ctrl_TIF, 
           ctrl_balance,
           IPR)

HDAC_tib = read.table("rs20220708_HDAC_targets.txt", sep = "\t", fill = NA, header = T) %>%
  dplyr::select(-target_main, -target_second) %>%
  as_tibble()

clone5.matrix = indel.data %>% 
  dplyr::select(IPR, ends_with(".zscore")) %>% 
  distinct() %>% 
  rename_with(., ~gsub(".zscore", "", .x)) %>%
  # left_join(ctrl.vals) %>%
  column_to_rownames(var="IPR")

clone5_map_ipr = hclust(dist(clone5.matrix), method = "ward.D") %>% 
  # Rotate the middle node to math better the TIF in CTRL conditions
  dendextend::rotate(c(1:6, 10:7, 11:18))

# Chromatin annotation
chromatin <- readRDS(
  "/DATA/usr/x.vergara/XV_ChIPIndels/XV20200902_DDR_RS_revision/XV20200902_DDR_RS_revision/data/xv20200915_DDR_data_CHIP.rds") %>% 
  distinct(barcode, chromatin) %>%
  left_join(distinct(indel.data, barcode, IPR)) %>%
  left_join(dmso_freq)

# chromatin$IPR[is.na(chromatin$IPR)] <- "IPR1"
chromatin <- chromatin %>%
  filter(IPR != "<NA>") %>% 
  dplyr::select(-barcode) %>%
  column_to_rownames(var="IPR")

## Chromatin groups
group_colors = c("Euchromatin" = "#f7941d",
                 "other-heterochromatin" = "#838687",
                 "Triple Heterochromatin" = "#662d91",
                 "H3K27me3" = "#d21f8a")


viab_colors = sequential_hcl(11, "Peach", rev = TRUE)
names(viab_colors) = seq(0, 1, by = 0.1)

annotation_colors_scr = list(
  chromatin = group_colors,
  target = c(HDAC="#60988D", HAT="#C6D0A8", Sirtuin="#8FCAAC",
             HMT="#FDCDAC", DNMT="#D69C81", `Histone Demethylase`="#FFF2AE",
             HIF="#E3DCD5", JAK="#C8D7E2", PIM="#9AB8C4", `Aurora Kinase`="#F4CAE4",
             PARP="#CB546F", `Epigenetic Reader Domain`= "#476D61", `DNA-PK`="#E07A43"),
  conc_char = conc_colors,
  clusters_drugs = sequential_hcl(7, "Plasma"),
  # viab_rep1 = c('#ffeee9', '#ffdcd4', '#ffcbbe', '#ffb8a8', '#fda693', '#f9947e', '#f4826a', '#ef6f56', '#e95b43', '#e24530'),
  # viab_rep2 = c('#ffeee9', '#ffdcd4', '#ffcbbe', '#ffb8a8', '#fda693', '#f9947e', '#f4826a', '#ef6f56', '#e95b43', '#e24530'),
  viab = viab_colors,
  ctrl_TIF = colorRampPalette(c("white", "black"))(10))

# Order the annotation legend
ordered_marks <- c("Euchromatin", 
                   "H3K27me3", 
                   "other-heterochromatin",  
                   "Triple Heterochromatin")

annotation_colors_scr$chromatin <- annotation_colors_scr$chromatin[ordered_marks]


# First calculate means and sd of the DMSO efficiency per replicate and barcode
dmso.eff.score <- indel.data %>%
  filter(drug == "DMSO") %>%
  dplyr::select('DMSO' = freqCut, well, barcode, replicate, tech, plate) %>%
  mutate(DMSO = ave(DMSO, well, replicate, tech, plate, FUN = function(x) mean(x, na.rm=T))) %>%
  distinct(replicate, tech, well, plate, DMSO) %>% 
  pull(DMSO) %>%
  fitdistr(., "normal") %>% tidy()
```

#### processing screen data for heatmaps
```{r filter heatmap}
# Generate heatmap: select top drugs (highest change at any IPR) per concentration
eff.data <- indel.data %>%
  #filter(target != "Aurora Kinase") %>%
  filter(sample == "Epigenetic Compound", indel_freq_filter, viab_reproducibility, !(freqCut_diff_mean < -0.1 & viab_mean < 0.5)) %>%
  # filter(abs(zscore_TIF_global_comb_rep) >= 2.58) %>%  # we can choose to display all drugs that have significant global hits
  distinct(conc_char, drug, target, IPR, zscore_TIF_comb_rep, freqCut_diff_mean, replicate, tech, filter_zscore_TIF_comb_rep)

eff.data.heatmap <-  eff.data %>%
  group_by(drug, conc_char) %>%
  # ave to add IPR grouping level to calculate mean of the replicates
  mutate(freqCut_diff_mean = ave((freqCut_diff_mean*100), IPR, FUN = function(x) mean(x, na.rm = TRUE)),
         ## Select drugs that have a large effect size in at least one location 
         ## (some drugs might have a small global effect size, but large local)
         zscore_TIF_comb_rep_temp = ifelse(!filter_zscore_TIF_comb_rep, 0, zscore_TIF_comb_rep)) %>%
  ungroup() %>%
  mutate(p_value = 2*pnorm(q = abs(zscore_TIF_comb_rep_temp), lower.tail = F),
         # Adjust p value for false discovery rate
         adj_pvalue = p.adjust(p_value, method = 'BH')) %>%
  # Pick the smallest per compound & concentration
  group_by(drug, conc_char) %>%
  mutate(min_score = min(adj_pvalue)) %>%
  # filter for adjusted p.val < 0.01 so we keep compounds with at least one signif effect.
  filter(min_score < 0.01) %>%
  distinct(drug, target, conc_char, freqCut_diff_mean, IPR, min_score) %>%
  ungroup()

#eff.data.heatmap2$freqCut_norm_mean[eff.data.heatmap2$mean_eff_zscore_comb > 0.05] <- NA
eff.data.heatmap2 <- eff.data.heatmap %>%
  dplyr::select(-min_score) %>%
  spread(IPR, freqCut_diff_mean) %>%
  filter(!is.nan(IPR3))

eff.data.heatmap2 <- eff.data.heatmap2[,c("drug", "target", "conc_char", iprs)]
eff.data.heatmap2$drug <- gsub(" [(].*[)]", "", eff.data.heatmap2$drug) #shorten name of drug
```

# CCD Data Screen
###Load data
```{r}
file = list.files(in.dir, pattern = "CCD_table", full.names = T) %>% tail(n = 1)
CCD_tib <- readRDS(file)
file = list.files(in.dir, pattern = "IPR_annotation_heatmap", full.names = T) %>% tail(n = 1)
chromatin_TIF <- readRDS(file)
file = list.files(in.dir, pattern = "compound_annotation_heatmap", full.names = T) %>% tail(n = 1)
target <- readRDS(file)
file = list.files(in.dir, pattern = "chrom_cluster_heatmap", full.names = T) %>% tail(n = 1)
TRIP_chrom_clust <- readRDS(file)
```

### Annotation

```{r heatmap annotation CCD}
colors_drugs <- c("#FA8D62", "#65BFA4", "#8CA0C4", "#808184", "#F6D289", "#E5E5E5", "#ACACAC", "#737373")
conc_control_colors = c("control" = colors_drugs[2], "100 nM" = colors_drugs[7], "1 µM" = colors_drugs[4], "10 µM" = "#000000")
conc_colors = c("100 nM" = colors_drugs[7], "1 µM" = colors_drugs[4], "10 µM" = "#000000")

group_colors = c("Euchromatin" = "#f7941d",
                 "other-heterochromatin" = "#838687",
                 "Triple Heterochromatin" = "#662d91",
                 "H3K27me3" = "#d21f8a")

viab_colors = sequential_hcl(11, "Peach", rev = TRUE)
names(viab_colors) = as.numeric(seq(0, 1, by = 0.1))

# Annotation colors heatmaps
annotation_colors_scr = list(
  chromatin = group_colors,
  target = c(HDAC="#60988D", HAT="#C6D0A8", Sirtuin="#8FCAAC",
             HMT="#FDCDAC", DNMT="#D69C81", `Histone Demethylase`="#FFF2AE",
             HIF="#E3DCD5", JAK="#C8D7E2", PIM="#9AB8C4", `Aurora Kinase`="#F4CAE4",
             PARP="#CB546F", `Epigenetic Reader Domain`= "#476D61", `DNA-PK`="#E07A43", 
             `Negative Control` = "#65BFA4", `MRN` = "#5D79AC"),
  conc_char = conc_colors,
  conc_char_ctrl = conc_control_colors,
  clusters_drugs = sequential_hcl(7, "Plasma"),
  # viab_rep1 = c('#ffeee9', '#ffdcd4', '#ffcbbe', '#ffb8a8', '#fda693', '#f9947e', '#f4826a', '#ef6f56', '#e95b43', '#e24530'),
  # viab_rep2 = c('#ffeee9', '#ffdcd4', '#ffcbbe', '#ffb8a8', '#fda693', '#f9947e', '#f4826a', '#ef6f56', '#e95b43', '#e24530'),
  viab = viab_colors,
  ctrl_TIF = colorRampPalette(c("white", "black"))(20),
  MMEJratio = c('#e1251b', '#d32c2d', '#c5313c', '#b6354a', '#a73857', '#963a65', '#843d72', '#6e3e7f', '#53408d', '#26419a'),
  group = c(insulator="#a6b5a3", repressive="#304441", euchromatin="#cda4cc",
            transcribing="#FBB040", HDAC="#aac9e0", accessibility='#cebc85',
            methylation='#7dc98f'))


ordered_marks2 <- c("Negative Control", "MRN", "DNA-PK","Epigenetic Reader Domain", "HDAC",  "Sirtuin", "HAT", 
                    "Histone Demethylase", "HMT", "DNMT", 
                    "HIF", "JAK", "PIM",
                    "Aurora Kinase", "PARP")

#Epistatic interaction colors
slope.colors <- tibble(color = c("#8c510a", "#f5f5f5","#01665e"),label = c("negative","none","positive"), feature = "epistasis")

annotation_colors_scr$target <- annotation_colors_scr$target[ordered_marks2]

clustering = data.frame(row.names=c('LMNB1', 'late_replicating', 'H3K9me2', 'H3K9me3', 'H3K27me3', 'EZH2', 
                                    'CTCF', 'SMC3',
                                    'HDAC3', 
                                    'H3K4me1', 'H3K4me2', 'H3K4me3', 'H3K27ac', 'H4K5acK8ac', 'H2AFZ', 
                                    'DNAse', 'Dam', 
                                    'm5C',
                                    'H3K79me2', 'TTseq',  'H3K36me3', 'POL2AS2','POL2'),
                        group=factor(c(rep('repressive', 6), 
                                       rep('insulator', 2),
                                       rep('HDAC', 1), 
                                       rep('euchromatin', 6),
                                       rep('accessibility', 2), 
                                       'methylation',
                                       rep('transcribing', 5)),
                                     levels=c('transcribing', 'accessibility',
                                              'methylation', 'euchromatin',
                                              'HDAC', 'insulator', 'repressive')))
```


### Synergies Heatmaps

```{r Synergies chromatin prep, fig.width= 10, fig.height= 20}
all_compounds_epistasis_TIF = CCD_tib %>%
  dplyr::select(drug_conc, indelrate_slope_plot, feature) %>%
  group_by(drug_conc) %>%
  mutate(onlyNA = sum(is.na(indelrate_slope_plot))) %>% 
  filter(onlyNA != 25) %>%
  reshape2::dcast(drug_conc ~ feature, value.var = "indelrate_slope_plot", fill = 0) %>% 
  column_to_rownames(var = "drug_conc")

target_ctrl = dplyr::select(target, -conc_char)

quant_breaks <- quantile(all_compounds_epistasis_TIF, c(.01,.99), na.rm = T)
palette_length <- 100
breaks <- c(seq(quant_breaks[1], 0, length.out=ceiling(palette_length/2) + 1), 
            seq(quant_breaks[2]/palette_length, quant_breaks[2], length.out=floor(palette_length/2)))
colors = colorRampPalette(slope.colors %>% pull(color))(length(breaks))
colors[which(quant_breaks == 0)-1] = "gray90"
```

---

# Main figures

## 3ABC CCD Examples
```{r 3A CCD examples, fig.width= 12, fig.height= 4}
CCD_example_tib = indel.data %>% 
  mutate(drug = gsub(" [(].*[)]", "", drug),
         drug_conc = paste(drug, conc_char, sep = " ")) %>%
  filter(drug_conc %in% c("EPZ-6438 10 µM", "Decitabine 1 µM", "Scriptaid 1 µM", "Apicidin 100 nM", "DMSO control", "SNS-314 Mesylate 1 µM")) %>%
  distinct(IPR, drug, freqCut, H3K27me3.zscore, LMNB1.zscore, m5C.zscore, tech, replicate, H3K36me3.zscore) %>%
  mutate(freqCut_mean = ave(freqCut, IPR, drug, FUN = mean)) %>%
  distinct(IPR, drug, freqCut_mean, H3K27me3.zscore, LMNB1.zscore, m5C.zscore, H3K36me3.zscore) %>%
  pivot_wider(names_from = drug, values_from = freqCut_mean) %>%
  mutate(Scriptaid_log2_FC_TIF = log2(Scriptaid/DMSO),
         Api_log2_FC_TIF = log2(Apicidin/DMSO),
         EPZ6438_log2_FC_TIF = log2(`EPZ-6438`/DMSO),
         SNSMesyl_log2_FC_TIF = log2(`SNS-314 Mesylate`/DMSO),
         Decitabine_log2_FC_TIF = log2(Decitabine/DMSO))

#Create mock examples
mock.slopes <- tibble(z_score = seq(-3,3,0.25), A = z_score*0 + 0.6,B = z_score*0.1 +0.6, C = z_score*-0.1 - 0.6) %>% reshape2::melt(id.vars = "z_score")

#Plot mock examples
p1 = ggplot(mock.slopes, aes(z_score,value,color = variable)) + theme_bw() + 
  geom_smooth() +
  geom_vline(xintercept = 0, linetype = 2) + 
  geom_hline(yintercept = 0, linetype = 2) + 
  ylab("log2(fold change TIF)") + 
  xlab("Feature signal (z-score)") + 
  theme(legend.position = "none") + 
  coord_cartesian(ylim= c(-1.5,1.5)) + 
  scale_color_manual(values = c("black","#01665e","#8c510a"))+
  theme_classic2(base_size = 16)

# p2 = ggplot(CCD_example_tib,
#        aes(x = m5C.zscore, y = Decitabine_log2_FC_TIF)) +
#   geom_point() +
#   geom_smooth(method = "lm", color = "#01665e",se = FALSE) +
#   stat_cor(method = "pearson") +
#   xlab("m5C z-score") +
#   ylab("delta log2(TIF)") + 
#   coord_cartesian()
# 
# CCD_tib %>% filter(drug_conc == "Decitabine 1 µM" & feature == "m5C")

p2 = ggplot(CCD_example_tib,
            aes(x = H3K27me3.zscore, y = EPZ6438_log2_FC_TIF)) +
  geom_point() +
  geom_smooth(method = "lm", color = "#01665e",se = FALSE) +
  stat_cor(method = "pearson") +
  xlab("H3K27me3 z-score") +
  ylab("delta log2(TIF)") + 
  coord_cartesian()+
  theme_classic2(base_size = 16)

CCD_tib %>% filter(drug_conc == "EPZ-6438 10 µM" & feature == "H3K27me3")


p3 = ggplot(CCD_example_tib, 
            aes(x = H3K36me3.zscore, y = SNSMesyl_log2_FC_TIF)) + 
  geom_point() +
  geom_smooth(method = "lm", color = "black",se = FALSE) +
  stat_cor(method = "pearson") +
  xlab("H3K36me3 z-score") +
  ylab("delta log2(TIF)") + 
  coord_cartesian()+
  theme_classic2(base_size = 16)

plot_grid(p1, p2, p3, nrow = 1, labels = c("A", "B", "C"))
```


## 3D All compound synergies with chromatin
```{r 3D Synergies chromatin - TIF all, fig.width= 10, fig.height= 20}
# change these numbers according to where you want to set the cutoff
quant_breaks <- quantile(all_compounds_epistasis_TIF, c(.01,.99), na.rm = T)
palette_length <- 100
breaks <- c(seq(quant_breaks[1], 0, length.out=ceiling(palette_length/2) + 1), 
            seq(quant_breaks[2]/palette_length, quant_breaks[2], length.out=floor(palette_length/2)))
colors = colorRampPalette(slope.colors %>% pull(color))(length(breaks))
colors[which(quant_breaks == 0)-1] = "gray90"

pheatmap(all_compounds_epistasis_TIF[, TRIP_chrom_clust$labels],
         border_color = F,
         cellwidth = 8,
         cellheight = 8,
         breaks = breaks,
         cluster_cols = TRIP_chrom_clust,
         clustering_method = "ward.D",
         annotation_row = target_ctrl,
         main = "Chromatin synergies - TIF all",
         color = colors,
         annotation_colors = annotation_colors_scr, 
         annotation_legend = T)
```

```{r top CCDs for all chromatin features, fig.width= 6, fig.height= 6, eval = FALSE}
for(i in unique(CCD_tib$feature)) {
  rank_dt = CCD_tib %>% 
    left_join(rownames_to_column(target, var = "drug_conc"), by = "drug_conc") %>%
    filter(feature == i, !is.na(indelrate_slope_plot)) %>%
    distinct() %>%
    arrange(desc(indelrate_slope_plot))
  
  if(nrow(rank_dt) > 1) {
    rank_dt %<>%
      mutate(rank = seq(1:n()))
    
    maxrank = nrow(rank_dt)
    
    p1 = rank_dt %>% 
      ggplot(., aes(x = rank, fill = target)) +
      geom_bar() + 
      scale_x_reverse() +
      coord_flip() +
      xlab("Rank") +
      theme_classic2(base_size = 16) +
      scale_fill_manual(values = colors_targets) +
      # geom_label(aes(label = ifelse(width == 2, drug_conc,''))) +
      theme(legend.position = "none",
            axis.title.x=element_blank(),
            axis.text.x=element_blank(),
            axis.ticks.x=element_blank())+
      # add horizontal line for 0.01 slope cutoff which is highlighted in the heatmap
      geom_vline(xintercept = rank_dt[rank_dt$indelrate_slope_plot  < -0.01, ]$rank[1] - 0.5)  +
      geom_vline(xintercept = max(rank_dt[rank_dt$indelrate_slope_plot  > 0.01, ]$rank) + 0.5)
    
    
    p2 = rank_dt %>% 
      filter(!between(rank, 6, maxrank-5)) %>%
      mutate(rank = seq(1:n())) %>% 
      ggplot(., aes(x = rank, y = indelrate_slope_plot, fill = target)) +
      geom_bar(stat = "identity") + 
      scale_x_reverse() +
      coord_flip() +
      ylab("Synergy score") +
      theme_classic2(base_size = 16) +
      scale_fill_manual(values = colors_targets) +
      # geom_label(aes(label = ifelse(width == 2, drug_conc,''))) +
      geom_text(aes(label=drug_conc)) +
      theme(legend.position = "none",
            axis.title.y=element_blank(),
            axis.text.y=element_blank(),
            axis.ticks.y=element_blank()) +
      ggtitle(paste0("Top synergies with: ", i))
    
    
    p = plot_grid(p1, p2, labels = c("E", "F"), rel_widths = c(1, 5))
    print(p)
  }
}
```


# Supplementary Figures
## S3AB Example CCD
```{r S3AB example of filtered out CCDs,  fig.width = 8, fig.height = 3}

#Create mock examples
mock.slopes <- tibble(z_score = seq(-3,3,0.25), 
                      A = z_score*0.1 -0.6, 
                      B = z_score*-0.1 +0.6) %>% 
  reshape2::melt(id.vars = "z_score")

#Plot mock examples
p1 = ggplot(mock.slopes, aes(z_score, value, color = variable)) + 
  geom_smooth() +
  geom_vline(xintercept = 0, linetype = 2) + 
  geom_hline(yintercept = 0, linetype = 2) + 
  ylab("log2(fold change TIF)") + 
  xlab("Feature signal (z-score)") + 
  coord_cartesian(ylim= c(-1.5,1.5)) + 
  scale_color_manual(values = c("#01665e","#8c510a"))+
  theme_classic2(base_size = 16) + 
  theme(legend.position = "none")

p2 = ggplot(CCD_example_tib,
            aes(x = H3K36me3.zscore, y = EPZ6438_log2_FC_TIF)) +
  geom_point() +
  geom_smooth(method = "lm", color = "#8c510a",se = FALSE) +
  stat_cor(method = "pearson") +
  geom_vline(xintercept = 0, linetype = 2) + 
  geom_hline(yintercept = 0, linetype = 2) + 
  xlab("H3K27me3 z-score") +
  ylab("delta log2(TIF)") + 
  coord_cartesian() +
  theme_classic2(base_size = 16)

plot_grid(p1, p2, nrow = 1, labels = c("C", "D"))

CCD_tib %>% filter(feature == "H3K36me3" & drug_conc == "EPZ-6438 10 µM") %>% pull(indelrate_slope_plot)
```

```{r}

```



```{r eval=FALSE}
indel.data %>% filter(!between(zscore_TIF_comb_rep, -3.2905, 3.2905) & conc_char != "control" & viab_reproducibility & cor_TIF_filter & filter_zscore_TIF_comb_rep) %>% 
  # distinct(IPR, drug, replicate, target, tech, conc_char, zscore_TIF, zscore_TIF_comb_tech, zscore_TIF_comb_rep) %>% 
  distinct(IPR, drug, conc_char, target, zscore_TIF_comb_rep) %>% 
  ggplot(., aes(IPR, zscore_TIF_comb_rep, color = target)) + 
  geom_quasirandom() + 
  facet_grid(. ~ conc_char) + 
  scale_color_manual(values = colors_targets) +
  theme_classic2(base_size = 16)
```

# Session Info
```{r}
paste("Run time: ",format(Sys.time()-StartTime))
getwd()
date()
sessionInfo()
```

