---
title: "DSB Episcreen data analysis & plotting"
author: 
  - name: "Max Trauernicht"
    email: "m.trauernicht@nki.nl"
    affiliation: "Netherlands Cancer Institute"
date: '`r format(Sys.time(), "%d/%m/%Y")`'
output: 
  html_document:
    theme: united
    highlight: pygments
    fig_caption: yes
    code_folding: hide
    df_print: kable
    toc: true
    toc_depth: 4
    toc_float:
      collapsed: false
---

*knitr document van Steensel lab*


# Introduction
Clone 5 with ~18 sgRNA target site IPRs was cut and treated with 160 epigenetic drugs. The repair outcomes of that experiment will be visualized here. I previously processed the raw sequencing data, and calculated the repair outcomes. In this script, the main results plots will be generated.


## Setup {.tabset}

<!-- little HTML script to do indentation of the table of contents -->
<script>
$(document).ready(function() {
$items = $('div#TOC li');
$items.each(function(idx) {
num_ul = $(this).parentsUntil('#TOC').length;
$(this).css({'text-indent': num_ul * 10, 'padding-left': 0});
});

});
</script>

```{css, echo = FALSE}
div.sourceCode {
overflow-x: hidden;
}
```
### Libraries
### Libraries

```{r load_packages, results = 'hide', warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
StartTime <-Sys.time()

# 8-digit Date tag:
Date <- substr(gsub("-","",Sys.time()),1,8) 
# libraries:
library(knitr)
library(tidyverse)
library(outliers)
library(ggplot2)
library(ggbeeswarm)
library(data.table)
# library(ggrepel)
library(GGally)
library(MASS)
# library(DESeq2)
library(gghighlight)
library(ggrepel)
# library(platetools)
library(ggpubr)
# library(Laurae)
library(pheatmap)
library(RColorBrewer)
# # library(ggrastr)
library(broom)
# library(ggsignif)
# library(rstatix)
library(colorspace)
library(analogue)

## Select outdir
out.dir = paste0("./figures/", Date, "/")
dir.create(out.dir)
in.dir = list.dirs(path = "/DATA/projects/DSBrepair/data/R") %>% .[grepl("episcreen", .)] %>% tail(n = 1)
opts_chunk$set(dev=c('png', 'pdf'), fig.path = file.path(out.dir))
pdf.options(useDingbats = FALSE)
```

### Functions

```{r load_functions, echo = FALSE, warning = FALSE}

p.adjust <- function(p, method = p.adjust.methods, n = length(p))
{
  ## Methods 'Hommel', 'BH', 'BY' and speed improvements
  ## contributed by Gordon Smyth
  method <- match.arg(method)
  if(method == "fdr") method <- "BH"	# back compatibility
  nm <- names(p)
  p <- as.numeric(p)
  p0 <- setNames(p, nm)
  if(all(nna <- !is.na(p))) nna <- TRUE
  p <- p[nna]
  lp <- length(p)
  # stopifnot(n >= lp)
  if (n <= 1) return(p0)
  if (n == 2 && method == "hommel") method <- "hochberg"
  
  p0[nna] <-
    switch(method,
           bonferroni = pmin(1, n * p),
           holm = {
             i <- seq_len(lp)
             o <- order(p)
             ro <- order(o)
             pmin(1, cummax( (n - i + 1L) * p[o] ))[ro]
           },
           hommel = { ## needs n-1 >= 2 in for() below
             if(n > lp) p <- c(p, rep.int(1, n-lp))
             i <- seq_len(n)
             o <- order(p)
             p <- p[o]
             ro <- order(o)
             q <- pa <- rep.int( min(n*p/i), n)
             for (j in (n-1):2) {
               ij <- seq_len(n-j+1)
               i2 <- (n-j+2):n
               q1 <- min(j*p[i2]/(2:j))
               q[ij] <- pmin(j*p[ij], q1)
               q[i2] <- q[n-j+1]
               pa <- pmax(pa,q)
             }
             pmax(pa,p)[if(lp < n) ro[1:lp] else ro]
           },
           hochberg = {
             i <- lp:1L
             o <- order(p, decreasing = TRUE)
             ro <- order(o)
             pmin(1, cummin( (n - i + 1L) * p[o] ))[ro]
           },
           BH = {
             i <- lp:1L
             o <- order(p, decreasing = TRUE)
             ro <- order(o)
             pmin(1, cummin( n / i * p[o] ))[ro]
           },
           BY = {
             i <- lp:1L
             o <- order(p, decreasing = TRUE)
             ro <- order(o)
             q <- sum(1L/(1L:n))
             pmin(1, cummin(q * n / i * p[o]))[ro]
           },
           none = p)
  p0
}

CorrelationPlotMean <- function(data, feature_name, targets, title = "with treatment", p.value=0.001, alfa = c(0.5, 1)){
  cor.df = data.frame()
  for (i in feature_name){
    corr.list <- data %>% dplyr::select(i, as.character(targets)) %>% as.matrix() %>% rcorr(type = "spearman")
    
    df <- data.frame(rowname = names(corr.list$r[1, -1]),
                     corr = corr.list$r[1, -1],
                     pvalue = corr.list$P[1, -1],
                     feature = i)
    cor.df <- rbind(cor.df, df)
  }
  p <- cor.df %>%
    transform(rowname = factor(rowname, levels = levels(targets)),
              corr = corr,
              pvalue = pvalue) %>%
    ggplot(aes(x = rowname, y = corr, fill = pvalue < p.value)) +
    geom_bar(stat = "identity", position = "dodge") +
    ylab(paste("Spearman correlation ",title)) +
    scale_fill_manual(values = c("grey60", "black")) +
    xlab("Chromatin Feature") +
    # ggtitle(paste("Repair pathways correlation with chromatin features in", condition_name, "exps")) +
    scale_y_continuous(breaks=seq(-1,1, .1)) +
    coord_flip()
  return(list(dataframe=cor.df, plot = p))
}

SetFileName <- function(filename, initials) {
  # Set filename with extension and initials to make filename with date integrated.
  filename <- substitute(filename)
  initials <- substitute(initials)
  filename <- paste0(initials, Date, filename)
  filename
}

plot_comb_grid_beeswarm <- function(dt, opt_df, domains, y_name, min_count=0,
                                    max_count=Inf, group_by=NULL, color=NULL, y_plabel=c(-1.5, 1.5)){
  
  opt_melt = reshape2::melt(opt_df)
  colnames(opt_melt) = c('group', 'domain', 'is_in')
  
  opt_melt$is_in = ifelse(opt_melt$is_in, 'in', 'out')
  opt_melt$domain = factor(opt_melt$domain, levels=domains)
  
  
  group_list = lapply(rownames(opt_df), function(opt_name){
    opt = opt_df[opt_name, ]
    in_vec = domains[opt]
    out_vec = domains[!opt]
    
    in_sum = rowSums(dt[,..in_vec])
    out_sum = rowSums(dt[,..out_vec])
    if (length(in_vec)==0){
      in_group = which(out_sum == 0)
    } else if (length(out_vec) == 0){
      in_group = which(in_sum==length(in_vec))
    } else  {
      in_group = which(in_sum == length(in_vec) & out_sum==0)
    }
    return(in_group)
  })
  
  names(group_list) = rownames(opt_df)
  for (opt_name in names(group_list)){
    i_vec = group_list[[opt_name]]
    dt[i_vec, group := opt_name]
  }
  if (is.null(group_by)){
    group_count_df = data.frame(dt[,table(group)])
  } else {
    group_count_df = data.frame(dt[,table(group)/length(unique(get(group_by)))])
  }
  group_count_df = group_count_df[order(match(group_count_df$group, rownames(opt_df))),]
  # print(group_count_df)
  group_count_sel = group_count_df[group_count_df$Freq>=min_count,]
  group_count_sel = group_count_sel[group_count_sel$Freq<=max_count,]
  group_levels =  group_count_sel$group
  opt_sel = opt_melt[opt_melt$group %in% group_levels, ]
  
  opt_sel$group = factor(opt_sel$group, levels=group_levels)
  # print(opt_sel)
  
  indel_selection = dt[group%in%group_levels, ]
  indel_selection[,group:=factor(group, levels=group_levels)]
  
  # wilcox_list = lapply(group_levels[group_levels!='iDomain'],
  #                      function(g){
  #                          x = indel_selection[group==g, y_name, with=F]
  #                          y = indel_selection[group=='iDomain', y_name,
  #                                              with=F]
  #                          wilcox.test(unlist(x),unlist(y))
  #                      })
  
  group_count_sel$group = factor(group_count_sel$group, levels=group_levels)
  
  count_plot = ggplot(group_count_sel, aes(x=group,y=1, label=Freq)) +
    geom_label() +
    theme_void()
  
  
  
  dots = ggplot(opt_sel, aes(x=group, y=domain, color=is_in)) +
    geom_point(size=2) +
    geom_line(data=opt_sel[opt_sel$is_in=='in',], aes(group=group)) +
    scale_color_manual(values=c('in'='black', 'out'='grey70')) +
    guides(color="none")  +
    geom_vline(xintercept=seq(1.5, length(unique(dt$group))-0.5, 1),
               lwd=0.5, colour="grey90") +
    theme(axis.title=element_blank(),
          axis.text=element_blank(),
          axis.ticks=element_blank(),
          panel.grid.major = element_blank())
  
  symnum = list(cutpoints = c(0, 0.0001, 0.001, 0.01, 0.05, 1),
                symbols = c("****", "***", "**", "*", " "))
  
  
  if (is.null(group_by)){
    beeswarm = ggplot(indel_selection, aes_string(x='group', y=y_name)) +
      geom_quasirandom(width=0.3) +
      stat_summary(fun=median, fun.min = median,
                   fun.max = median,
                   geom="crossbar", width = 0.5,
                   color='red') +
      stat_compare_means(aes(label = ..p.signif..),
                         symnum.args = symnum,
                         method = 'wilcox.test',
                         ref.group='iDomain')
    
  } else {
    wilcox_dt = indel_selection[,test_wilcox(.SD, group_by, y_name),
                                by='group']
    wilcox_dt[,p_adj:=p.adjust(p_value)]
    wilcox_dt[,p_signif:=symnum(p_adj, cutpoints=symnum$cutpoints,
                                symbols=symnum$symbols)]
    
    
    beeswarm = ggplot(indel_selection,
                      aes_string(x='group', y=y_name, color=group_by)) +
      geom_quasirandom(dodge.width=1) +
      scale_color_manual(values=color) +
      stat_summary(aes_string(group=group_by), fun=median,
                   fun.min = median, fun.max = median,
                   geom="crossbar", width = 0.5,
                   color='red',
                   position= position_dodge(width =1)) +
      geom_text(data=wilcox_dt, aes(x=group, y=0.8, label=p_signif),
                inherit.aes=F)
    print(wilcox_dt)
  }
  beeswarm = beeswarm +
    geom_vline(xintercept=seq(1.5, length(unique(dt$group))-0.5, 1),
               lwd=0.5, colour="grey90") +
    theme(axis.title=element_blank(),
          axis.text.x=element_blank(),
          panel.grid.major.x = element_blank(),
          axis.ticks.x=element_blank()) +
    ylim(y_plabel)
  
  count = data.frame(domain=gsub('_domain', '', domains),
                     count=colSums(dt[,..domains], na.rm=T))
  
  count$domain = factor(count$domain, levels=domain_levels)
  
  histogram = ggplot(count, aes(x=1, y=domain, label = count)) +
    geom_label() +
    theme(plot.background = element_blank(),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.border = element_blank(),
          axis.title=element_blank(),
          axis.text.x=element_blank(),
          axis.ticks=element_blank())
  
  empty = ggplot(data=data.frame(1,1),aes(x=1,y=1)) +
    geom_point() +
    theme_void()
  
  count_beeswarm = plot_grid(beeswarm, count_plot, ncol=1, rel_heights=c(20,1),
                             align='v', axis='lr')
  return(plot_grid(empty, beeswarm, empty, count_plot,
                   histogram, dots, nrow=3,
                   rel_widths=c(1,10), rel_heights=c(15,1,5), align='vh',
                   axis='tlbr'))
}


get_combinations <- function(domain_vec){
  n_domain = length(domain_vec)
  get_ivec <-function (last_x, n_domain){
    if (last_x < n_domain){
      i_vec = c(0, c((last_x+1):n_domain))
    } else{
      i_vec = c(0)
    }
  }
  
  combine <- function(x_new, x_vec, n_domain){
    new_vec = c(x_vec, x_new)
    if (length(new_vec)==n_domain){
      return(new_vec)
    } else {
      i_vec = get_ivec(max(new_vec), n_domain)
      x_list = lapply(i_vec, combine, x_vec=new_vec,
                      n_domain=n_domain)
      return(do.call(rbind, x_list))
    }
  }
  
  comb_list = lapply(c(0:n_domain), combine, x_vec=c(), n_domain=n_domain)
  
  comb = do.call(rbind, comb_list)
  comb_df = unique(t(apply(comb, 1, function(x){x[order(x)]})))
  
  name_list = apply(comb_df, 1, function(x){
    name = paste(domain_vec[x], collapse='-')
    if (name == ''){
      name = 'iDomain'
    }
    return(name)
  })
  
  opt_list = lapply(domain_vec, function(d){
    apply(comb_df, 1, function(x){
      opt = domain_vec[x]
      return(d%in%opt)
    })
  })
  
  opt_df = do.call(cbind, opt_list)
  colnames(opt_df) = domain_vec
  rownames(opt_df) = name_list
  
  return(opt_df)
}

# From Fede:
# ggpairs custom functions
corColor <- function(data, mapping, color = I("black"), sizeRange = c(1, 3), ...) {
  
  x   <- eval_data_col(data, mapping$x)
  y   <- eval_data_col(data, mapping$y)
  r   <- cor(x, y)
  rt  <- format(r, digits = 3)
  tt  <- as.character(rt)
  cex <- max(sizeRange)
  
  # helper function to calculate a useable size
  percent_of_range <- function(percent, range) {
    percent * diff(range) + min(range, na.rm = TRUE)
  }
  
  # plot correlation coefficient
  p <- ggally_text(label = tt, mapping = aes(), xP = 0.5, yP = 0.5,
                   size = I(percent_of_range(cex * abs(r), sizeRange)), color = color, ...) +
    theme(panel.grid.minor=element_blank(),
          panel.grid.major=element_blank())
  
  corColors <- RColorBrewer::brewer.pal(n = 7, name = "RdYlBu")[2:6]
  
  if (r <= boundaries[1]) {
    corCol <- corColors[1]
  } else if (r <= boundaries[2]) {
    corCol <- corColors[2]
  } else if (r < boundaries[3]) {
    corCol <- corColors[3]
  } else if (r < boundaries[4]) {
    corCol <- corColors[4]
  } else {
    corCol <- corColors[5]
  }
  
  p <- p +
    theme(panel.background = element_rect(fill = corCol))
  
  return(p)
}


theme_classic_lines <- function() {
  theme_pubr(border = T, legend = "top") +
    theme(panel.grid.major = element_line(colour = "#adb5bd", size = 0.25),
          strip.background = element_rect(fill = "white", color = "white")
    )
}
theme_classic_lines_45 <- function() {
  theme_pubr(border = T, legend = "top", x.text.angle = 45) +
    theme(panel.grid.major = element_line(colour = "#adb5bd", size = 0.25),
          strip.background = element_rect(fill = "white", color = "white")
    )
}

theme_classic_lines_90 <- function() {
  theme_pubr(border = T, legend = "top", x.text.angle = 90) +
    theme(panel.grid.major = element_line(colour = "#adb5bd", size = 0.25),
          strip.background = element_rect(fill = "white", color = "white")
    )
}

theme_classic_lines_90_noborder <- function() {
  theme_pubr(border = F, legend = "top", x.text.angle = 90) +
    theme(strip.background = element_rect(fill = "white", color = "white")
    )
}

theme_set(theme_classic_lines())

colors_diverse <- c("#264653", "#9AC1AE", "#5D987B", "#f2cc8f", "#e76f51")
colors_drugs <- c("#FA8D62", "#65BFA4", "#8CA0C4", "#808184", "#F6D289", "#E5E5E5", "#ACACAC", "#737373")
#colors_diverse <- c("#264653", "#2a9d8f", "#e9c46a", "#f4a261", "#e76f51") 
colors_targets <-  c(HDAC="#60988D", HAT="#C6D0A8", Sirtuin="#8FCAAC",
                     HMT="#FDCDAC", DNMT="#D69C81", `Histone Demethylase`="#FFF2AE",
                     HIF="#E3DCD5", JAK="#C8D7E2", PIM="#9AB8C4", `Aurora Kinase`="#F4CAE4",
                     PARP="#CB546F", `Epigenetic Reader Domain`= "#476D61", `DNA-PK`="#E07A43")
colors_pathways = c("wt" = "#808184", "other" = "black", "NHEJ" = "#e1251b", "MMEJ" = "#26419a", "SSTR" = "#007A4C")

conc_colors = c("100 nM" = colors_drugs[7], "1 µM" = colors_drugs[4], "10 µM" = "#000000")



ggplot_custom <- function(...) ggplot2::ggplot(...) + 
  scale_color_manual(values = colors_diverse) + 
  scale_fill_manual(values = colors_diverse)

sigFunc = function(x){
  if(x < 0.001){"***"} 
  else if(x < 0.01){"**"}
  else if(x < 0.05){"*"}
  else{NA}}

```


### Data import

```{r data import, echo = FALSE, warning = FALSE}
# # Import data from preprocessing script
# in.dir = list.dirs(path = "/DATA/projects/DSBrepair/data/R") %>% .[grepl("episcreen", .)] %>% tail(n = 1)
# file = list.files(in.dir, pattern = "episcreen_ratios", full.names = T) %>% tail(n = 1)
# indel.data <- readRDS(file) # %>% filter(pathway_bal_filter == TRUE & viab_reproducibility == TRUE)
indel.data <- readRDS("data/episcreen_ratios_mt20220811.RDS") 

# Import data from preprocessing validation script
file.validation = list.files(in.dir, pattern = "validation_ratio", full.names = T) %>% tail(n = 1)
indel.data.validation <- readRDS(file.validation)

chromatin.features <- colnames(indel.data)[grepl("[.]zscore", colnames(indel.data))]
chrom_data = indel.data %>% dplyr::select(barcode, all_of(chromatin.features)) %>% distinct()  %>%
  rename_with(., ~gsub(".zscore", "", .x))

chromatin.features.short = gsub(".zscore", "", chromatin.features)

# List chromatin name columns from this table
chromatin_cols = colnames(indel.data.validation)[42:75]
# Import data from DSB TRIP paper
RSTP2_IndelRatios_Chromatin_2kb <- readRDS("/DATA/usr/r.schep/DSBrepair/data/R/rs20210311/RSTP2_IndelRatios_Chromatin_2kb.RDS") %>% 
  filter(replicate == "mean" & drug %in% c("DMSO_GSK", "GSK126") & plasmid == "LBR2") %>%
  dplyr::select(barcode, drug, MMEJ, NHEJ, efficiency, all_of(chromatin_cols)) %>%
  distinct() %>%
  mutate(MMEJ_NHEJ = MMEJ/NHEJ) %>%
  dplyr::select(-NHEJ, -MMEJ)
```

## Prepare heatmaps
```{r prepare data for heatmaps}
# Order IPRs by their starting freqCut value
dmso_freq <- indel.data %>%
  filter(drug == "DMSO") %>%
  distinct(IPR, MMEJratio) %>%
  mutate(MMEJratio = ave(MMEJratio, IPR, FUN = function(x) mean(x, na.rm = TRUE))*100) %>%
  dplyr::select('ctrl_MMEJratio' = MMEJratio, IPR) %>%
  distinct() %>%
  arrange(desc(ctrl_MMEJratio))

iprs <- dmso_freq$IPR
iprs <- as.character(iprs)


# Set annotation for heatmap
# viability
viab_drug = indel.data %>% filter(drug != "PAO") %>% 
  mutate(drug = gsub(" [(].*[)]", "", drug),
         drug_conc = paste(drug, conc_char, sep = " ")) %>% 
  distinct(viab_mean, drug_conc, replicate) %>% 
  group_by(drug_conc) %>%
  summarise(viab = mean(viab_mean))
# Row annotation: annotate drugs with the target group
target_set_drugs <- indel.data %>% 
  filter(!target %in% c("Negative Control", "MRN", "DNA-PK")) %>%
  mutate(drug = gsub(" [(].*[)]", "", drug),
         drug_conc = paste(drug, conc_char, sep = " ")) %>%
  left_join(viab_drug) %>%
  distinct(drug, drug_conc, conc_char, target, viab) %>% 
  column_to_rownames(var="drug_conc")

ctrl.vals <- indel.data %>%
  filter(drug == "DMSO") %>%
  distinct(IPR, freqCut, MMEJratio) %>%
  mutate(ctrl_TIF = ave(freqCut, IPR, FUN = function(x) mean(x, na.rm = TRUE))*100,
         ctrl_balance = ave(MMEJratio, IPR, FUN = function(x) mean(x, na.rm = TRUE))*100) %>%
  distinct(ctrl_TIF, 
           ctrl_balance,
           IPR)

clone5.matrix = indel.data %>% 
  dplyr::select(IPR, ends_with(".zscore")) %>% 
  distinct() %>% 
  rename_with(., ~gsub(".zscore", "", .x)) %>%
  # left_join(ctrl.vals) %>%
  column_to_rownames(var="IPR")

clone5_map_ipr = hclust(dist(clone5.matrix), method = "ward.D") %>% 
  # Rotate the middle node to math better the TIF in CTRL conditions
  dendextend::rotate(c(1:6, 10:7, 11:18))

# Chromatin annotation
chromatin <- readRDS(
  "/DATA/usr/x.vergara/XV_ChIPIndels/XV20200902_DDR_RS_revision/XV20200902_DDR_RS_revision/data/xv20200915_DDR_data_CHIP.rds") %>% 
  distinct(barcode, chromatin) %>%
  left_join(distinct(indel.data, barcode, IPR)) %>%
  left_join(dmso_freq)

# chromatin$IPR[is.na(chromatin$IPR)] <- "IPR1"
chromatin <- chromatin %>%
  filter(IPR != "<NA>") %>% 
  dplyr::select(-barcode) %>%
  column_to_rownames(var="IPR")

## Chromatin groups
group_colors = c("Euchromatin" = "#f7941d",
                 "other-heterochromatin" = "#838687",
                 "Triple Heterochromatin" = "#662d91",
                 "H3K27me3" = "#d21f8a")


viab_colors = sequential_hcl(11, "Peach", rev = TRUE)
names(viab_colors) = seq(0, 1, by = 0.1)

annotation_colors_scr = list(
  chromatin = group_colors,
  target = c(HDAC="#60988D", HAT="#C6D0A8", Sirtuin="#8FCAAC",
             HMT="#FDCDAC", DNMT="#D69C81", `Histone Demethylase`="#FFF2AE",
             HIF="#E3DCD5", JAK="#C8D7E2", PIM="#9AB8C4", `Aurora Kinase`="#F4CAE4",
             PARP="#CB546F", `Epigenetic Reader Domain`= "#476D61", `DNA-PK`="#E07A43"),
  conc_char = conc_colors,
  clusters_drugs = sequential_hcl(4, "YlGn"),
  # viab_rep1 = c('#ffeee9', '#ffdcd4', '#ffcbbe', '#ffb8a8', '#fda693', '#f9947e', '#f4826a', '#ef6f56', '#e95b43', '#e24530'),
  # viab_rep2 = c('#ffeee9', '#ffdcd4', '#ffcbbe', '#ffb8a8', '#fda693', '#f9947e', '#f4826a', '#ef6f56', '#e95b43', '#e24530'),
  viab = viab_colors,
  ctrl_MMEJratio = colorRampPalette(c("white", "black"))(10))

# Order the annotation legend
ordered_marks <- c("Euchromatin", 
                   "H3K27me3", 
                   "other-heterochromatin",  
                   "Triple Heterochromatin")

annotation_colors_scr$chromatin <- annotation_colors_scr$chromatin[ordered_marks]

# Order IPRs by their starting MMEJ_log2_FC value
dmso_bal <- indel.data %>%
  filter(drug == "DMSO") %>%
  distinct(IPR, MMEJ_log2_FC) %>%
  mutate(MMEJ_log2_FC = ave(MMEJ_log2_FC, IPR, FUN = function(x) mean(x, na.rm = TRUE))*100) %>%
  dplyr::select('dmso_bal' = MMEJ_log2_FC, IPR) %>%
  distinct() %>%
  arrange(desc(dmso_bal))

iprs <- dmso_bal$IPR
iprs <- as.character(iprs)
```


## 4A MMEJ ratio change heatmap
Aim: Give an overview of the most important MMEJ ratio fold-changes. 
```{r prepare data for heatmaps}
# Generate heatmap: select top drugs (highest change at any IPR) per concentration
balance.data <- indel.data %>%
  #filter(target != "Aurora Kinase") %>%
  filter(sample == "Epigenetic Compound") %>% 
  filter(pathway_bal_filter == TRUE & viab_reproducibility == TRUE & cor_bal_filter) %>%
  filter(abs(zscore_bal_global_comb_rep) >= 2.58) %>%  # we can choose to display all drugs that have significant global hits
  distinct(conc_char, drug, target, IPR, zscore_bal_comb_rep, MMEJ_log2_FC, replicate, tech)

dmso_bal <- indel.data %>%
  filter(drug == "DMSO") %>%
  dplyr::select('dmso_bal' = MMEJ_log2_FC , IPR) %>%
  mutate(dmso_bal = ave(dmso_bal, IPR, FUN = function(x) mean(x, na.rm = TRUE))) %>%
  distinct()

balance.data.heatmap <- balance.data %>%
  # left_join(dmso_bal) %>%
  group_by(drug, conc_char) %>%
  # ave to add IPR grouping level to calculate mean of the replicates
  mutate(MMEJ_log2_FC_mean = ave(MMEJ_log2_FC, IPR, FUN = function(x) mean(x, na.rm = TRUE)),
         ## Select drugs that have a large effect size in at least one location 
         ## (some drugs might have a small global effect size, but large local)
         p_value = 2*pnorm(q = abs(zscore_bal_comb_rep), lower.tail = F),
         # Adjust p value for false discovery rate
         adj_pvalue = p.adjust(p_value, method = 'bonferroni'),
         # Pick the smallest per compound & concentration
         min_score = min(adj_pvalue)) %>%
  # filter for adjusted p.val < 0.01 so we keep compounds with at least one signif effect.
  filter(min_score < 0.01) %>%
  distinct(drug, target, conc_char, MMEJ_log2_FC_mean, IPR, min_score) %>%
  ungroup() %>%
  mutate(drug_conc = paste(drug, conc_char, sep = "_"))


# replace -Inf with minimal value of the table.
min.value = min(balance.data.heatmap$MMEJ_log2_FC_mean[is.finite(balance.data.heatmap$MMEJ_log2_FC_mean)])

top_50_hits = balance.data.heatmap %>% 
  distinct(drug_conc, min_score) %>%
  arrange(desc(abs(min_score))) %>% 
  dplyr::slice_tail(n = 60) %>%
  distinct(drug_conc)


#eff.data.heatmap2$freqCut_norm_mean[eff.data.heatmap2$mean_eff_zscore_comb > 0.05] <- NA
balance.data.heatmap2 <- balance.data.heatmap %>% 
  filter(drug_conc %in% top_50_hits$drug_conc) %>%
  dplyr::select(-min_score) %>%
  spread(IPR, MMEJ_log2_FC_mean) %>%
  filter(!is.nan(IPR3))

balance.data.heatmap2 <- balance.data.heatmap2[,c("drug", "target", "conc_char", iprs)]
balance.data.heatmap2$drug <- gsub(" [(].*[)]", "", balance.data.heatmap2$drug) #shorten name of drug


breaks <- quantile(unlist(balance.data.heatmap2 %>% 
                            dplyr::select(contains('IPR'))), c(.05,.95), na.rm = T) # change these numbers according to where you want to set the cutoff
palette_length <- 100

myBreaks <- c(seq(breaks[1], 0, length.out=ceiling(palette_length/2) + 1), 
              seq(-breaks[1]/palette_length, -breaks[1], length.out=floor(palette_length/2)))


# myBreaks <- c(seq(-0.5, 0, length.out=ceiling(palette_length/2) + 1), 
#               seq(0.5/palette_length, 0.5, length.out=floor(palette_length/2)))

ordered_marks2 <- balance.data.heatmap2 %>% 
  dplyr::select(target) %>%
  distinct() %>%
  arrange(target)

ordered_marks2 <- ordered_marks2$target

annotation_colors_scr_set = annotation_colors_scr
annotation_colors_scr_set$target <- annotation_colors_scr$target[ordered_marks2]

# Add cluster annotation
clusters_drugs <- hclust(dist(balance.data.heatmap2), method = "ward.D")
clusters_drugs <- cutree(tree = clusters_drugs, k = 5)
clusters_drugs <- data.frame(clusters_drugs)

balance.data.heatmap2.clusters <- cbind(balance.data.heatmap2, clusters_drugs) %>% 
  as_tibble() %>%
  mutate(drug_conc = paste(drug, conc_char, sep = " "))

clusters_drugs <- balance.data.heatmap2.clusters %>%
  distinct(drug_conc, clusters_drugs)

target_set_all <- target_set_drugs %>%
  rownames_to_column("drug_conc") %>%
  left_join(clusters_drugs) %>%
  na.omit()
rownames(target_set_all) <- target_set_all$drug_conc
target_all <- target_set_all %>%
  dplyr::select(target, 
                clusters_drugs,
                conc_char, viab)
```

```{r S2A TIF heatmaps, warning = FALSE, fig.width = 10, fig.height = 20}
# pdf(paste0(figure_out, "rs20220623_indelfrequency_heatmap_allconc.pdf"), width=20, height=8)
balance.data.matrix = balance.data.heatmap2.clusters %>%
  arrange(target_all) %>%
  column_to_rownames('drug_conc') %>%
  dplyr::select(contains('IPR'))



plt1 = pheatmap(balance.data.matrix[ , clone5_map_ipr$labels],
                cluster_cols = clone5_map_ipr,
                border_color = F,
                cellwidth = 8,
                cellheight = 8,
                breaks = myBreaks,
                clustering_method = "ward.D",
                cluster_rows = T,
                na_col = "white",
                main = 'MMEJ ratio FC (log2), all concentrations',
                annotation_row = target_all,
                annotation_col = chromatin,
                # color = colorRampPalette(c("#e07a5f", "#e5e5e5", "#3d405b"))(palette_length),
                color = diverge_hcl(palette_length, "Blue-Red"),
                annotation_colors = annotation_colors_scr_set, 
                annotation_legend = T,
                cutree_row = 5)
```
*Conclusion Figure 4A: HDAC-rich cluster seems to show H3K27me3-specific MMEJ ratio decrease. No strong MMEJ ratio increases.*


## 5A: Cluster differences per chromatin group
Aim: Is there any chromatin-specificity in the defined MMEJ ratio FC clusters?
```{r}
chromatin_ipr <- chromatin %>%
  rownames_to_column("IPR")

balance.data.heatmap3 <- balance.data.heatmap2 %>%
  pivot_longer(cols = contains("IPR"), names_to = "IPR", values_to = "MMEJratio_FC")

chromatin_effect <- left_join(balance.data.heatmap3, chromatin_ipr) %>%
  #filter(target == "HDAC") %>%
  mutate(MMEJratio_FC = ave(MMEJratio_FC, drug, chromatin, conc_char, FUN = function(x) mean(x, na.rm = TRUE)),
         drug_conc = paste(drug, conc_char, sep = " ")) %>%
  distinct(MMEJratio_FC, drug, chromatin, conc_char, target, drug_conc)

clusters_drugs = target_set_all # %>% filter(clusters_drugs %in% c(3, 4))


chromatin_effect <- left_join(clusters_drugs, chromatin_effect)

ggplot(chromatin_effect,
       aes(x = chromatin, y = MMEJratio_FC, fill = chromatin)) +
  geom_boxplot(outlier.shape = NA) +
  geom_quasirandom() +
  geom_signif(comparisons = combn(sort(unique(chromatin_effect$chromatin)), 2, simplify = F), 
              map_signif_level=sigFunc, 
              step_increase = 0.1) +
  theme_classic_lines_45() +
  scale_fill_manual(values = group_colors) +
  theme(legend.position = "none") +
  facet_wrap(~ clusters_drugs, nrow = 1)
```
*Conclusion Figure 5A: Cluster 3 is most H3K27me3-specific and doesn't change the MMEJ ratio at all in euchromatin.*


## S4A MMEJ ratio change heatmap
Aim: Give an overview of the most important MMEJ ratio fold-changes. 
```{r prepare data for heatmaps}
# Generate heatmap: select top drugs (highest change at any IPR) per concentration
balance.data <- indel.data %>%
  #filter(target != "Aurora Kinase") %>%
  filter(sample == "Epigenetic Compound") %>% 
  filter(pathway_bal_filter == TRUE & viab_reproducibility == TRUE & cor_bal_filter) %>%
  filter(abs(zscore_bal_global_comb_rep) >= 2.58) %>%  # we can choose to display all drugs that have significant global hits
  distinct(conc_char, drug, target, IPR, zscore_bal_comb_rep, MMEJ_log2_FC, replicate, tech)

dmso_bal <- indel.data %>%
  filter(drug == "DMSO") %>%
  dplyr::select('dmso_bal' = MMEJ_log2_FC , IPR) %>%
  mutate(dmso_bal = ave(dmso_bal, IPR, FUN = function(x) mean(x, na.rm = TRUE))) %>%
  distinct()

balance.data.heatmap <- balance.data %>%
  # left_join(dmso_bal) %>%
  group_by(drug, conc_char) %>%
  # ave to add IPR grouping level to calculate mean of the replicates
  mutate(MMEJ_log2_FC_mean = ave(MMEJ_log2_FC, IPR, FUN = function(x) mean(x, na.rm = TRUE)),
         ## Select drugs that have a large effect size in at least one location 
         ## (some drugs might have a small global effect size, but large local)
         p_value = 2*pnorm(q = abs(zscore_bal_comb_rep), lower.tail = F),
         # Adjust p value for false discovery rate
         adj_pvalue = p.adjust(p_value, method = 'bonferroni'),
         # Pick the smallest per compound & concentration
         min_score = min(adj_pvalue)) %>%
  # filter for adjusted p.val < 0.01 so we keep compounds with at least one signif effect.
  filter(min_score < 0.01) %>%
  distinct(drug, target, conc_char, MMEJ_log2_FC_mean, IPR, min_score) %>%
  ungroup()


# replace -Inf with minimal value of the table.
min.value = min(balance.data.heatmap$MMEJ_log2_FC_mean[is.finite(balance.data.heatmap$MMEJ_log2_FC_mean)])

top_50_hits = balance.data.heatmap %>% 
  arrange(desc(abs(MMEJ_log2_FC_mean))) %>% 
  distinct(drug, conc_char, min_score) 


#eff.data.heatmap2$freqCut_norm_mean[eff.data.heatmap2$mean_eff_zscore_comb > 0.05] <- NA
balance.data.heatmap2 <- top_50_hits %>% 
  left_join(balance.data.heatmap)%>%
  dplyr::select(-min_score) %>%
  spread(IPR, MMEJ_log2_FC_mean) %>%
  filter(!is.nan(IPR3))

balance.data.heatmap2 <- balance.data.heatmap2[,c("drug", "target", "conc_char", iprs)]
balance.data.heatmap2$drug <- gsub(" [(].*[)]", "", balance.data.heatmap2$drug) #shorten name of drug


breaks <- quantile(unlist(balance.data.heatmap2 %>% 
                            dplyr::select(contains('IPR'))), c(.05,.95), na.rm = T) # change these numbers according to where you want to set the cutoff
palette_length <- 100

myBreaks <- c(seq(breaks[1], 0, length.out=ceiling(palette_length/2) + 1), 
              seq(-breaks[1]/palette_length, -breaks[1], length.out=floor(palette_length/2)))


# myBreaks <- c(seq(-0.5, 0, length.out=ceiling(palette_length/2) + 1), 
#               seq(0.5/palette_length, 0.5, length.out=floor(palette_length/2)))

ordered_marks2 <- balance.data.heatmap2 %>% 
  dplyr::select(target) %>%
  distinct() %>%
  arrange(target)

ordered_marks2 <- ordered_marks2$target

annotation_colors_scr_set = annotation_colors_scr
annotation_colors_scr_set$target <- annotation_colors_scr$target[ordered_marks2]

# Add cluster annotation
clusters_drugs <- hclust(dist(balance.data.heatmap2), method = "ward.D")
clusters_drugs <- cutree(tree = clusters_drugs, k = 8)
clusters_drugs <- data.frame(clusters_drugs)

balance.data.heatmap2.clusters <- cbind(balance.data.heatmap2, clusters_drugs) %>% 
  as_tibble() %>%
  mutate(drug_conc = paste(drug, conc_char, sep = " "))

clusters_drugs <- balance.data.heatmap2.clusters %>%
  distinct(drug_conc, clusters_drugs)

target_set_all <- target_set_drugs %>%
  rownames_to_column("drug_conc") %>%
  left_join(clusters_drugs) %>%
  na.omit()
rownames(target_set_all) <- target_set_all$drug_conc
target_all <- target_set_all %>%
  dplyr::select(target, 
                clusters_drugs,
                conc_char, viab)
```

```{r S2A TIF heatmaps, warning = FALSE, fig.width = 10, fig.height = 20}
# pdf(paste0(figure_out, "rs20220623_indelfrequency_heatmap_allconc.pdf"), width=20, height=8)
balance.data.matrix = balance.data.heatmap2.clusters %>%
  arrange(target_all) %>%
  column_to_rownames('drug_conc') %>%
  dplyr::select(contains('IPR'))



plt1 = pheatmap(balance.data.matrix[ , clone5_map_ipr$labels],
                cluster_cols = clone5_map_ipr,
                border_color = F,
                cellwidth = 8,
                cellheight = 8,
                breaks = myBreaks,
                clustering_method = "ward.D",
                cluster_rows = T,
                na_col = "white",
                main = 'MMEJ ratio FC (log2), all concentrations',
                annotation_row = target_all,
                annotation_col = chromatin,
                # color = colorRampPalette(c("#e07a5f", "#e5e5e5", "#3d405b"))(palette_length),
                color = diverge_hcl(palette_length, "Blue-Red"),
                annotation_colors = annotation_colors_scr_set, 
                annotation_legend = T,
                cutree_row = 8)
```



## S5A: Cluster differences per chromatin group
Aim: Is there any chromatin-specificity in the defined MMEJ ratio FC clusters?
```{r}
chromatin_ipr <- chromatin %>%
  rownames_to_column("IPR")

balance.data.heatmap3 <- balance.data.heatmap2 %>%
  pivot_longer(cols = contains("IPR"), names_to = "IPR", values_to = "MMEJratio_FC")

chromatin_effect <- left_join(balance.data.heatmap3, chromatin_ipr) %>%
  filter(target %in% c("HDAC", "Aurora Kinase", "Epigenetic Reader Domain")) %>%
  filter(drug != "Barasertib") %>%
  mutate(MMEJratio_FC = ave(MMEJratio_FC, drug, chromatin, conc_char, FUN = function(x) mean(x, na.rm = TRUE)),
         drug_conc = paste(drug, conc_char, sep = " ")) %>%
  distinct(MMEJratio_FC, drug, chromatin, conc_char, target, drug_conc)

clusters_drugs = target_set_all # %>% filter(clusters_drugs %in% c(3, 4))


chromatin_effect <- left_join(clusters_drugs, chromatin_effect)

ggplot(chromatin_effect,
       aes(x = chromatin, y = MMEJratio_FC, fill = chromatin)) +
  geom_boxplot(outlier.shape = NA) +
  geom_quasirandom() +
  geom_signif(comparisons = combn(sort(unique(chromatin_effect$chromatin)), 2, simplify = F), 
              map_signif_level=sigFunc, 
              step_increase = 0.1) +
  theme_classic_lines_45() +
  scale_fill_manual(values = group_colors) +
  theme(legend.position = "none") +
  facet_wrap(~ clusters_drugs, nrow = 1)
```
*Conclusion Figure S5A: Again, the same cluster is most H3K27me3-specific and doesn't change the MMEJ ratio at all in euchromatin.*



```{r}
#data frame to save data
mean_sd_fits <- tibble(term = NA, estimate = NA, std.error = NA, barcode = NA, rep = NA)


# First calculate means and sd of the DMSO efficiency per replicate and barcode
dmso.eff.score <- indel.data %>%
  filter(drug == "DMSO") %>%
  dplyr::select('DMSO' = MMEJratio, well, barcode, replicate, tech, plate) %>%
  mutate(DMSO = ave(DMSO, well, replicate, tech, plate, FUN = function(x) mean(x, na.rm=T))) %>%
  distinct(replicate, tech, well, plate, DMSO) %>% 
  pull(DMSO) %>%
  fitdistr(., "normal") %>% tidy()
```

```{r Compounds that affect the balance at all the sites}
## Which compounds change the balance at all the sites in both screens?
all_balance = indel.data %>% 
  distinct(drug, conc_char, replicate, zscore_bal_comb_rep, target, 
           barcode, viab_reproducibility, pathway_bal_filter) %>% 
  filter(!is.na(zscore_bal_comb_rep) & 
           !between(zscore_bal_comb_rep, -1.96, 1.96) & 
           viab_reproducibility == T & 
           pathway_bal_filter == T) %>% 
  group_by(drug, conc_char, target) %>% 
  count() %>% 
  filter(n == 36) # 2 x 18

kable(all_balance)
all_balance %>% ungroup() %>% distinct(drug, target)

## Now with 16/18 (~ 90%) of the sites with an effect. Near global
near_all_balance = indel.data %>% 
  distinct(drug, conc_char, replicate, zscore_bal_comb_rep, target, 
           barcode, viab_reproducibility, pathway_bal_filter) %>% 
  filter(!is.na(zscore_bal_comb_rep) & 
           !between(zscore_bal_comb_rep, -1.96, 1.96) & 
           viab_reproducibility == T & 
           pathway_bal_filter == T) %>% 
  group_by(drug, conc_char, target) %>% 
  count() %>% 
  filter(n >= 32) # 2 x 16


kable(near_all_balance)
near_all_balance %>% ungroup() %>% distinct(drug, target)
```

```{r warning = FALSE}
# Plot raw MMEJ ratio
dmso_bal <- indel.data %>%
  filter(drug == "DMSO") %>%
  dplyr::select('dmso_bal' = MMEJratio, IPR, replicate, tech, well, plate) %>%
  group_by(replicate, tech, well, plate) %>%
  summarise(dmso_bal = mean(dmso_bal)) %>%
  ungroup() %>%
  mutate(dmso_mean = mean(dmso_bal),
         dsmo_sd = sd(dmso_bal)) %>%
  distinct(dmso_mean, dsmo_sd)


indel.data %>%
  filter(sample == "Epigenetic Compound" & pathway_bal_filter == TRUE & viab_reproducibility == TRUE) %>%
  distinct(MMEJratio, drug, conc_char, zscore_bal_global_comb_rep) %>%
  mutate(GlobalMMEJratio = ave(MMEJratio, drug, conc_char, FUN = mean)) %>%
  distinct() %>%
  mutate(zscore_bal_global_comb_rep = 2*pnorm(q = abs(zscore_bal_global_comb_rep), lower.tail = F)) %>%
  mutate(MMEJ_pval_adj_global_comb = log10(p.adjust(zscore_bal_global_comb_rep, method = 'fdr')),
         col_bal = ifelse(MMEJ_pval_adj_global_comb < 0.01 & 
                            GlobalMMEJratio > (dmso_bal$dmso_mean + dmso_bal$dsmo_sd), "green", 
                          ifelse(MMEJ_pval_adj_global_comb < 0.01 & 
                                   GlobalMMEJratio < (dmso_bal$dmso_mean - dmso_bal$dsmo_sd), "yellow","blue"))) %>%  
  filter(col_bal != "blue") %>%
  distinct(drug, conc_char)
```

---

## 4B: MMEJ ratio change per target group
Aim: Are there any target groups that are prone to changes in MMEJ ratio?
```{r warning = FALSE}
# C: Bar plot
indel.data.bal <- indel.data %>%
  filter((sample == "Epigenetic Compound" | drug == "DMSO") & 
           pathway_bal_filter & viab_reproducibility &
           cor_bal_filter & !is.na(MMEJratio)) %>%
  distinct(IPR, drug, conc_char, sample, target, viab_mean, replicate,
           MMEJratio_global, MMEJratio, IPR,
           zscore_bal_global_comb_rep , filter_zscore_bal_global_comb_rep, 
           zscore_bal_comb_rep , filter_zscore_bal_comb_rep) %>%
  filter(!(viab_mean < 0.5)) %>%
  mutate(MMEJratio_mean = ave(MMEJratio_global, drug, conc_char))


bal.data <- indel.data.bal %>%
  filter(sample == "Epigenetic Compound") %>%
  distinct(IPR, zscore_bal_global_comb_rep , filter_zscore_bal_global_comb_rep, 
           zscore_bal_comb_rep , filter_zscore_bal_comb_rep, 
           drug, target, conc_char) %>%
  mutate(mean_eff_pval_comb = 2*pnorm(q = abs(zscore_bal_comb_rep ), lower.tail = F),
         mean_eff_pval_adj_comb = p.adjust(mean_eff_pval_comb, method = 'fdr'),
         sign_local = ifelse((mean_eff_pval_adj_comb < 0.01 & 
                                filter_zscore_bal_comb_rep &
                                 zscore_bal_comb_rep > 2.58), "MMEJbal_increase", 
                             ifelse((mean_eff_pval_adj_comb < 0.01 &
                                       filter_zscore_bal_global_comb_rep &
                                 zscore_bal_comb_rep < -2.58), "MMEJbal_decrease", 
                                    "No_Change"))) %>%
  group_by(drug, conc_char) %>%
  mutate(no_change_n = length(table(sign_local))) %>%
  mutate(sign_global = ifelse(no_change_n == 1,  names(table(sign_local)), "No_Change")) %>%
  distinct(IPR, drug, conc_char, target, sign_local, sign_global)

global.effects = bal.data %>%
  distinct(drug, conc_char, target, sign_global) %>%
  group_by(target, sign_global, conc_char) %>% 
  count()

local.effects = bal.data %>% 
  filter(sign_global == "No_Change") %>% 
  distinct(drug, conc_char, target, sign_local) %>%
  group_by(target, sign_local, conc_char) %>% 
  count() %>% 
  pivot_wider(names_from = sign_local, values_from = n, values_fill = 0) %>%
  mutate(No_Change = No_Change - MMEJbal_increase - MMEJbal_decrease) %>%
  dplyr::rename(Local_MMEJbal_increase = MMEJbal_increase, Local_MMEJbal_decrease = MMEJbal_decrease) %>% 
  pivot_longer(cols = No_Change:Local_MMEJbal_decrease, values_to = "n", names_to = "sign_global")

target_effects = global.effects %>% 
  filter(sign_global != "No_Change") %>%
  rbind(local.effects) %>%
  mutate(sign_global = factor(sign_global, levels = c(c("MMEJbal_decrease", "Local_MMEJbal_decrease", "No_Change","Local_MMEJbal_increase", "MMEJbal_increase"))))

ggplot(target_effects,
       aes(y = reorder(target, n), x = n, fill = sign_global)) +
  geom_bar(stat = "identity") +
  facet_wrap(~conc_char)+
  scale_fill_manual(values = c("MMEJbal_decrease" = "#023FA5", "Local_MMEJbal_decrease" = "#A1A6C8", "No_Change" = "#E2E2E2","Local_MMEJbal_increase" = "#CA9CA4", "MMEJbal_increase" = "#8E063B")) +
  theme_pubr(border = T)

target_effects %>% filter(sign_global != "No_Change") %>% distinct(target, conc_char, n) %>% group_by(conc_char) %>% summarise(count = sum(n))
```

*Conclusion Figure 4B: Again, HDACs have a significant fraction of local changes. Also AURK and ERD inhibitors stick out.*


---


## 4C: Highlighting the strongest MMEJ ratio reducers. 
Aim: Show which drugs and which target groups are responsible for the biggest global MMEJ ratio decrease.
```{r 2C Top TIF boosting, , fig.width = 12, fig.height = 10}
rank_dt = indel.data %>% 
  filter(drug != "DMSO" & pathway_bal_filter & viab_reproducibility & cor_bal_filter, filter_zscore_bal_global_comb_rep) %>%
  distinct(drug, conc_char, MMEJratio_global_mean , target, cor_bal, cor_bal_filter) %>% 
  arrange(MMEJratio_global_mean) %>%
  mutate(rank = seq(1:n()),
         drug = gsub(" [(].*[)]", "", drug),
         drug_conc = paste(drug, conc_char, sep = " "))

rank_dt %>%
  filter(rank <= 150) %>%
  ggplot(., aes(rank, MMEJratio_global_mean , color = target, size = 1)) +
  geom_point()  +
  scale_color_manual(values = colors_targets) +
  gghighlight(rank <= 34, use_direct_label = F) +
  geom_hline(linetype = "dashed", yintercept = 0.3711068) +
  ylab("Global MMEJ ratio")



top_hits = indel.data  %>% 
  filter(conc_char != "control" & pathway_bal_filter == T & viab_reproducibility == TRUE) %>% 
  distinct(drug, conc_char, zscore_bal_global_comb_rep , target) %>%
  arrange(desc(zscore_bal_global_comb_rep )) %>% slice_head(n = 10) %>% distinct(drug, conc_char, target, zscore_bal_global_comb_rep ) %>% 
  mutate(rank = seq(1:n()),
         drug = gsub(" [(].*[)]", "", drug),
         drug_conc = paste(drug, conc_char, sep = " "))
```
*Conclusion Figure 4C: Many AURKi are the most stron MMEJ ratio reducers. Also Decitabine and a HDAC inhibitor are amongst the strongest.*

## 5C: Indel plot Vorinostat
Aim: Zoom in on Vorinostat to showcase the chromatin-specific MMEJ ratio reduction. 
```{r}
# Import data from preprocessing script
indel.data <- readRDS("data/episcreen_ratios_mt20220811.RDS")
mutation.data <- readRDS("data/episcreen_mutations_mt20220811.RDS")

chrom <- chromatin %>%
  rownames_to_column("IPR") 

label_ratio <- indel.data %>%
  mutate(drug_conc = paste(drug, conc_char, sep = " ")) %>%
  filter(drug_conc %in% c("DMSO control", "Vorinostat (SAHA, MK0683) 1 µM")) %>% 
  na.omit() %>%
  group_by(drug_conc, IPR) %>%
  dplyr::summarise(MMEJratio_mean = mean(MMEJratio),
                   MMEJratio_sd = sd(MMEJratio),
                   freqCut_mean = mean(freqCut),
                   freqCut_sd = sd(freqCut)) %>%
  ungroup() %>%
  left_join(chrom, by = "IPR")

FC_df <- indel.data %>% 
  filter(drug == "Vorinostat (SAHA, MK0683)" & conc_char == "1 µM") %>% 
  distinct(drug, MMEJ_log2_FC, freqCut_diff_mean, IPR) %>% 
  mutate(MMEJ_log2_FC = ave(MMEJ_log2_FC, IPR, FUN = function(x) mean(x, na.rm = T))) %>% 
  mutate(freqCut_diff_mean = ave(freqCut_diff_mean, IPR, FUN = function(x) mean(x, na.rm = T))) %>% 
  distinct() %>%   
  left_join(chrom, by = "IPR") %>% 
  mutate(chrom_mmej_FC = ave(MMEJ_log2_FC, chromatin, FUN = mean)) %>% 
  mutate(chrom_tif_FC = ave(freqCut_diff_mean, chromatin, FUN = mean)) %>% 
  mutate(chrom_mmej_FC_sd = ave(MMEJ_log2_FC, chromatin, FUN = sd)) %>% 
  mutate(chrom_tif_FC_sd = ave(freqCut_diff_mean, chromatin, FUN = sd)) 


label_ratio <- label_ratio %>%
  mutate(MMEJratio_mean_chrom = ave(MMEJratio_mean, drug_conc, chromatin, FUN = mean)) %>%
  mutate(MMEJratio_sd_chrom = ave(MMEJratio_mean, drug_conc, chromatin, FUN = sd)) %>%
  mutate(freqCut_mean_chrom = ave(freqCut_mean, drug_conc, chromatin, FUN = mean)) %>%
  mutate(freqCut_sd_chrom = ave(freqCut_mean, drug_conc, chromatin, FUN = sd))  %>%
  mutate(label_efficiency = paste0(drug_conc, " TIF = ", round(freqCut_mean_chrom*100, 1),"% ±", round(freqCut_sd_chrom*100, 1),"%"), 
         label_ratio = paste0(drug_conc, "MMEJ ratio = ", round(MMEJratio_mean_chrom, 2)," ±", round(MMEJratio_sd_chrom, 2))) 

label_ratio$drug_conc[label_ratio$drug_conc == "Vorinostat (SAHA, MK0683) 1 µM"] <- "Vorinostat 1 µM"

bc_mut_tib_mean <- mutation.data %>% 
  mutate(drug = gsub(" [(].*[)]", "", drug),
         drug_conc = paste(drug, conc_char, sep = " ")) %>%
  distinct(IPR, drug_conc, freq_mean,freq_sd, color, mutation, type) %>%
  filter(drug_conc %in% c("DMSO control", "Vorinostat 1 µM"), type %in% c("core", "grouped_indel")) %>% 
  na.omit() %>%
  left_join(label_ratio, by = c("drug_conc", "IPR"))

ggplot(bc_mut_tib_mean, 
            aes(mutation, freq_mean, color = color, fill = drug_conc)) + 
  geom_boxplot(position = position_dodge(0.75), aes(alpha = 0.7)) +
  geom_quasirandom(dodge.width = 0.75, shape = 21) + 
  scale_fill_manual("legend", values = c("white", "#60988D")) + 
  scale_color_manual("legend", values = colors_pathways) + 
  #scale_y_continuous(name="frequency", labels = percent) +
  scale_x_discrete(name="indel size", breaks = c("<-14", -7, 0, 1, ">2")) + 
  facet_wrap( ~ chromatin, ncol = 2) +
  geom_text(aes(x = 7, y = .7, label = label_efficiency)) +
  geom_text(aes(x = 7, y = .6, label = label_ratio)) + 
  theme_pubr(border = T)
```
*Conclusion Figure 5C: Vorinostat reduces MMEJ more in H3K27me3 regions*


## 4D: Indel plot decitabine
Aim: Zoom in on Decitabine to showcase the strong MMEJ reduction. 
```{r}
# Import data from preprocessing script
in.dir = list.dirs(path = "/DATA/projects/DSBrepair/data/R") %>% .[grepl("episcreen", .)] %>% tail(n = 1)
file = list.files(in.dir, pattern = "episcreen_ratios", full.names = T) %>% tail(n = 1)
indel.data <- readRDS(file)

# Import data from preprocessing script
in.dir = list.dirs(path = "/DATA/projects/DSBrepair/data/R") %>% .[grepl("episcreen", .)] %>% tail(n = 1)
file.muts = list.files(in.dir, pattern = "episcreen_mutations", full.names = T) %>% tail(n = 1)

mutation.data <- readRDS(file.muts)

chrom <- chromatin %>%
  rownames_to_column("IPR")

label_ratio <- indel.data %>%
  mutate(drug_conc = paste(drug, conc_char, sep = " ")) %>%
  filter(drug_conc %in% c("DMSO control", "Decitabine 1 µM")) %>% 
  na.omit() %>%
  group_by(drug_conc, IPR) %>%
  dplyr::summarise(MMEJratio_mean = mean(MMEJratio),
                   MMEJratio_sd = sd(MMEJratio),
                   freqCut_mean = mean(freqCut),
                   freqCut_sd = sd(freqCut)) %>%
  ungroup() %>%
  left_join(chrom, by = "IPR") %>%
  mutate(MMEJratio_mean_chrom = ave(MMEJratio_mean, drug_conc, chromatin, FUN = mean)) %>%
  mutate(MMEJratio_sd_chrom = ave(MMEJratio_mean, drug_conc, chromatin, FUN = sd)) %>%
  mutate(freqCut_mean_chrom = ave(freqCut_mean, drug_conc, chromatin, FUN = mean)) %>%
  mutate(freqCut_sd_chrom = ave(freqCut_mean, drug_conc, chromatin, FUN = sd))  %>%
  mutate(label_efficiency = paste0(drug_conc, " TIF = ", round(freqCut_mean_chrom*100, 1),"% ±", round(freqCut_sd_chrom*100, 1),"%"), 
         label_ratio = paste0(drug_conc, "MMEJ ratio = ", round(MMEJratio_mean_chrom, 2)," ±", round(MMEJratio_sd_chrom, 2))) 

bc_mut_tib_mean <- mutation.data %>% 
  mutate(drug = gsub(" [(].*[)]", "", drug),
         drug_conc = paste(drug, conc_char, sep = " ")) %>%
  distinct(IPR, drug_conc, freq_mean,freq_sd, color, mutation, type) %>%
  filter(drug_conc %in% c("DMSO control", "Decitabine 1 µM"), type %in% c("core", "grouped_indel")) %>% 
  na.omit() %>%
  left_join(label_ratio, by = c("drug_conc", "IPR"))

ggplot(bc_mut_tib_mean, 
            aes(mutation, freq_mean, color = color, fill = drug_conc)) + 
  geom_boxplot(position = position_dodge(0.75), aes(alpha = 0.7)) +
  geom_quasirandom(dodge.width = 0.75, shape = 21) + 
  scale_fill_manual("legend", values = c("#D69C81", "white")) + 
  scale_color_manual("legend", values = colors_pathways) + 
  #scale_y_continuous(name="frequency", labels = percent) +
  scale_x_discrete(name="indel size", breaks = c("<-14", -7, 0, 1, ">2")) + 
  #facet_wrap( ~ chromatin, ncol = 2) +
  geom_text(aes(x = 7, y = .7, label = label_efficiency)) +
  geom_text(aes(x = 7, y = .6, label = label_ratio)) + 
  theme_pubr(border = T)
```
*Conclusion Figure 4D: Decitabine reduces MMEJ drastically*





```{r prepare heatmaps, fig.width= 10, fig.height= 20}
clone5.matrix = indel.data %>% 
  dplyr::select(IPR, ends_with(".zscore")) %>% 
  distinct() %>% 
  rename_with(., ~gsub(".zscore", "", .x)) %>%
  column_to_rownames(var="IPR")

clone5_map = hclust(dist(t(clone5.matrix)))
clone5_map_ipr = dendextend::rotate(hclust(dist(clone5.matrix)), c(1:2,5:8,3:4, 15:14, 17, 16, 18 ,13,10:12,9))

chromatin <- readRDS(
  "/DATA/usr/x.vergara/XV_ChIPIndels/XV20200902_DDR_RS_revision/XV20200902_DDR_RS_revision/data/xv20200915_DDR_data_CHIP.rds") %>% 
  distinct(barcode, chromatin) %>%
  left_join(distinct(indel.data, barcode, IPR)) 

chromatin <- chromatin %>%
  filter(IPR != "<NA>") %>% 
  dplyr::select(-barcode) %>%
  column_to_rownames(var="IPR")

# Set annotation for heatmap
viab_drug = indel.data %>% filter(drug != "PAO") %>% 
  mutate(drug = gsub(" [(].*[)]", "", drug),
         drug_conc = paste(drug, conc_char, sep = " ")) %>% 
  distinct(viab_mean, drug_conc, replicate) %>% group_by(drug_conc) %>%
  summarise(viab = mean(viab_mean))

# Row annotation: annotate drugs with the target group
target <- indel.data %>% 
  # filter(!target %in% c("Negative Control", "MRN", "DNA-PK")) %>%
  mutate(drug = gsub(" [(].*[)]", "", drug),
         drug_conc = paste(drug, conc_char, sep = " ")) %>%
  left_join(viab_drug) %>%
  distinct(drug_conc, conc_char, target, viab) %>% 
  column_to_rownames(var="drug_conc")

TIF_IPR = indel.data %>% filter(drug == "DMSO") %>%
  distinct(replicate, tech, freqCut,MMEJratio, IPR, well, plate) %>%
  group_by(IPR) %>%
  summarise(TIF = mean(freqCut),
            MMEJratio = mean(MMEJratio))

chromatin_TIF = chromatin %>% rownames_to_column(var = "IPR") %>%
  left_join(TIF_IPR) %>%
  column_to_rownames("IPR")
```

## 5D: Vorinostat pool follow-up
Aim: Show that Vorinostat also specifically acts on H3K27me3 domains in the pool. 
```{r}
vorinostat_validation = indel.data.validation %>% 
  filter(replicate == "mean" & plasmid == "LBR2", drug %in% c("DMSO", "Vorinostat"), sum_bc_reads > 1000)%>% 
  dplyr::select(barcode, drug, MMEJ_NHEJ, efficiency, all_of(chromatin_cols)) %>%
  distinct() 

barcde_all_reps_vorino = vorinostat_validation %>% 
  distinct(barcode, drug) %>% 
  group_by(barcode) %>%
  summarise(bccount = n()) %>% 
  filter(bccount == 2) %>%
  pull(barcode)

vorinostat_ratios = vorinostat_validation  %>% 
  filter(barcode %in% barcde_all_reps_vorino) %>%
  pivot_wider(names_from = drug, values_from = c(efficiency, MMEJ_NHEJ)) %>%
  mutate(diff_TIF = efficiency_Vorinostat - efficiency_DMSO,
         ratio_TIF = efficiency_Vorinostat/efficiency_DMSO,
         log2_FC_TIF = log2(ratio_TIF),
         diff_MMEJ_NHEJ = MMEJ_NHEJ_Vorinostat - MMEJ_NHEJ_DMSO,
         ratio_MMEJ_NHEJ = MMEJ_NHEJ_Vorinostat/MMEJ_NHEJ_DMSO,
         log2_FC_MMEJ_NHEJ = log2(ratio_MMEJ_NHEJ)) 

barcde_all_reps_gsk = RSTP2_IndelRatios_Chromatin_2kb %>% 
  distinct(barcode, drug) %>% 
  group_by(barcode) %>%
  summarise(bccount = n()) %>% 
  filter(bccount == 2) %>%
  pull(barcode)

GSK_ratios = RSTP2_IndelRatios_Chromatin_2kb %>% 
  filter(barcode %in% barcde_all_reps_gsk) %>%
  # distinct(barcode, drug, MMEJ_NHEJ , efficiency) %>% 
  pivot_wider(names_from = drug, values_from = c(efficiency, MMEJ_NHEJ)) %>%
  mutate(diff_TIF = efficiency_GSK126 - efficiency_DMSO_GSK,
         ratio_TIF = efficiency_GSK126/efficiency_DMSO_GSK,
         log2_FC_TIF = log2(ratio_TIF),
         diff_MMEJ_NHEJ = MMEJ_NHEJ_GSK126 - MMEJ_NHEJ_DMSO_GSK,
         ratio_MMEJ_NHEJ = MMEJ_NHEJ_GSK126/MMEJ_NHEJ_DMSO_GSK,
         log2_FC_MMEJ_NHEJ = log2(ratio_MMEJ_NHEJ))


domain_levels = c('late_replicating', 'LAD', 'H3K9me2', 'H3K27me3')

domains = paste0(domain_levels, '_domain')

vorino.dt = vorinostat_ratios %>%
  filter(!is.na(diff_TIF)) %>%
  dplyr::select(barcode, log2_FC_MMEJ_NHEJ, domains) %>%
  distinct(barcode, .keep_all = TRUE) %>%
  data.table()

dim(vorino.dt)




vorino.dt[vorino.dt=='iDomain'] <- 0
vorino.dt[vorino.dt=='Domain'] <- 1
vorino.dt[vorino.dt=='DHS'] <- 1

vorino.dt[is.na(LAD_domain), LAD_domain:=0]
# vorino.dt[is.na(DHS_domain), DHS_domain:=0]

vorino.dt[, (domains) := lapply(.SD, as.numeric), .SDcols = domains]

colnames(vorino.dt) = gsub('_domain', '', colnames(vorino.dt))

opt_df = get_combinations(domain_levels)

k27me3 = vorino.dt %>% filter(late_replicating == 0, 
                             LAD == 0,
                             H3K9me2 == 0,
                             H3K27me3 == 1) %>% 
  pull(log2_FC_MMEJ_NHEJ)

any.k27me3 = vorino.dt %>% filter(H3K27me3 == 1) %>% 
  pull(log2_FC_MMEJ_NHEJ)

not.k27me3 = vorino.dt %>% filter(H3K27me3 != 1) %>% 
  pull(log2_FC_MMEJ_NHEJ)

hetcomb = vorino.dt %>% filter(late_replicating == 1, 
                              LAD == 1,
                              H3K9me2 == 1,
                              H3K27me3 == 0) %>% 
  pull(log2_FC_MMEJ_NHEJ)

idomain = vorino.dt %>% filter(late_replicating == 0, 
                              LAD == 0,
                              H3K9me2 == 0,
                              H3K27me3 == 0) %>% 
  pull(log2_FC_MMEJ_NHEJ)

# DHS = vorino.dt %>% filter(late_replicating == 0,
#                           LAD == 0,
#                           H3K9me2 == 0,
#                           H3K27me3 == 0,
#                           DHS == 1) %>%
#   pull(diff_TIF)

wilcox.test(idomain, k27me3, alternative = "two.sided")

wilcox.test(idomain, any.k27me3, alternative = "two.sided")
wilcox.test(not.k27me3, any.k27me3, alternative = "two.sided")

wilcox.test(idomain, hetcomb, alternative = "two.sided")

vorino.dt %>% mutate(state = ifelse(late_replicating == 1 & 
                                      LAD == 1 &
                                      H3K9me2 == 1 &
                                      H3K27me3 == 0, "Triple Heterochromatin", ifelse(
                                        late_replicating == 0 & 
                                          LAD == 0 & 
                                          H3K9me2 == 0 & 
                                          H3K27me3 == 0, "Euchromatin", ifelse(late_replicating == 0 &  
                                                                                 LAD == 0 & 
                                                                                 H3K9me2 == 0 & 
                                                                                 H3K27me3 == 1, "H3K27me3", "other-heterochromatin")))) %>%
  mutate(state = factor(state, levels = c("Euchromatin", "other-heterochromatin", "H3K27me3", "Triple Heterochromatin"))) %>%
  ggplot(., aes(log2_FC_MMEJ_NHEJ, fill = state)) +
  geom_density(alpha = 0.6, color = "white", size = .25) +
  scale_fill_manual(values = group_colors)
```
*Conclusion Figure 5D: MMEJ ratio FC is the strongest in H3K27me3 domains*

# Supplementary Figure 5

## S5B: Vorinostat MMEJ FC per chromatin group
Aim: Present a detailed view of the Vorinostat pool data
```{r}
plot_comb_grid_beeswarm(vorino.dt, opt_df,
                        domain_levels, 'log2_FC_MMEJ_NHEJ', min_count = 10, y_plabel=c(-2.5, 2.5))
```
*Conclusion Figure S5B: Triple heterochromatin is unaltered from treatment, H3K27me3 domains are affected*

## S5C Vorinostat MMEJ FC correlation with chromatin features
Aim: Show that H3K27me3 is the strongest correlating factor. 
```{r S3C Vorino TIF FC Correlations}
## Vorinostat
chrom.mods.cols <- seq(grep("CTCF", names(vorinostat_ratios)), grep("H3K27me3_domain", names(vorinostat_ratios))-1)

vorinostat.hist.corr <- CorrelationDFMean(vorinostat_ratios, "log2_FC_MMEJ_NHEJ", targets = chrom.mods.cols) %>%
  arrange(corr) %>% pull(rowname) %>% factor(.,levels = .)

lims = 1

CorrelationPlotMean(vorinostat_ratios, "log2_FC_MMEJ_NHEJ", targets = vorinostat.hist.corr, title = "with log2 FC TIF after Vorinostat treatment")
```
*Conclusion Figure S5B: H3K27me3 is the feature that correlates the most with reduction in MMEJ ratio upon Vorinostat treatment*


## S4C: Cell cycle analysis
Aim: Show which drugs are affected by G2 arrest. 
```{r, fig.width = 8, fig.height = 4, eval = FALSE}
# rmarkdown::render("3_Processing_FACS.Rmd")
all.samples = read_csv2("/DATA/usr/m.trauernicht/projects/EpiScreen/FACS/Data/mt20220802_facs_all_drugs.csv")

all.samples <- all.samples %>%
  mutate(drug_conc = paste(drug, conc, sep = " ")) %>%
  mutate(drug_conc_exp = paste(drug, conc, exp, sep = "_")) %>%
  mutate(mean_value = ave(value, drug_conc, variable, exp, FUN = mean)) %>%
  mutate(SD = ave(value, drug_conc, variable, exp, FUN = sd)) %>%
  filter(exp == "effective_drugs") %>%
  distinct(drug, conc, drug_conc_exp, variable, mean_value, SD, value, exp, drug_conc) %>%
  mutate(pval = "")

for (i in unique(all.samples$drug_conc_exp)) {
  if (length(all.samples$value[all.samples$drug_conc_exp == i & all.samples$variable == "G2" & !is.na(all.samples$value)]) > 1) { 
    if (all.samples$exp[all.samples$drug_conc_exp == i] == "five_drugs") {

      all.samples$pval[all.samples$drug_conc_exp == i] <- t.test(all.samples$value[all.samples$drug_conc == "DMSO 10um" & all.samples$exp == "five_drugs" & all.samples$variable == "G2"], 
                                                                 all.samples$value[all.samples$drug_conc_exp == i & all.samples$variable == "G2"])$p.value
      } else {
        all.samples$pval[all.samples$drug_conc_exp == i] <- t.test(all.samples$value[all.samples$drug_conc == "DMSO 10um" & all.samples$exp == "effective_drugs" & all.samples$variable == "G2"], 
                                                               all.samples$value[all.samples$drug_conc_exp == i & all.samples$variable == "G2"])$p.value
  }
  }
}

all.samples <- all.samples %>%
  dplyr::select(-value) %>%
  distinct()

vadj.G2 = all.samples %>% filter(variable %in% c("G2", "S", "G1")) %>% group_by(drug_conc_exp) %>% summarise(vadj = sum(mean_value)) %>% mutate(variable = "G2")
vadj.S = all.samples %>% filter(variable %in% c("S", "G1")) %>% group_by(drug_conc_exp) %>% summarise(vadj = sum(mean_value)) %>% mutate(variable = "S")
vadj.G1 = all.samples %>% filter(variable %in% c("G1")) %>% group_by(drug_conc_exp) %>% summarise(vadj = sum(mean_value)) %>% mutate(variable = "G1")

vadj.tib = rbind(vadj.G1, vadj.S, vadj.G2) %>% na.omit()
targets <- target_set_drugs %>% 
  distinct(drug, target)

all.samples.plot = all.samples  %>%
  as_tibble() %>%
  filter(variable != "Apoptotic cells") %>%
  left_join(vadj.tib) %>%
  left_join(targets, by = "drug") %>%
  mutate(variable = factor(variable, levels = c("G2", "S", "G1")),
         target = ifelse(drug == "CX-6258", "PIM",
                         ifelse(drug == "DMSO", "Vehicle",
                                ifelse(drug == "DZNeP", "HMT",
                                       ifelse(drug == "Suramin", "Sirtuin",target))))) %>%
  na.omit()


#aurk.inhibitors <- c("Alisertib 100 nM", "Aurora A Inhibitor I 1 µM", "Barasertib 1 µM", "CYC116 1 µM", "SNS-314 Mesylate 10 µM")

g2_order <- all.samples.plot %>%
  filter(variable == "G2") %>%
  dplyr::select(drug_conc_exp, "g2_value" = mean_value) %>%
  distinct()

all.samples.plot <- merge(all.samples.plot, g2_order, all = T)

drugs_order <- all.samples.plot %>%
  distinct(drug_conc, g2_value) %>%
  filter(str_detect(drug_conc, "DMSO", negate = T)) %>%
  arrange(desc(g2_value))

all.samples.plot <- all.samples.plot %>% 
  filter(drug != "PAO") %>%
  mutate(drug_conc = factor(drug_conc, levels = c("DMSO 10um", drugs_order$drug_conc)))

ggplot(all.samples.plot, aes(y = mean_value, x = drug_conc)) +
  geom_linerange(aes(ymin = vadj, ymax = vadj+SD, color = target), size = 1) +
  geom_bar(stat="identity", aes(alpha = variable, fill = target), color = ifelse(all.samples.plot$pval < 0.05, "grey40", NA), size = .5, width = .75) +
  labs(title = "cell cycle arrest upon treatment with Aurora Kinase inhibitors") +
  scale_color_manual(values = annotation_colors_scr$target) +
  scale_fill_manual(values = annotation_colors_scr$target) +
  scale_alpha_discrete(range = c(0.3,1)) +
  theme_pubr(border = T, x.text.angle = 90)
```
*Conclusion Figure S4C: Mostly AURKi affect cell cycle, some other drugs (Azacitdine, NSC-3852) also do to a limited extent which might explain their MMEJ ratio reduction capacity.*


## HDACi only heatmap
```{r}
HDAC_tib = read.table("rs20220708_HDAC_targets.txt", sep = "\t", fill = NA, header = T) %>%
  dplyr::select(-target_main, -target_second) %>%
  as_tibble()

# cluster1 = target_all %>% filter(target == "HDAC") %>% rownames()
# cluster1 = cluster1[!grepl("Nullscript", cluster1)]
# 
# top_50_hits_hdacs <- top_50_hits %>%
#   mutate(drug_conc = paste(drug, conc_char, sep = " "))

eff.data.heatmap.cluster1 = balance.data.heatmap2 %>% 
  mutate(drug_conc = paste(drug, conc_char, sep = " ")) %>% 
  filter(target == "Epigenetic Reader Domain")

  # filter(drug_conc %in% cluster1) %>%
  # filter(drug_conc %in% top_50_hits_hdacs$drug_conc) 

# Add cluster annotation
cluster_HDAC <- hclust(dist(eff.data.heatmap.cluster1), method = "ward.D")
cluster_HDAC <- cutree(tree = cluster_HDAC, k = 3)
cluster_HDAC <- data.frame(cluster_HDAC)

eff.data.heatmap.cluster1 <- cbind(eff.data.heatmap.cluster1, cluster_HDAC) %>% 
  as_tibble()

cluster_HDAC <- eff.data.heatmap.cluster1 %>%
  distinct(drug_conc, cluster_HDAC)

target.cluster1 <- target_set_drugs %>%
  rownames_to_column("drug_conc") %>%
  left_join(HDAC_tib) %>%
  left_join(cluster_HDAC) %>%
  filter(!is.na(cluster_HDAC)) %>%
  replace(is.na(.), 0)

rownames(target.cluster1) <- target.cluster1$drug_conc
target.cluster1 <- target.cluster1 %>%
  dplyr::select(cluster_HDAC, conc_char, pan_HDAC, HDAC1,HDAC2, HDAC3,
                HDAC4, HDAC6, HDAC8, HDAC10, viab)

target.cluster1

# target_set_all <- target_set_drugs %>%
#   rownames_to_column("drug_conc") %>%
#   left_join(clusters_drugs) %>%
#   na.omit()
# rownames(target_set_all) <- target_set_all$drug_conc
# target_all <- target_set_all %>%
#   dplyr::select(target, 
#                 clusters_drugs,
#                 conc_char, viab)



HDAC_matrix = eff.data.heatmap.cluster1 %>%
  arrange(target.cluster1) %>%
  column_to_rownames('drug_conc') %>%
  dplyr::select(contains('IPR')) %>%
  t()

breaks <- quantile(unlist(balance.data.heatmap2 %>% 
                            dplyr::select(contains('IPR'))), c(.05,.95), na.rm = T) # change these numbers according to where you want to set the cutoff
palette_length <- 100

myBreaks <- c(seq(breaks[1], 0, length.out=ceiling(palette_length/2) + 1), 
              seq(-breaks[1]/palette_length, -breaks[1], length.out=floor(palette_length/2)))


# pdf(paste0(figure_out, "rs20220623_indelfrequency_heatmap_HDAC.pdf"), width=16, height=8)
plt3 = pheatmap(HDAC_matrix[clone5_map_ipr$labels,],
                border_color = F,
                cellwidth = 8,
                cellheight = 8,
                breaks = myBreaks,
                cluster_rows = clone5_map_ipr,
                cluster_cols = T,
                angle_col = 90,
                na_col = "white",
                clustering_method = "ward.D",
                main = 'MMEJ ratio FC (log2), all concentrations',
                annotation_col = target.cluster1,
                annotation_row = chromatin,
                color = diverge_hcl(palette_length, "Blue-Red"),
                annotation_colors = annotation_colors_scr,
                annotation_legend = T,
                cutree_cols = 3)
```


```{r}
chromatin_ipr <- chromatin %>%
  rownames_to_column("IPR")

eff.data.heatmap.cluster2 <- eff.data.heatmap.cluster1 %>%
  pivot_longer(cols = contains("IPR"), names_to = "IPR", values_to = "MMEJratio_FC")

chromatin_effect <- left_join(eff.data.heatmap.cluster2, chromatin_ipr) %>%
  #filter(target == "HDAC") %>%
  mutate(MMEJratio_FC = ave(MMEJratio_FC, drug, chromatin, conc_char, FUN = function(x) mean(x, na.rm = TRUE)),
         drug_conc = paste(drug, conc_char, sep = " ")) %>%
  distinct(MMEJratio_FC, drug, chromatin, conc_char, target, drug_conc)

clusters_drugs = target_set_all # %>% filter(clusters_drugs %in% c(3, 4))


chromatin_effect <- left_join(cluster_HDAC, chromatin_effect)

ggplot(chromatin_effect,
       aes(x = chromatin, y = MMEJratio_FC, fill = chromatin)) +
  geom_boxplot(outlier.shape = NA) +
  geom_quasirandom() +
  geom_signif(comparisons = combn(sort(unique(chromatin_effect$chromatin)), 2, simplify = F), 
              map_signif_level=sigFunc, 
              step_increase = 0.1) +
  theme_classic_lines_45() +
  scale_fill_manual(values = group_colors) +
  theme(legend.position = "none") +
  facet_wrap(~ cluster_HDAC, nrow = 1)
```





# Session Info
```{r}
paste("Run time: ",format(Sys.time()-StartTime))
getwd()
date()
sessionInfo()
```

