---
title: "DSB Episcreen data analysis & plotting"
author: 
  - name: "Max Trauernicht"
    email: "m.trauernicht@nki.nl"
    affiliation: "Netherlands Cancer Institute"
date: '`r format(Sys.time(), "%d/%m/%Y")`'
output: 
  html_document:
    theme: united
    highlight: pygments
    fig_caption: yes
    code_folding: hide
    df_print: kable
    toc: true
    toc_depth: 4
    toc_float:
      collapsed: false
---

*knitr document van Steensel lab*


# Introduction
Clone 5 with ~18 sgRNA target site IPRs was cut and treated with 160 epigenetic drugs. The repair outcomes of that experiment will be visualized here. I previously processed the raw sequencing data, and calculated the repair outcomes. In this script, the main results plots will be generated.


## Setup {.tabset}

<!-- little HTML script to do indentation of the table of contents -->
<script>
    $(document).ready(function() {
      $items = $('div#TOC li');
      $items.each(function(idx) {
        num_ul = $(this).parentsUntil('#TOC').length;
        $(this).css({'text-indent': num_ul * 10, 'padding-left': 0});
      });

    });
</script>

```{css, echo = FALSE}
div.sourceCode {
  overflow-x: hidden;
}
```
### Libraries

```{r load_packages, results = 'hide', warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
StartTime <-Sys.time()

# 8-digit Date tag:
Date <- substr(gsub("-","",Sys.time()),1,8) 
# libraries:
library(dplyr)
library(outliers)
library(ggplot2)
library(ggbeeswarm)
library(data.table)
library(ggrepel)
library(GGally)
library(DESeq2)
library(gghighlight)
library(platetools)
library(ggpubr)
library(tidyr)
library(Laurae)
library(tibble)
library(pheatmap)
library(RColorBrewer)
library(ggrastr)
```

### Functions

```{r load_functions, echo = FALSE, warning = FALSE}

p.adjust <- function(p, method = p.adjust.methods, n = length(p))
{
    ## Methods 'Hommel', 'BH', 'BY' and speed improvements
    ## contributed by Gordon Smyth
    method <- match.arg(method)
    if(method == "fdr") method <- "BH"	# back compatibility
    nm <- names(p)
    p <- as.numeric(p)
    p0 <- setNames(p, nm)
    if(all(nna <- !is.na(p))) nna <- TRUE
    p <- p[nna]
    lp <- length(p)
    # stopifnot(n >= lp)
    if (n <= 1) return(p0)
    if (n == 2 && method == "hommel") method <- "hochberg"

    p0[nna] <-
	switch(method,
	       bonferroni = pmin(1, n * p),
	       holm = {
		   i <- seq_len(lp)
		   o <- order(p)
		   ro <- order(o)
		   pmin(1, cummax( (n - i + 1L) * p[o] ))[ro]
	       },
	       hommel = { ## needs n-1 >= 2 in for() below
		   if(n > lp) p <- c(p, rep.int(1, n-lp))
		   i <- seq_len(n)
		   o <- order(p)
		   p <- p[o]
		   ro <- order(o)
		   q <- pa <- rep.int( min(n*p/i), n)
		   for (j in (n-1):2) {
		       ij <- seq_len(n-j+1)
		       i2 <- (n-j+2):n
		       q1 <- min(j*p[i2]/(2:j))
		       q[ij] <- pmin(j*p[ij], q1)
		       q[i2] <- q[n-j+1]
		       pa <- pmax(pa,q)
		   }
		   pmax(pa,p)[if(lp < n) ro[1:lp] else ro]
	       },
	       hochberg = {
		   i <- lp:1L
		   o <- order(p, decreasing = TRUE)
		   ro <- order(o)
		   pmin(1, cummin( (n - i + 1L) * p[o] ))[ro]
	       },
	       BH = {
		   i <- lp:1L
		   o <- order(p, decreasing = TRUE)
		   ro <- order(o)
		   pmin(1, cummin( n / i * p[o] ))[ro]
	       },
	       BY = {
		   i <- lp:1L
		   o <- order(p, decreasing = TRUE)
		   ro <- order(o)
		   q <- sum(1L/(1L:n))
		   pmin(1, cummin(q * n / i * p[o]))[ro]
	       },
	       none = p)
    p0
}

SetFileName <- function(filename, initials) {
  # Set filename with extension and initials to make filename with date integrated.
  filename <- substitute(filename)
  initials <- substitute(initials)
  filename <- paste0(initials, Date, filename)
  filename
}

# From Fede:
# ggpairs custom functions
corColor <- function(data, mapping, color = I("black"), sizeRange = c(1, 3), ...) {

  x   <- eval_data_col(data, mapping$x)
  y   <- eval_data_col(data, mapping$y)
  r   <- cor(x, y)
  rt  <- format(r, digits = 3)
  tt  <- as.character(rt)
  cex <- max(sizeRange)

  # helper function to calculate a useable size
  percent_of_range <- function(percent, range) {
    percent * diff(range) + min(range, na.rm = TRUE)
  }

  # plot correlation coefficient
  p <- ggally_text(label = tt, mapping = aes(), xP = 0.5, yP = 0.5,
                   size = I(percent_of_range(cex * abs(r), sizeRange)), color = color, ...) +
    theme(panel.grid.minor=element_blank(),
          panel.grid.major=element_blank())

  corColors <- RColorBrewer::brewer.pal(n = 7, name = "RdYlBu")[2:6]

  if (r <= boundaries[1]) {
    corCol <- corColors[1]
  } else if (r <= boundaries[2]) {
    corCol <- corColors[2]
  } else if (r < boundaries[3]) {
    corCol <- corColors[3]
  } else if (r < boundaries[4]) {
    corCol <- corColors[4]
  } else {
    corCol <- corColors[5]
  }

  p <- p +
    theme(panel.background = element_rect(fill = corCol))

  return(p)
}


# stat_smooth_func <- function(mapping = NULL, data = NULL,
#                         geom = "smooth", position = "identity",
#                         ...,
#                         method = "auto",
#                         formula = y ~ x,
#                         se = TRUE,
#                         n = 80,
#                         span = 0.75,
#                         fullrange = FALSE,
#                         level = 0.95,
#                         method.args = list(),
#                         na.rm = FALSE,
#                         show.legend = NA,
#                         inherit.aes = TRUE,
#                         xpos = NULL,
#                         ypos = NULL) {
#   layer(
#     data = data,
#     mapping = mapping,
#     stat = StatSmoothFunc,
#     geom = geom,
#     position = position,
#     show.legend = show.legend,
#     inherit.aes = inherit.aes,
#     params = list(
#       method = method,
#       formula = formula,
#       se = se,
#       n = n,
#       fullrange = fullrange,
#       level = level,
#       na.rm = na.rm,
#       method.args = method.args,
#       span = span,
#       xpos = xpos,
#       ypos = ypos,
#       ...
#     )
#   )
# }
# 
# StatSmoothFunc <- ggproto("StatSmooth", Stat,
#                       
#                       setup_params = function(data, params) {
#                         # Figure out what type of smoothing to do: loess for small datasets,
#                         # gam with a cubic regression basis for large data
#                         # This is based on the size of the _largest_ group.
#                         if (identical(params$method, "auto")) {
#                           max_group <- max(table(data$group))
#                           
#                           if (max_group < 1000) {
#                             params$method <- "loess"
#                           } else {
#                             params$method <- "gam"
#                             params$formula <- y ~ s(x, bs = "cs")
#                           }
#                         }
#                         if (identical(params$method, "gam")) {
#                           params$method <- mgcv::gam
#                         }
#                         
#                         params
#                       },
#                       
#                       compute_group = function(data, scales, method = "auto", formula = y~x,
#                                                se = TRUE, n = 80, span = 0.75, fullrange = FALSE,
#                                                xseq = NULL, level = 0.95, method.args = list(),
#                                                na.rm = FALSE, xpos=NULL, ypos=NULL) {
#                         if (length(unique(data$x)) < 2) {
#                           # Not enough data to perform fit
#                           return(data.frame())
#                         }
#                         
#                         if (is.null(data$weight)) data$weight <- 1
#                         
#                         if (is.null(xseq)) {
#                           if (is.integer(data$x)) {
#                             if (fullrange) {
#                               xseq <- scales$x$dimension()
#                             } else {
#                               xseq <- sort(unique(data$x))
#                             }
#                           } else {
#                             if (fullrange) {
#                               range <- scales$x$dimension()
#                             } else {
#                               range <- range(data$x, na.rm = TRUE)
#                             }
#                             xseq <- seq(range[1], range[2], length.out = n)
#                           }
#                         }
#                         # Special case span because it's the most commonly used model argument
#                         if (identical(method, "loess")) {
#                           method.args$span <- span
#                         }
#                         
#                         if (is.character(method)) method <- match.fun(method)
#                         
#                         base.args <- list(quote(formula), data = quote(data), weights = quote(weight))
#                         model <- do.call(method, c(base.args, method.args))
#                         
#                         m = model
#                         eq <- substitute(italic(y) == a + b %.% italic(x)*","~~italic(r)^2~"="~r2, 
#                                          list(a = format(coef(m)[1], digits = 3), 
#                                               b = format(coef(m)[2], digits = 3), 
#                                               r2 = format(summary(m)$r.squared, digits = 3)))
#                         func_string = as.character(as.expression(eq))
#                         
#                         if(is.null(xpos)) xpos = min(data$x)*0.9
#                         if(is.null(ypos)) ypos = max(data$y)*0.9
#                         data.frame(x=xpos, y=ypos, label=func_string)
#                         
#                       },
#                       
#                       required_aes = c("x", "y")
# )

theme_classic_lines <- function() {
  theme_pubr(border = F, legend = "top") +
            theme(panel.grid.major = element_line(colour = "#adb5bd", size = 0.25),
                  strip.background = element_rect(fill = "#ced4da")
            )
}
theme_classic_lines_45 <- function() {
  theme_pubr(border = T, legend = "top", x.text.angle = 45) +
            theme(panel.grid.major = element_line(colour = "#adb5bd", size = 0.25),
                  strip.background = element_rect(fill = "#ced4da")
            )
}

theme_classic_lines_90 <- function() {
  theme_pubr(border = T, legend = "top", x.text.angle = 90) +
            theme(panel.grid.major = element_line(colour = "#adb5bd", size = 0.25),
                  strip.background = element_rect(fill = "#ced4da")
            )
}

theme_set(theme_classic_lines())

colors_diverse <- c("#264653", "#9AC1AE", "#5D987B", "#f2cc8f", "#e76f51")
colors_drugs <- c("#FA8D62", "#65BFA4", "#8CA0C4", "#808184", "#F6D289", "#E5E5E5", "#ACACAC", "#737373")
#colors_diverse <- c("#264653", "#2a9d8f", "#e9c46a", "#f4a261", "#e76f51") 

ggplot_custom <- function(...) ggplot2::ggplot(...) + 
  scale_color_manual(values = colors_diverse) + 
  scale_fill_manual(values = colors_diverse)

```


### Data import

```{r data import, echo = FALSE, warning = FALSE}
# Import data from preprocessing script
indel.data <- readRDS("/DATA/usr/m.trauernicht/projects/EpiScreen/files_scripts/mt20220105_episcreen.RDS")
```

---

## A: Cutting efficiency change per drug
Are there any drugs (per concentration) that significantly change the cutting efficiency?
```{r}
# A: Cutting efficiency per drug - vulcano plot
eff.data <- indel.data %>%
                dplyr::distinct(drug, freqCut_norm_mean_drug, freqCut_norm_split, conc_char, sample, target,
                                tech, replicate) %>%
  mutate(freqCut_norm_split = ave(freqCut_norm_split, drug, conc_char, replicate, tech, FUN = mean)) %>%
  unique()

for (i in unique(eff.data$drug)) {
  for (j in unique(eff.data$conc_char)) {
    if (nrow(eff.data[eff.data$drug == i & eff.data$conc_char == j]) >= 2) {
    eff.data$pval[eff.data$drug == i & eff.data$conc_char == j] <- 
      t.test(eff.data$freqCut_norm_split[eff.data$drug == i & eff.data$conc_char == j], 
                            eff.data$freqCut_norm_split[eff.data$drug == "DMSO"])$p.value
    }
  }
}
#eff.data$pval <- p.adjust(eff.data$pval, method = "fdr")


ggplot_custom(eff.data %>%
                filter(sample == "drug") %>%
                dplyr::select(-freqCut_norm_split) %>%
                unique(),
              aes(x = log2(freqCut_norm_mean_drug), y = -log10(pval), color = ifelse(pval < 0.05, "red", "black"))) +
  geom_quasirandom() +
  facet_wrap(~conc_char)
```
*Figure 2A: Compound-induced change in cutting efficiency of all tested compounds. Mean per drug & concentration vs. p-value from t-test.*

Conclusion: Yes, some drugs do alter cutting efficiency significantly. Changes go in both directions, at lower concentration the efficiency is mostly increased (at higher concentrations the decrease might be due to toxicity)

---

## B: Cutting efficiency change per target group
Are there any target groups that are prone to changes in cutting efficiency?
```{r}
# B: Cutting efficiency change per target group, all concentrations together
eff.data.2 <- eff.data %>%
                filter(sample != "DNA-PKi (NHEJ control)", sample != "Mirin (MMEJ control)", sample != "DMSO (neg control)", conc_char == "100 nM") %>%
                  dplyr::select(freqCut_norm_mean_drug, pval, sample, drug, conc_char, target) %>%
                  unique() %>%
                mutate(mean_eff_target = ave(freqCut_norm_mean_drug, target, FUN = mean),
                       pval = -log10(pval),
                       freqCut_norm_mean_drug = log2(freqCut_norm_mean_drug))

ggplot_custom(data = eff.data.2, 
              aes(y = freqCut_norm_mean_drug, x = reorder(target, -mean_eff_target))) + 
  geom_quasirandom(aes(color = eff.data.2$pval)) +
    geom_boxplot(position = position_dodge(0.75), alpha = 0.4) +
    xlab("") + ylab("log2(cutting efficiency)") +
      theme_classic_lines_90()+
  scale_color_viridis_b()
```
*Figure 2B: Compound-induced change in cutting efficiency per compound target groups.*

Conclusion: Yes, specifically HDACs seem to increase cutting efficiency. Some other interesting observations can also be made from this plot.





```{r export}
setwd("/DATA/usr/m.trauernicht/projects/EpiScreen/files_scripts/")
filename <- SetFileName("_indel.data", "mt")
save(indel.data, file = filename)
```

# Session Info
```{r}
paste("Run time: ",format(Sys.time()-StartTime))
getwd()
date()
sessionInfo()
```

