---
title: "Clone 5 annotation"
author: "Max Trauernicht"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output: 
  html_document:
    theme: journal #cerulean
    highlight: monochrome
    toc: true
    toc_float: true
    code_folding: show
  editor_options:
    chunk_output_type: console
---

# knitr document van Steensel lab

# TRIP Clone identification

# Introduction

## Description of Data
How to make a good rendering table: 
```{r table1, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
tabl <- "
| column1	|	column2	|	column3	|	
|----|----|----|
|1	|	2	|	3	|	
|a	|	b	|	c	|	
"
cat(tabl) # output the table in a format good for HTML/PDF/docx conversion
```

# Data processing
## Path, Libraries, Parameters and Useful Functions
```{r setup}
knitr::opts_chunk$set(echo = TRUE)
StartTime <-Sys.time()

# 8-digit Date tag:
Date <- substr(gsub("-","",Sys.time()),1,8) 
# libraries:
library(dplyr)
library(data.table)
library(devtools)
```

### Custom functions
Functions used thoughout this script.
```{r}
SetFileName <- function(filename, initials) {
  # Set filename with extension and initials to make filename with date integrated.
  filename <- substitute(filename)
  initials <- substitute(initials)
  filename <- paste0(initials, Date, filename)
  filename
}
```


## Data import
```{r}
# Load ChIP data
setwd("~/mydata/Mapping_Clone5_9/nearest_domains-1000-0.9")
file.list <- list.files(pattern = '*.txt')
ChIP.list <- lapply(file.list, read.table, header = TRUE)
mapping_integration <- as.data.frame(ChIP.list)

setwd("~/mydata/Mapping_Clone5_9/means")
file.list_mean <- list.files(pattern = '*.txt')
mapping.list_mean <- lapply(file.list_mean, read.table, header = TRUE)
mapping_mean <- as.data.frame(mapping.list_mean)

# Load Lmnb1 data
setwd("~/mydata/Mapping_Clone5_9")
lmn.df <- get(load("rs20190116_RSTP2.2000.Chromatin.RData"))

# Load clone 5 barcode data
clone5.bc <- read.delim("mt20190204_clone5_barcodes.txt")
```

# Analysis

## Pre-processing mapping_integration df
```{r}
# Remove not needed data
mapping_integration <- select(mapping_integration, ends_with("e"), contains("Distance"))

# Rename variables 
names(mapping_integration)[2:38] <- gsub(".*_.*_(.*?)_.*.txt", "\\1", file.list)

# Process lmn dataset
lmn.df.mean <- lmn.df[,c(1,101)]
lmn.df.mean <- lmn.df.mean[!duplicated(lmn.df.mean), ]
lmn.df.mean <- merge(lmn.df.mean, clone5.bc)
setnames(mapping_integration, old = "Barcode", new = "barcode")
mapping_integration <- merge(mapping_integration, clone5.bc)

# Combine the two chr. state dfs
mapping_integration <- merge(lmn.df.mean, mapping_integration, all = TRUE)

# Change from distance to integration: yes (1) or no (0)
mapping_integration[,3:ncol(mapping_integration)] <- ifelse(mapping_integration[,3:ncol(mapping_integration)] > 0, 0, 1)
mapping_integration[,2] <- ifelse(mapping_integration[,2] == "AD", 1, 0)

# Remove .B from barcode
mapping_integration$barcode <- gsub(".B","",mapping_integration$barcode)
```

## Pre-processing mapping_mean df
```{r}
# Remove not needed data
mapping_mean <- select(mapping_mean, ends_with("id"), starts_with("mean"))

# Rename variables 
names(mapping_mean)[2:47] <- gsub(".*_.*_(.*?)_.*.txt", "\\1", file.list)

# Process mapping_mean dataset
lmn.df <- lmn.df[,c(1,100)]
lmn.df <- lmn.df[!duplicated(lmn.df), ]
lmn.df <- merge(lmn.df, clone5.bc)
setnames(mapping_mean, old = "transcript_id", new = "barcode")
mapping_mean <- merge(mapping_mean, clone5.bc)

# Combine the two chr. state dfs
mapping_mean <- merge(lmn.df, mapping_mean, all = TRUE)

# Remove .B from barcode
mapping_mean$barcode <- gsub(".B","",mapping_mean$barcode)
```

## Normalization mean values
```{r}
# Average repetitive ChIP datasets & remove duplets
mapping_mean$H3K79me2 <- rowMeans(mapping_mean[,c("H3K79me2","H3K79me2.1","H3K79me2.2","H3K79me2.3")])
mapping_mean$H3K9me3 <- rowMeans(mapping_mean[,c("H3K9me3","H3K9me3.1","H3K9me3.2","H3K9me3.3","H3K9me3.4")])
mapping_mean$H3K9me2 <- rowMeans(mapping_mean[,c("H3K9me2","H3K9me2.1","H3K9me2.2")])
mapping_mean$H3K27ac <- rowMeans(mapping_mean[,c("H3K27ac","H3K27ac.1","H3K27ac.2")])
mapping_mean$H3K27me3 <- rowMeans(mapping_mean[,c("H3K27me3","H3K27me3.1","H3K27me3.2")])
mapping_mean$H3K36me3 <- rowMeans(mapping_mean[,c("H3K36me3","H3K36me3.1","H3K36me3.2")])
mapping_mean$H3K4me1 <- rowMeans(mapping_mean[,c("H3K4me1","H3K4me1.1","H3K4me1.2")])
mapping_mean$PU1 <- rowMeans(mapping_mean[,c("PU1","PU1.1")])
mapping_mean$POL2 <- rowMeans(mapping_mean[,c("POL2","POL2.1")])
mapping_mean$H2AFZ <- rowMeans(mapping_mean[,c("H2AFZ","H2AFZ.1")])
mapping_mean$EZH2 <- rowMeans(mapping_mean[,c("EZH2","EZH2.1")])
mapping_mean$H4K5acK8ac <- rowMeans(mapping_mean[,c("H4K5acK8ac","H4K5acK8ac.1")])
mapping_mean$H3K122ac <- rowMeans(mapping_mean[,c("H3K122ac","H3K122ac.1")])
mapping_mean$H3K122ac <- rowMeans(mapping_mean[,c("H3K122ac","H3K122ac.1")])
mapping_mean <- select(mapping_mean, -ends_with(".1"), -ends_with(".2"), -ends_with(".3"), -ends_with(".4"))

# Change df format to normalize mean values
mapping_mean_long <- melt(mapping_mean)
mapping_mean_long$value[is.na(mapping_mean_long$value)] <- 0
mapping_mean_long$log <- mapping_mean_long$value
mapping_mean_long$log[mapping_mean_long$variable != "LMNB1"] <- ave(mapping_mean_long$value[mapping_mean_long$variable != "LMNB1"], FUN = function(x) log2(x))
mapping_mean_long$mean <- ave(mapping_mean_long$log, mapping_mean_long$variable, FUN = function(x) x/max(x))
mapping_mean_long <- mapping_mean_long[,c(-3,-4)]
mapping_mean <- dcast(mapping_mean_long, variable ~ barcode, value.var = "mean")

# Add type of chr. mark (repressive or active)
mapping_mean_long$type <- "Active"
mapping_mean_long[grep("EZH2|K562|K9me|K79me3|K27me|macroH2A|BMI1|CBX2|119ub|
                          MEL18|RING|PCGF|MEL|RCOR|LMNB", mapping_mean_long$variable),]$type <- "Repressive"
```

## Create heatmaps
```{r}
## mapping_mean dataset
# Set Barcode column as rownames and remove Barcode column
row.names(mapping_mean) <- mapping_mean$variable
mapping_mean <- mapping_mean[,-1]

# Generate heatmap
pheatmap(as.matrix(mapping_mean)) 

## mapping_mean_essent dataset
mapping_mean_essent <- mapping_mean[c("LMNB1","EZH2","H3K79me2","H3K9me3","H3K9me2","H3K122ac","H3K27ac","H3K4me1","H3K27me3","H3K36me3","POL2"),]


# Annotate chr. marks
chromatin.mark_type1 <- mapping_mean_long[,c(-1,-3)]
chromatin.mark_type1 <- chromatin.mark_type1[!duplicated(chromatin.mark_type1), ]
chromatin.mark_type <- chromatin.mark_type1[,2]
chromatin.mark_type1 <- chromatin.mark_type1[,-2]
chromatin.mark_type1 <- as.data.frame(chromatin.mark_type1)
chromatin.mark_type <- as.data.frame(chromatin.mark_type)
row.names(chromatin.mark_type) <- chromatin.mark_type1$chromatin.mark_type1

# Annotate barcodes
LAD_integration <- mapping_integration[,c("LAD_state")]
LAD_integration <- as.data.frame(LAD_integration)
row.names(LAD_integration) <- mapping_integration$barcode
LAD_integration$LAD_integration[is.na(LAD_integration$LAD_integration)] <- 0
LAD_integration[LAD_integration$LAD_integration == 0,] <- "No"
LAD_integration[LAD_integration$LAD_integration == 1,] <- "Yes"

H3K27me3_integration <- mapping_integration[,c("H3K27me3")]
H3K27me3_integration <- as.data.frame(H3K27me3_integration)
row.names(H3K27me3_integration) <- mapping_integration$barcode
H3K27me3_integration$H3K27me3_integration[is.na(H3K27me3_integration$H3K27me3_integration)] <- 0
H3K27me3_integration[H3K27me3_integration$H3K27me3_integration == 0,] <- "No"
H3K27me3_integration[H3K27me3_integration$H3K27me3_integration == 1,] <- "Yes"

LAD_integration$H3K27me3_integration <- H3K27me3_integration$H3K27me3_integration

# Change colors of annotations
annotation_colors = list(
  LAD_integration = c(Yes="#66CC00", No="#CCCCCC"),
  H3K27me3_integration = c(Yes="#66CC00", No="#CCCCCC"),
  chromatin.mark_type = c(Active="#66CC00", Repressive="#CCCCCC"))

# Generate heatmap
pheatmap(as.matrix(mapping_mean_essent), annotation_col = LAD_integration, 
         annotation_row = chromatin.mark_type, annotation_colors = annotation_colors, legend_labels = "log2(")




test = matrix(rnorm(200), 20, 10)
test[1:10, seq(1, 10, 2)] = test[1:10, seq(1, 10, 2)] + 3
test[11:20, seq(2, 10, 2)] = test[11:20, seq(2, 10, 2)] + 2
test[15:20, seq(2, 10, 2)] = test[15:20, seq(2, 10, 2)] + 4
colnames(test) = paste("Test", 1:10, sep = "")
rownames(test) = paste("Gene", 1:20, sep = "")
#plot
annotation <- data.frame(Var1 = factor(1:10 %% 2 == 0, labels = c("Exp1", "None")), Var1 = factor(1:10 %% 2 == 0, labels = c("Exp1", "None")))
rownames(annotation) <- colnames(test) # check out the row names of annotation






```

# Results
```{r}
```

# Conclusions
```{r}
```

## Exporting potential data. 
```{r}
```

# Session Info
```{r}
paste("Run time: ",format(Sys.time()-StartTime))
getwd()
date()
sessionInfo()
```

