---
title: "Clone 5 annotation"
author: "Max Trauernicht"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  pdf_document:
    toc: yes
  html_document:
    code_folding: show
    highlight: monochrome
    theme: journal
    toc: yes
    toc_float: yes
  editor_options:
    chunk_output_type: console
---

# knitr document van Steensel lab

# TRIP Clone annotation

# Introduction
In this document, ChIP data from the location of the barcodes in RSTP2#5 will be imported in order to be able to characterize and annotate each barcode. In this way, we can identify which barcodes are in a rather active chromatin environment, but also barcodes that are in repressed environments. This will help us to understand the results we might get in future experiments (e.g. the epigenetic screening).

## Description of Data
How to make a good rendering table: 
```{r table1, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
tabl <- "
| column1	|	column2	|	column3	|	
|----|----|----|
|1	|	2	|	3	|	
|a	|	b	|	c	|	
"
cat(tabl) # output the table in a format good for HTML/PDF/docx conversion
```

# Data processing
## Path, Libraries, Parameters and Useful Functions
```{r setup}
knitr::opts_chunk$set(echo = TRUE)
StartTime <-Sys.time()

# 8-digit Date tag:
Date <- substr(gsub("-","",Sys.time()),1,8) 
# libraries:
library(dplyr)
library(data.table)
library(devtools)
library(RColorBrewer)
library(pheatmap)
```

### Custom functions
Functions used thoughout this script.
```{r}
SetFileName <- function(filename, initials) {
  # Set filename with extension and initials to make filename with date integrated.
  filename <- substitute(filename)
  initials <- substitute(initials)
  filename <- paste0(initials, Date, filename)
  filename
}
```


## Data import
```{r}
# Load ChIP data
setwd("/DATA/usr/m.trauernicht/Mapping_Clone5_9/nearest_domains-1000-0.9")
file.list <- list.files(pattern = '*.txt')
ChIP.list <- lapply(file.list, read.table, header = TRUE)
mapping_integration <- as.data.frame(ChIP.list)

setwd("/DATA/projects/DSBrepair/data/pipeline_20181208/means/")
file.list_mean <- list.files(pattern = '*.txt')
mapping.list_mean <- lapply(file.list_mean, read.table, stringsAsFactors = F, header = TRUE)
mapping_mean <- as.data.frame(mapping.list_mean)

# Load Lmnb1 data
setwd("/DATA/usr/m.trauernicht/Mapping_Clone5_9")
lmn.df <- get(load("rs20190116_RSTP2.2000.Chromatin.RData"))

# Load clone 5 barcode data
clone5.bc <- read.delim("mt20190204_clone5_barcodes.txt")

# Alternative LMNB1 data
setwd("/DATA/usr/c.leemans/projects/chip_ruben/")
lmnb.df <- read.delim("cl20181025_mean_dam_10kb.txt")


```

# Analysis

## Pre-processing mapping_integration df
```{r}
# Remove not needed data
mapping_integration <- select(mapping_integration, ends_with("e"), contains("Distance"))

# Rename variables 
names(mapping_integration)[2:38] <- gsub(".*_(.*?_.*?_.*?).txt", "\\1", file.list)

mapping_integration <- mapping_integration[,c("Barcode", "Bernstein2011_H2AFZ_ENCSR000APC", 
                       "Schmidl2015_H3K27me3_chipmentation", "Schmidl2015_H3K27ac_chipmentation",
                       "Schmidl2015_H3K4me1_chipmentation", "Schmidl2015_H3K36me3_chipmentation",
                       "Salzberg2017_H3K9me3_ab8898", "Salzberg2017_H3K9me2_ab1220",
                       "Pradeepa2016_H3K122ac_ab33309", "Fu2013_H3K79me2_GSM865283",
                       "Minoda2018_H4K5acK8ac_GSE113635", "vdBoom2016_H2AK119ub_GSM1319308",
                       "Snyder2016_POL2_ENCSR388QZF")]

# Rename variables 
colnames(mapping_integration) <- gsub(".*_(.*?)_.*", "\\1", colnames(mapping_integration))

# Process lmn dataset
lmn.df.mean <- lmn.df[,c(1,101)]
lmn.df.mean <- lmn.df.mean[!duplicated(lmn.df.mean), ]
lmn.df.mean <- merge(lmn.df.mean, clone5.bc)
setnames(mapping_integration, old = "Barcode", new = "barcode")
mapping_integration <- merge(mapping_integration, clone5.bc)

# Combine the two chr. state dfs
mapping_integration <- merge(lmn.df.mean, mapping_integration, all = TRUE)

# Change from distance to integration: yes (1) or no (0)
mapping_integration[,3:ncol(mapping_integration)] <- 
  ifelse(mapping_integration[,3:ncol(mapping_integration)] > 0, 0, 1)
mapping_integration[,2] <- ifelse(mapping_integration[,2] == "AD", 1, 0)

# Remove .B from barcode
mapping_integration$barcode <- gsub(".B","",mapping_integration$barcode)
```

## Pre-processing mapping_mean df
```{r}
# Remove not needed data
mapping_mean <- select(mapping_mean, ends_with("id"), starts_with("mean"))


# Rename variables 
names(mapping_mean)[2:47] <- gsub(".*_(.*?_.*?_.*?).txt", "\\1", file.list_mean)

mapping_mean <- mapping_mean[,c("transcript_id", "Bernstein2011_H2AFZ_ENCSR000APC", 
                       "Schmidl2015_H3K27me3_chipmentation", "Schmidl2015_H3K27ac_chipmentation",
                       "Schmidl2015_H3K4me1_chipmentation", "Schmidl2015_H3K36me3_chipmentation",
                       "Salzberg2017_H3K9me3_ab8898", "Salzberg2017_H3K9me2_ab1220",
                       "Pradeepa2016_H3K122ac_ab33309", "Fu2013_H3K79me2_GSM865283",
                       "Minoda2018_H4K5acK8ac_GSE113635", "vdBoom2016_H2AK119ub_GSM1319308",
                       "Snyder2016_POL2_ENCSR388QZF")]

# Rename variables 
colnames(mapping_mean) <- gsub(".*_(.*?)_.*", "\\1", colnames(mapping_mean))



# Process mapping_mean dataset
# Load alternative lmnb1 df
setnames(lmnb.df, old = "name", new = "barcode")
lmnb.df <- merge(lmnb.df, clone5.bc)
lmnb.df <- lmnb.df[,-3]
setnames(mapping_mean, old = "transcript_id", new = "barcode")
mapping_mean <- merge(mapping_mean, clone5.bc)

# Combine the two chr. state dfs
mapping_mean <- merge(lmnb.df, mapping_mean, all = TRUE)
setnames(mapping_mean, old = "mean", new = "LMNB1")

# Remove .B from barcode
mapping_mean$barcode <- gsub(".B","",mapping_mean$barcode)
```

## Normalization mean values
```{r}
# Change df format to normalize mean values
mapping_mean_long <- melt(mapping_mean)
mapping_mean_long$value[is.na(mapping_mean_long$value)] <- 0
mapping_mean_long$log <- mapping_mean_long$value
mapping_mean_long$log[mapping_mean_long$variable != "LMNB1"] <- 
  ave(mapping_mean_long$value[mapping_mean_long$variable != "LMNB1"], FUN = function(x) log2(x))
mapping_mean_long$mean <- 
  ave(mapping_mean_long$log, mapping_mean_long$variable, FUN = function(x) x/max(x))
mapping_mean_long <- mapping_mean_long[,c(-3,-4)]
mapping_mean <- dcast(mapping_mean_long, variable ~ barcode, value.var = "mean")

# Add type of chr. mark (repressive or active)
mapping_mean_long$type <- "Active"
mapping_mean_long[grep("EZH2|K562|K9me|K79me3|K27me|macroH2A|BMI1|CBX2|119ub|
                          MEL18|RING|PCGF|MEL|RCOR|LMNB", 
                       mapping_mean_long$variable),]$type <- "Repressive"
```

## Create code for heatmaps: mapping_mean dataset
```{r}
# Annotate chr. marks: here I create a df to annotate which chromatin marks are active/repressive, 
# this is being used in the heatmap later
chromatin.mark_type1 <- mapping_mean_long[,c(-1,-3)]
chromatin.mark_type1 <- chromatin.mark_type1[!duplicated(chromatin.mark_type1), ]
chromatin.mark_type <- chromatin.mark_type1[,2]
chromatin.mark_type1 <- chromatin.mark_type1[,-2]
chromatin.mark_type1 <- as.data.frame(chromatin.mark_type1)
chromatin.mark_type <- as.data.frame(chromatin.mark_type)
row.names(chromatin.mark_type) <- chromatin.mark_type1$chromatin.mark_type1

# Annotate barcodes
LAD_integration <- mapping_integration[,c("LAD_state")]
LAD_integration <- as.data.frame(LAD_integration)
row.names(LAD_integration) <- mapping_integration$barcode
LAD_integration$LAD_integration[is.na(LAD_integration$LAD_integration)] <- 0
LAD_integration[LAD_integration$LAD_integration == 0,] <- "No"
LAD_integration[LAD_integration$LAD_integration == 1,] <- "Yes"

H3K27me3_integration <- mapping_integration[,c("H3K27me3")]
H3K27me3_integration <- as.data.frame(H3K27me3_integration)
row.names(H3K27me3_integration) <- mapping_integration$barcode
H3K27me3_integration$H3K27me3_integration[is.na(H3K27me3_integration$H3K27me3_integration)] <- 0
H3K27me3_integration[H3K27me3_integration$H3K27me3_integration == 0,] <- "No"
H3K27me3_integration[H3K27me3_integration$H3K27me3_integration == 1,] <- "Yes"

LAD_integration$H3K27me3_integration <- H3K27me3_integration$H3K27me3_integration

# Change colors of annotations
annotation_colors = list(
  LAD_integration = c(Yes="#669966", No="grey90"),
  H3K27me3_integration = c(Yes="#669966", No="grey90"),
  chromatin.mark_type = c(Active="#669966", Repressive="grey90"))




# Set Barcode column as rownames and remove Barcode column
row.names(mapping_mean_essent) <- mapping_mean_essent$variable
mapping_mean_essent <- mapping_mean_essent[,-1]


## mapping_mean dataset
# Set Barcode column as rownames and remove Barcode column
row.names(mapping_mean) <- mapping_mean$variable
mapping_mean <- mapping_mean[,-1]





```


## Create code for heatmaps: mapping_integration dataset
```{r}
# Turn around cols and rows
mapping_integration_long <- melt(mapping_integration)
mapping_integration <- 
  dcast(mapping_integration_long, variable ~ barcode, value.var = "value")

# Annotation color
chromatin.mark_type2 <- chromatin.mark_type
chromatin.mark_type2 <- mapping_mean_long[,c(-1,-3)]
chromatin.mark_type2 <- 
  chromatin.mark_type2[!duplicated(chromatin.mark_type2), ]
chromatin.mark_type2$variable <- 
  as.character(chromatin.mark_type2$variable)
chromatin.mark_type2[chromatin.mark_type2$variable == "LMNB1",]$variable <- 
  "LAD_state"
chromatin.mark_type3 <- chromatin.mark_type2[,2]
chromatin.mark_type2 <- chromatin.mark_type2[,-2]
chromatin.mark_type2 <- as.data.frame(chromatin.mark_type2)
chromatin.mark_type3 <- as.data.frame(chromatin.mark_type3)
row.names(chromatin.mark_type3) <- 
  chromatin.mark_type2$chromatin.mark_type2
annotation_colors_2 = list(
  chromatin.mark_type3 = c(Active="chocolate1", Repressive="grey82"))


# Preprocess dataset for heatmap
mapping_integration[is.na(mapping_integration)] <- 0
row.names(mapping_integration) <- mapping_integration$variable
mapping_integration <- mapping_integration[,-1]


```

Next, heatmaps can be visualized, using either the dataframe containing only essential ChIP data or all ChIP data.

\newpage
# Results
```{r}
## Generate the heatmaps

# The following two heatmaps represent in color the intensity -
# of the chromatin marks at the barcodes integrated in RSTP2 clone 5

# Each chromatin mark is annotated with either repressive or active

# Additionally, LAD & H3K27me3 integration calling states are represented -
# per barcode (meaning: do these barcodes reside in a LAD/H3K27me3 domain?)

breaksList <- seq(-1.2,1,.001)

# Generate heatmap: chr. mark intensity of only essential ChIP datasets vs. barcodes
x <- pheatmap(as.matrix(mapping_mean), annotation_col = LAD_integration, 
         annotation_row = chromatin.mark_type, 
         annotation_colors = annotation_colors,
         color = colorRampPalette(rev(brewer.pal(n = 7, name = "RdGy")))(length(breaksList)),
         breaks = breaksList,
         annotation_legend = T,
         main = "RSTP2#5: chromatin mark intensity 
         per barcode (essential ChIP datasets)")

superheat(as.matrix(mapping_mean),bottom.label.text.angle = 90, yt = cor(mapping_mean)[,"TCTTTTGAGGAGCTGA"])
```

## Conclusions from this heatmap:
- LAD barcodes cluster together, but still have some intra-group variability
- The H3K27me3 integration state calling doesn't always overlap with high ChIP mean values (why?)
- There seems to be a high variability between barcodes (we have pure LADs (e.g. CATCC...), repressed barcodes (e.g. GAGC...), rather
boring barcodes (no marks, e.g. GCGCA...), and very active barcodes(e.g. TATG...))
- possible barcode clustering (from left to right): barcode 1-2 = very active, barcode 3-6: active, barcode 7-10: slightly repressed, barcode 11-14: slightly repressed 2, barcode 15-18: LADs
- active marks tend to cluster together
- LMNB1, H3K9me2, EZH2 and H3K122ac are very different from other marks
- possible barcode clustering (from left to right): 1-3: mainly active, 4: repressed, 5-6: active, 7-10: LADs, 11: active, 12-15: boring/slightly repressed, 16-18: repressed

```{r fig.height = 10, fig.width = 20}
# The following two heatmaps represent in color the status of integration -
# in a chromatin domain, since for some chromatin marks multiple datasets -
# (up to 3) were used, the state calling can be 0.333, 0.5, or 0.666 instead of 0 or 1

# Each chromatin mark is annotated with either repressive or active

# Generate heatmap: ChIP integration calling for all datasets vs. barcode
pheatmap(as.matrix(mapping_integration), annotation_row = chromatin.mark_type3, 
         annotation_colors = annotation_colors_2, color = brewer.pal(5, "PuBu"),
         annotation_legend = F,
         main = "RSTP2#5: chromatin integration state per barcode")
```

## Conclusions from these heatmaps:
- active/repressive marks tend to cluster together
- possible barcode clustering (from left to right): 1-5: LADs, 6-10: repressed, 11-18: active


"RSTP2#5: chromatin integration state per barcode (all ChIP datasets)":

- active/repressive marks tend to cluster together
- possible barcode clustering (from left to right): 1-3: repressed, 4-8: active, 9-10: slightly repressed, 11-15: LADs, 16-18: slightly active



# Conclusion
The heatmaps reveal many valuable information about the integrations of clone 5. 
We can now use this information to look at how individual barcodes responded in the drug screening.
We can also try to cluster barcodes based on the above heatmaps and use these clusters in the analysis of the epigenetic drug screening.







## Exporting potential data. 
```{r}
setwd("/DATA/usr/m.trauernicht/EpiScreen/epigenetic-screening-on-trip-clone/Data/")
filename <- SetFileName("_mapping_mean", "mt")
save(mapping_mean, file = filename)

filename <- SetFileName("_mapping_integration", "mt")
save(mapping_integration, file = filename)

filename <- SetFileName("_mapping_integration_long", "mt")
save(mapping_integration_long, file = filename)

filename <- SetFileName("_mapping_mean_long", "mt")
save(mapping_mean_long, file = filename)

```

# Session Info
```{r}
paste("Run time: ",format(Sys.time()-StartTime))
getwd()
date()
sessionInfo()
```

