---
title: "Clone 5 mapping"
author: "Max Trauernicht"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output: 
  html_document:
    theme: journal #cerulean
    highlight: monochrome
    toc: true
    toc_float: true
    code_folding: show
  editor_options:
    chunk_output_type: console
---

# knitr document van Steensel lab

# RSTP2 clone 5 mapping using tagmentation data

# Introduction
In this script, tagmentation mapping data is processed and analyzed by comparing it to iPCR mapping data. This way, I want to filter out the yet unknown 2 barcode locations of RSTP2#5.


## Description of Data
How to make a good rendering table: 
```{r table1, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
tabl <- "
| column1	|	column2	|	column3	|	
|----|----|----|
|1	|	2	|	3	|	
|a	|	b	|	c	|	
"
cat(tabl) # output the table in a format good for HTML/PDF/docx conversion
```

# Data processing
## Path, Libraries, Parameters and Useful Functions
```{r setup}
knitr::opts_chunk$set(echo = TRUE)
StartTime <-Sys.time()

# 8-digit Date tag:
Date <- substr(gsub("-","",Sys.time()),1,8) 
# libraries:
```

### Custom functions
Functions used thoughout this script.
```{r}
SetFileName <- function(filename, initials) {
  # Set filename with extension and initials to make filename with date integrated.
  filename <- substitute(filename)
  initials <- substitute(initials)
  filename <- paste0(initials, Date, filename)
  filename
}


roundUp <- function(x,to=5000)
{
  to*(x%/%to + as.logical(x%%to))
}

```


## Data import
```{r}
setwd("/DATA/usr/r.schep/projects/tagmap_ruben/")
header <- c(1:20)
tag.df.raw <- read.delim("rs20190218_Clone5_Rev.txt", header = F)

setwd("/DATA/usr/m.trauernicht/Mapping_Clone5_9/")
clone5.mapping <- read.table("mt20190218_RSTP2_clone5.bed")


```

# Analysis
## Pre-processing
```{r}
# Select important columns
tag.df.raw <- tag.df.raw[,c(3:10)]

# Remove NAs
tag.df.raw <- na.omit(tag.df.raw)

# Choose quality cutoff
# tag.df.raw <- tag.df.raw[tag.df.raw$V5 >= 40,]

# Remove low read counts
tag.df.raw <- tag.df.raw[tag.df.raw$V4 %in% names(which(table(tag.df.raw$V4) > 30)), ]

# Put close locations together by rounding
tag.df.raw$V4 <- roundUp(tag.df.raw$V4)

# Reduce df length by only selecting unique locations
tag.df <- tag.df.raw[!duplicated(tag.df.raw$V4), ]

# Import mapping df and round the same way
clone5.mapping <- clone5.mapping[,c(-3,-5)]
clone5.mapping$V2 <- roundUp(clone5.mapping$V2)

# Create dfs that only contain unmatched locations of both dfs
tag.df.unmatched <- tag.df[! tag.df$V4 %in% clone5.mapping$V2,]
clone5.mapping.unmatched <- clone5.mapping[! clone5.mapping$V2 %in% tag.df$V4,]
```



## Visualisation
```{r}
```

# Results
```{r}
```

# Conclusions
```{r}
```

## Exporting potential data. 
```{r}
```

# Session Info
```{r}
paste("Run time: ",format(Sys.time()-StartTime))
getwd()
date()
sessionInfo()
```

