---
title: "Clone 5 annotation"
author: "Max Trauernicht"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  pdf_document:
    toc: yes
  html_document:
    code_folding: show
    highlight: monochrome
    theme: journal
    toc: yes
    toc_float: yes
  editor_options:
    chunk_output_type: console
---

# knitr document van Steensel lab

# TRIP Clone annotation

# Introduction
In this document, ChIP data from the location of the barcodes in RSTP2#5 will be imported in order to be able to characterize and annotate each barcode. In this way, we can identify which barcodes are in a rather active chromatin environment or barcodes that are in repressed environments. This will help us to understand the results we might get in future experiments (e.g. the epigenetic screening).

## Description of Data
How to make a good rendering table: 
```{r table1, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
tabl <- "
| column1	|	column2	|	column3	|	
|----|----|----|
|1	|	2	|	3	|	
|a	|	b	|	c	|	
"
cat(tabl) # output the table in a format good for HTML/PDF/docx conversion
```

# Data processing
## Path, Libraries, Parameters and Useful Functions
```{r setup}
knitr::opts_chunk$set(echo = TRUE)
StartTime <-Sys.time()

# 8-digit Date tag:
Date <- substr(gsub("-","",Sys.time()),1,8) 
# libraries:
library(dplyr)
library(data.table)
library(devtools)
library(RColorBrewer)
library(pheatmap)
library(rtracklayer)
library(GenomicRanges)
```

### Custom functions
Functions used thoughout this script.
```{r}
SetFileName <- function(filename, initials) {
  # Set filename with extension and initials to make filename with date integrated.
  filename <- substitute(filename)
  initials <- substitute(initials)
  filename <- paste0(initials, Date, filename)
  filename
}
```


## Data import
```{r out.width= "80%", fig.align= "center", echo = FALSE, warning = FALSE}
# Load ChIP data
setwd("/DATA/usr/m.trauernicht/Mapping_Clone5_9/nearest_domains-1000-0.9")
file.list <- list.files(pattern = '*.txt')
ChIP.list <- lapply(file.list, read.table, header = TRUE)
mapping_integration <- as.data.frame(ChIP.list)

setwd("/DATA/projects/DSBrepair/data/cl20190220_chip_clone5/means/")
file.list_mean <- list.files(pattern = '*.txt')
mapping.list_mean <- lapply(file.list_mean, read.table, stringsAsFactors = F, header = TRUE)
mapping_mean <- as.data.frame(mapping.list_mean)

# Load clone 5 barcode data
setwd("/DATA/usr/m.trauernicht/Mapping_Clone5_9")
clone5.bc <- read.delim("mt20190204_clone5_barcodes.txt")
clone5_lmnb1 <- read.table("mt20190218_RSTP2_clone5.bed")

# Import LAD calling from Tom
LAD.HMM.gr <- import.bed("/home/t.v.schaik/mydata/proj/3D_nucleus/results/ts180110_4DN_DataProcessing/results/HMM/bin-5kb/K562_LMNB1-5kb-combined_HMM.txt.gz")

# K562 Lamina
setwd("/home/t.v.schaik/mydata/proj/3D_nucleus/results/ts180110_4DN_DataProcessing/results/normalized/bin-5kb/")
DamID.bins.norm.5kb <- import.bedGraph("K562_LMNB1-5kb-combined.norm.txt.gz")
colnames(mcols(DamID.bins.norm.5kb)) <- "Lamina"


```

# Analysis

## Import and compute LMNB1 data
```{r out.width= "80%", fig.align= "center", echo = FALSE, warning = FALSE}
# Rename columns of bed file
names(clone5_lmnb1) <- c("seqnames","start","end","barcode","ori")

# Increase width of integration window
clone5_lmnb1$end <- clone5_lmnb1$start + 1000

# Make genomic ranges
clone5_lmb1.1 <- makeGRangesFromDataFrame(clone5_lmnb1, start.field = "start", end.field = "end", keep.extra.columns = T)
clone5_lmb1.1 <- narrow(clone5_lmb1.1,start=500, end=-500) ##why

# Compute DamID scores for LmnB1
ovl <- findOverlaps(clone5_lmb1.1, DamID.bins.norm.5kb)
mcols(clone5_lmb1.1)$K562.lamina <- mcols(DamID.bins.norm.5kb)$Lamina [subjectHits(ovl)]

# LAD calling from Tom
ovl <- findOverlaps(clone5_lmb1.1, LAD.HMM.gr)

# overlaps
mcols(clone5_lmb1.1)$LAD_state <- "NA"
mcols(clone5_lmb1.1)$LAD_state[queryHits(ovl)] <- mcols(LAD.HMM.gr)$name[subjectHits(ovl)]

# Back to dataframe
clone5_lmnb1.dt <- GenomicRanges::as.data.frame(clone5_lmb1.1)
clone5_lmnb1.dt <- clone5_lmnb1.dt[,c(6,8,9)]
```



## Pre-processing mapping_integration df
```{r out.width= "80%", fig.align= "center", echo = FALSE, warning = FALSE}
# Remove not needed data
mapping_integration <- dplyr::select(mapping_integration, ends_with("e"), contains("Distance"))

# Rename variables 
names(mapping_integration)[2:38] <- gsub(".*_(.*?_.*?_.*?).txt", "\\1", file.list)

mapping_integration <- mapping_integration[,c("Barcode", "Bernstein2011_H2AFZ_ENCSR000APC", 
                       "Schmidl2015_H3K27me3_chipmentation", "Schmidl2015_H3K27ac_chipmentation",
                       "Schmidl2015_H3K4me1_chipmentation", "Schmidl2015_H3K36me3_chipmentation",
                       "Salzberg2017_H3K9me3_ab8898", "Salzberg2017_H3K9me2_ab1220",
                       "Pradeepa2016_H3K122ac_ab33309", "Fu2013_H3K79me2_GSM865283",
                       "Minoda2018_H4K5acK8ac_GSE113635", "vdBoom2016_H2AK119ub_GSM1319308",
                       "Snyder2016_POL2_ENCSR388QZF")]

# Rename variables 
colnames(mapping_integration) <- gsub(".*_(.*?)_.*", "\\1", colnames(mapping_integration))

# Only select clone 5 data
setnames(mapping_integration, old = "Barcode", new = "barcode")
mapping_integration <- merge(mapping_integration, clone5.bc)

# Combine the two chr. state dfs
clone5_lmnb1.dt_state <- clone5_lmnb1.dt[,c(1,3)]
mapping_integration <- merge(clone5_lmnb1.dt_state, mapping_integration, all = TRUE)

# Change from distance to integration: yes (1) or no (0)
mapping_integration[,3:ncol(mapping_integration)] <- 
  ifelse(mapping_integration[,3:ncol(mapping_integration)] > 0, 0, 1)
mapping_integration[,2] <- ifelse(mapping_integration[,2] == "AD", 1, 0)

# Remove .B from barcode
mapping_integration$barcode <- gsub(".B","",mapping_integration$barcode)
```

## Pre-processing mapping_mean df
```{r}
# Remove not needed data
mapping_mean_all <- dplyr::select(mapping_mean, ends_with("id"), starts_with("mean"))

# Rename variables 
names(mapping_mean_all)[2:ncol(mapping_mean_all)] <- gsub(".*_(.*?_.*?_.*?).txt", "\\1", file.list_mean)

mapping_mean <- mapping_mean_all[,c("transcript_id", "Bernstein2011_H2AFZ_ENCSR000APC", 
                       "Schmidl2015_H3K27me3_chipmentation", "Schmidl2015_H3K27ac_chipmentation",
                       "Schmidl2015_H3K4me1_chipmentation", "Schmidl2015_H3K36me3_chipmentation",
                       "Salzberg2017_H3K9me3_ab8898", "Salzberg2017_H3K9me2_ab1220",
                       "Pradeepa2016_H3K122ac_ab33309", "Fu2013_H3K79me2_GSM865283",
                       "Minoda2018_H4K5acK8ac_GSE113635", "vdBoom2016_H2AK119ub_GSM1319308",
                       "Snyder2016_POL2_ENCSR388QZF")]

# Rename variables 
colnames(mapping_mean) <- gsub(".*_(.*?)_.*", "\\1", colnames(mapping_mean))
colnames(mapping_mean_all) <- gsub(".*_(.*?)_.*", "\\1", colnames(mapping_mean_all))

# Rename variable
setnames(mapping_mean, old = "transcript_id", new = "barcode")
setnames(mapping_mean_all, old = "transcript_id", new = "barcode")

# Add lmnb1 data to df
clone5_lmnb1.dt_mean <- clone5_lmnb1.dt[,c(1,2)]
clone5_lmnb1.dt_mean$barcode <- gsub(".B","",clone5_lmnb1.dt_mean$barcode)
mapping_mean <- merge(clone5_lmnb1.dt_mean, mapping_mean, all = TRUE)
setnames(mapping_mean, old = "K562.lamina", new = "LMNB1")
mapping_mean_all <- merge(clone5_lmnb1.dt_mean, mapping_mean_all, all = TRUE)
setnames(mapping_mean_all, old = "K562.lamina", new = "LMNB1")
```



## Normalization mean values
```{r out.width= "80%", fig.align= "center", echo = FALSE, warning = FALSE}
# Change df format to normalize mean values
mapping_mean_long <- melt(mapping_mean)
mapping_mean_long$value[is.na(mapping_mean_long$value)] <- 0
mapping_mean_long$log <- mapping_mean_long$value
mapping_mean_long$log[mapping_mean_long$variable != "LMNB1"] <- 
  ave(mapping_mean_long$value[mapping_mean_long$variable != "LMNB1"], FUN = function(x) log2(x))
mapping_mean_long$mean <- 
  ave(mapping_mean_long$log, mapping_mean_long$variable, FUN = function(x) x/max(x))
mapping_mean_long <- mapping_mean_long[,c(-3,-4)]
mapping_mean <- dcast(mapping_mean_long, variable ~ barcode, value.var = "mean")

# The same for all ChIP datasets
mapping_mean_long_all <- melt(mapping_mean_all)
mapping_mean_long_all$value[is.na(mapping_mean_long_all$value)] <- 0
mapping_mean_long_all$log <- mapping_mean_long_all$value
mapping_mean_long_all$log[mapping_mean_long_all$variable != "LMNB1"] <- 
  ave(mapping_mean_long_all$value[mapping_mean_long_all$variable != "LMNB1"], FUN = function(x) log2(x))
mapping_mean_long_all$mean <- 
  ave(mapping_mean_long_all$log, mapping_mean_long_all$variable, FUN = function(x) x/max(x))
mapping_mean_long_all <- mapping_mean_long_all[,c(-3,-4)]
mapping_mean_all <- dcast(mapping_mean_long_all, variable ~ barcode, value.var = "mean")


# Add type of chr. mark (repressive or active)
mapping_mean_long$type <- "Active"
mapping_mean_long[grep("EZH2|K562|K9me|K79me3|K27me|macroH2A|BMI1|CBX2|119ub|
                          MEL18|RING|PCGF|MEL|RCOR|LMNB", 
                       mapping_mean_long$variable),]$type <- "Repressive"
```

## Create code for heatmaps: mapping_mean dataset
```{r out.width= "80%", fig.align= "center", echo = FALSE, warning = FALSE}
# Annotate chr. marks: here I create a df to annotate which chromatin marks are active/repressive, 
# this is being used in the heatmap later
chromatin.mark_type1 <- mapping_mean_long[,c(-1,-3)]
chromatin.mark_type1 <- chromatin.mark_type1[!duplicated(chromatin.mark_type1), ]
chromatin.mark_type <- chromatin.mark_type1[,2]
chromatin.mark_type1 <- chromatin.mark_type1[,-2]
chromatin.mark_type1 <- as.data.frame(chromatin.mark_type1)
chromatin.mark_type <- as.data.frame(chromatin.mark_type)
row.names(chromatin.mark_type) <- chromatin.mark_type1$chromatin.mark_type1
setnames(chromatin.mark_type, "chromatin.mark_type", "Type of Chromatin Mark")

# Annotate barcodes
LAD_integration <- mapping_integration[,c("LAD_state")]
LAD_integration <- as.data.frame(LAD_integration)
row.names(LAD_integration) <- mapping_integration$barcode
LAD_integration$LAD_integration[is.na(LAD_integration$LAD_integration)] <- 0
LAD_integration[LAD_integration$LAD_integration == 0,] <- "No"
LAD_integration[LAD_integration$LAD_integration == 1,] <- "Yes"

barcode_annotation <- LAD_integration
barcode_annotation$LAD_integration[c(11,20,5,3,19)] <- "Enhancer Region"
barcode_annotation$LAD_integration[c(14,9)] <- "H3K9me2/3 Domain"
barcode_annotation$LAD_integration[c(4,16)] <- "Active Gene Body"
barcode_annotation$LAD_integration[c(6,15,7,17)] <- "LAD"
barcode_annotation$LAD_integration[c(8,18)] <- "No Marks Present"
barcode_annotation$LAD_integration[c(2,12,1,10,13)] <- "H3K27me3 Domain"
setnames(barcode_annotation, "LAD_integration", "Integration Site")

H3K27me3_integration <- mapping_integration[,c("H3K27me3")]
H3K27me3_integration <- as.data.frame(H3K27me3_integration)
row.names(H3K27me3_integration) <- mapping_integration$barcode
H3K27me3_integration$H3K27me3_integration[is.na(H3K27me3_integration$H3K27me3_integration)] <- 0
H3K27me3_integration[H3K27me3_integration$H3K27me3_integration == 0,] <- "No"
H3K27me3_integration[H3K27me3_integration$H3K27me3_integration == 1,] <- "Yes"

LAD_integration$H3K27me3_integration <- H3K27me3_integration$H3K27me3_integration

# barcode_annotation$`Integration Site` <- 
#   factor(barcode_annotation$`Integration Site`, 
#          levels(barcode_annotation$`Integration Site`)[c("Enhancer Region", "H3K9me2/3 Domain", 
#                     "Active Gene Body", "LAD",  "No Marks Present", "H3K27me3 Domain")])

# Change colors of annotations
colors <- brewer.pal(length(unique(barcode_annotation$`Integration Site`)), "Pastel2")
names(colors) <- unique(barcode_annotation$`Integration Site`)

annotation_colors = list(
  LAD_integration = c(Yes="#669966", No="grey90"),
  H3K27me3_integration = c(Yes="#669966", No="grey90"),
  `Type of Chromatin Mark` = c(Active="#6AB163", Repressive="grey90"),
  `Integration Site` = colors)


# Set ChIP column as rownames and remove ChIP column
row.names(mapping_mean) <- mapping_mean$variable
mapping_mean <- mapping_mean[,-1]
```


## Create code for heatmaps: mapping_integration dataset
```{r out.width= "80%", fig.align= "center", echo = FALSE, warning = FALSE}
# Turn around cols and rows
mapping_integration_long <- melt(mapping_integration)
mapping_integration <- 
  dcast(mapping_integration_long, variable ~ barcode, value.var = "value")

# Annotation color
chromatin.mark_type2 <- chromatin.mark_type
chromatin.mark_type2 <- mapping_mean_long[,c(-1,-3)]
chromatin.mark_type2 <- 
  chromatin.mark_type2[!duplicated(chromatin.mark_type2), ]
chromatin.mark_type2$variable <- 
  as.character(chromatin.mark_type2$variable)
chromatin.mark_type2[chromatin.mark_type2$variable == "LMNB1",]$variable <- 
  "LAD_state"
chromatin.mark_type3 <- chromatin.mark_type2[,2]
chromatin.mark_type2 <- chromatin.mark_type2[,-2]
chromatin.mark_type2 <- as.data.frame(chromatin.mark_type2)
chromatin.mark_type3 <- as.data.frame(chromatin.mark_type3)
row.names(chromatin.mark_type3) <- 
  chromatin.mark_type2$chromatin.mark_type2
annotation_colors_2 = list(
  chromatin.mark_type3 = c(Active="chocolate1", Repressive="grey82"))


# Preprocess dataset for heatmap
mapping_integration[is.na(mapping_integration)] <- 0
row.names(mapping_integration) <- mapping_integration$variable
mapping_integration <- mapping_integration[,-1]


```

Next, heatmaps can be visualized, using either the dataframe containing only essential ChIP data or all ChIP data.

\newpage
# Results
```{r out.width= "80%", fig.align= "center", echo = FALSE, warning = FALSE}
## Generate the heatmaps

# The following two heatmaps represent in color the intensity -
# of the chromatin marks at the barcodes integrated in RSTP2 clone 5

# Each chromatin mark is annotated with either repressive or active

# Additionally, LAD & H3K27me3 integration calling states are represented -
# per barcode (meaning: do these barcodes reside in a LAD/H3K27me3 domain?)

breaksList <- c(seq(min(as.matrix(mapping_mean)), 0, 
                   length.out=ceiling(100/2) + 1), 
              seq(max(as.matrix(mapping_mean))/100, 
                  max(as.matrix(mapping_mean)), length.out=floor(100/2)))
breaksList2 <- seq(-1,1,0.02)

# Generate heatmap: chr. mark intensity of only essential ChIP datasets vs. barcodes
pheatmap(as.matrix(mapping_mean), 
         annotation_row = chromatin.mark_type, 
         annotation_col = barcode_annotation,
         annotation_colors = annotation_colors,
         fontsize_row = 12, annotation_names_row = F, 
         fontsize_col = 12, annotation_names_col = F,
         color = colorRampPalette(rev(brewer.pal(n = 7, name = "RdBu")))(100),
         breaks = breaksList2,
         annotation_legend = T,
         main = "RSTP2#5: chromatin mark intensity per barcode")


x <- pheatmap(as.matrix(mapping_mean[,c(-1,-11)]),
         annotation_row = chromatin.mark_type, 
         annotation_col = barcode_annotation,
         fontsize_row = 12, annotation_names_row = F, 
         fontsize_col = 12, annotation_names_col = F,
         annotation_colors = annotation_colors,
         color = colorRampPalette(rev(brewer.pal(n = 7, name = "RdBu")))(100),
         breaks = breaksList2,
         main = "RSTP2#5: chromatin mark intensity per barcode")
```

## Conclusions from this heatmap:
- LAD barcodes cluster together, but still have some intra-group variability
- The H3K27me3 integration state calling doesn't always overlap with high ChIP mean values (why?)
- There seems to be a high variability between barcodes (we have pure LADs (e.g. CATCC...), repressed barcodes (e.g. GAGC...), rather
boring barcodes (no marks, e.g. GCGCA...), and very active barcodes(e.g. TATG...))
- possible barcode clustering (from left to right): barcode 1-2 = very active, barcode 3-6: active, barcode 7-10: slightly repressed, barcode 11-14: slightly repressed 2, barcode 15-18: LADs
- active marks tend to cluster together
- LMNB1, H3K9me2, EZH2 and H3K122ac are very different from other marks
- possible barcode clustering (from left to right): 1-3: mainly active, 4: repressed, 5-6: active, 7-10: LADs, 11: active, 12-15: boring/slightly repressed, 16-18: repressed


## Conclusions from these heatmaps:
- active/repressive marks tend to cluster together
- possible barcode clustering (from left to right): 1-5: LADs, 6-10: repressed, 11-18: active


"RSTP2#5: chromatin integration state per barcode (all ChIP datasets)":

- active/repressive marks tend to cluster together
- possible barcode clustering (from left to right): 1-3: repressed, 4-8: active, 9-10: slightly repressed, 11-15: LADs, 16-18: slightly active



# Conclusion
The heatmaps reveal many valuable information about the integrations of clone 5. 
We can now use this information to look at how individual barcodes responded in the drug screening.
We can also try to cluster barcodes based on the above heatmaps and use these clusters in the analysis of the epigenetic drug screening.



## Exporting potential data. 
```{r}
setwd("/DATA/usr/m.trauernicht/EpiScreen/epigenetic-screening-on-trip-clone/Data/")
filename <- SetFileName("_mapping_mean", "mt")
save(mapping_mean, file = filename)

filename <- SetFileName("_mapping_integration", "mt")
save(mapping_integration, file = filename)

filename <- SetFileName("_mapping_integration_long", "mt")
save(mapping_integration_long, file = filename)

filename <- SetFileName("_mapping_mean_long", "mt")
save(mapping_mean_long, file = filename)

filename <- SetFileName("_chip_dendogram", "mt")
save(x, file = filename)

filename <- SetFileName("_mapping_mean_all", "mt")
save(mapping_mean_all, file = filename)
```

# Session Info
```{r}
paste("Run time: ",format(Sys.time()-StartTime))
getwd()
date()
sessionInfo()
```

