---
title: "xv20211214_epistasis_code"
output: html_document
---


# Libraries
```{r libraries}
# libraries:
library(tidyverse)
library(data.table)
library(parallel)
library(gtools)
library(pheatmap)
library(ggbeeswarm)
library(reshape2)
library(gridExtra)
library(broom)
library(ggrepel)
library(dendextend)
library(Hmisc)
library(RColorBrewer)
library(magrittr)
```

# Load libraries (Just need a data table with all the data and )
```{r}
#Chromatin data (it needs a barcode column)
screen_ratios_tib <- readRDS("/DATA/projects/DSBrepair/data/R/rs20220119_episcreen2/rs20220119_episcreen_ratios.RDS")
clone5_chrom_tib <- readRDS('/DATA/projects/DSBrepair/data/R/cl20201026_ChIP_zscore_selection.RDS')


chromatin.features <- colnames(screen_ratios_tib)[grepl("[.]zscore", colnames(screen_ratios_tib))]
chrom_data = screen_ratios_tib %>% select(barcode, all_of(chromatin.features)) %>% distinct()
```

#Example with three genes (POLL, FANCM, POLQ) & I will use "H3K4me1" as example
1st step = Compare POLL, FANCM, POLQ and WT data points vs. H3K4me1 (in lin or log space)

step 1 = plot 

```{r}
# Data table and compute log2_ratio (This line needs to be adjusted)
# log2.eff.screen.detail <- screen_ratios_tib %>% 
#   distinct(replicate, eff_score_mean, drug, conc_char, target, barcode) %>% 
#   mutate(log2_ratio = log2(eff_score_mean))

eff.screen.detail <- screen_ratios_tib %>% 
  select(replicate, tech, conc_char, barcode, drug, target, eff_zscore) %>% 
  distinct()

# DMSO data table set. (This lin of code needs to be adjusted)
# DMSO.set <- filter(screen_ratios_tib, drug == "DMSO") %>% 
#   select(replicate, eff_score_mean, drug, conc_char, target, barcode) %>% 
#   mutate(log2_ratio = log2(eff_score_mean)) %>% 
#   dplyr::group_by(barcode, replicate) %>% 
#   dplyr::summarise(DMSO.log2_ratio = mean(log2_ratio, na.rm = T), 
#                    DMSO.ratio = mean(eff_score_mean, na.rm = T))
DMSO.set <- filter(screen_ratios_tib, drug == "DMSO") %>% 
  distinct(replicate, tech, conc_char, barcode, drug, target, eff_zscore, mean_eff_zscore)

# Plot in log2 space
ggplot(data.for.plotting %>% filter(drug %in% c("DNA-PKi","Decitabine","GSK J4 HCl","DMSO"), conc_char == "10 µM"), aes(H3K4me1.zscore,m.log2_ratio, color = drug)) +
  geom_point() +
  geom_smooth(method = "lm") +
  stat_cor(method = "spearman") +
  theme_bw()

ggplot(data.for.plotting %>% filter(drug %in% c("DNA-PKi","Decitabine","GSK J4 HCl","DMSO"), conc_char == "10 µM"), aes(LMNB1.zscore,m.log2_ratio, color = drug)) +
  geom_point() +
  geom_smooth(method = "lm") +
  stat_cor(method = "spearman") +
  theme_bw()
```


# 2nd step check compute distances and plot them
```{r}
# Each replicate has a different value (compute differences by replicate)
log2.distance.eff <- log2.eff.screen.detail %>% 
  left_join(DMSO.set) %>% 
  mutate(log2.dist = sqrt((log2_ratio - DMSO.log2_ratio)^2), 
         lin.dist = sqrt((freqCut - DMSO.ratio)^2)) 

mean.log2.distance.eff <- log2.distance.eff %>% 
  dplyr::group_by(barcode, drug, conc_char) %>% 
  dplyr::summarise(m.log2.dist = mean(log2.dist, na.rm = T), 
                   m.lin.dist = mean(lin.dist, na.rm = T)) %>% 
  left_join(chrom_data)

#Plot slopes
ggplot(mean.log2.distance.eff %>% 
         filter(drug %in% c("DMSO","DNA-PKi","Decitabine")),
       aes(H3K27me3.zscore,m.log2.dist, color = drug)) + 
  geom_point() + 
  stat_cor(method = "spearman") + 
  geom_smooth(method = "lm") + 
  theme_bw() +
  facet_wrap(. ~ conc_char)


```



# Extract slopes for all genes
```{r}
# Perform analysis across all features for three test genes

mean.log2.distance.eff %<>% mutate(drug_conc = paste(drug, conc_char, sep = "_"))


slope.protein.features <- tibble(drug_conc = NA, feature = NA, slope.log2 = NA, term = NA, p.value = NA)

for (i in unique(mean.log2.distance.eff$drug_conc)) {
  for (j in chromatin.features) {
    model.dt <- mean.log2.distance.eff %>% filter(drug_conc == i & m.log2.dist != Inf) 
    
    model.epistasis.log2 <- lm(formula = m.log2.dist ~ unlist(model.dt[j]), 
                               data = model.dt) %>% 
      tidy()
    
    slope.protein.features <- slope.protein.features %>% 
      add_row(drug_conc = i, 
              feature = j, 
              slope.log2 = model.epistasis.log2 %>% 
                pull(estimate), 
              term = model.epistasis.log2 %>%
                pull(term),
              p.value = model.epistasis.log2 %>% 
                pull(p.value))
  }
}

# Plot matrix without scaling
# log2 space
ggplot(slope.protein.features %>% 
         filter(term != "(Intercept)" & 
                  complete.cases(.) & 
                  grepl("100 nM", drug_conc))) + 
  geom_tile(aes(drug_conc,feature, fill = slope.log2)) + 
  scale_fill_gradient2() + 
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust =0.5))

ggplot(slope.protein.features %>% 
         filter(term != "(Intercept)" & 
                  complete.cases(.) & 
                  grepl("1 µM", drug_conc))) + 
  geom_tile(aes(drug_conc,feature, fill = slope.log2)) + 
  scale_fill_gradient2() + 
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust =0.5))

ggplot(slope.protein.features %>% 
         filter(term != "(Intercept)" & 
                  complete.cases(.) & 
                  grepl("10 µM", drug_conc))) + 
  geom_tile(aes(drug_conc,feature, fill = slope.log2)) + 
  scale_fill_gradient2() + 
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust =0.5))

slope.protein.features2 = slope.protein.features 


hits2 = slope.protein.features %>% filter(term == "unlist(model.dt[j])" & p.value < 0.001) %>% pull(drug_conc) %>% unique()

hits = slope.protein.features %>% filter(term == "unlist(model.dt[j])" & abs(slope.log2) > 0.05) %>% pull(drug_conc) %>% unique()

slope.prot.features.scale.dcast.list <- slope.protein.features %>% 
  filter(term == "unlist(model.dt[j])" & drug_conc %in% hits) %>% 
  separate(drug_conc, into = c("drug", "conc"), sep = "_") %>% 
  group_by(conc) %>%
  group_split(.keep = FALSE) 

slope.prot.features.scale.dcast.1um = slope.prot.features.scale.dcast.list[[1]] %>%
  reshape2::dcast(drug ~ feature, value.var = "slope.log2") %>% 
  column_to_rownames(var = "drug")


slope.prot.features.scale.dcast.10um = slope.prot.features.scale.dcast.list[[2]] %>%
  reshape2::dcast(drug ~ feature, value.var = "slope.log2") %>% 
  column_to_rownames(var = "drug")


slope.prot.features.scale.dcast.100nm = slope.prot.features.scale.dcast.list[[3]] %>%
  reshape2::dcast(drug ~ feature, value.var = "slope.log2") %>% 
  column_to_rownames(var = "drug")
 
```



```{r}
# Set annotation for heatmap
# Row annotation: annotate drugs with the target group
target <- screen_ratios_tib %>% 
  distinct(drug, target) %>% 
  column_to_rownames(var="drug")

# colors <- brewer.pal(length(unique(barcode.annotation$cluster)), "Pastel2")
# names(colors) <- unique(barcode.annotation$cluster)

annotation_colors_scr = list(
  target = c(HDAC="#60988D", HAT="#C6D0A8", Sirtuin="#8FCAAC",
             HMT="#FDCDAC", DNMT="#D69C81", `Histone Demethylase`="#FFF2AE",
             HIF="#E3DCD5", JAK="#C8D7E2", PIM="#9AB8C4", `Aurora Kinase`="#F4CAE4",
             PARP="#CB546F", `Epigenetic Reader Domain`= "#476D61", `DNA-PK`="#E07A43", `Negative Control` = "gray50", MRN = "black"
             ))


ordered_marks2 <- c("Negative Control", "MRN", "DNA-PK","Epigenetic Reader Domain", "HDAC",  "Sirtuin", "HAT", 
                    "Histone Demethylase", "HMT", "DNMT", 
                    "HIF", "JAK", "PIM",
                    "Aurora Kinase", "PARP")

annotation_colors_scr$target <- annotation_colors_scr$target[ordered_marks2]

# 
# for (i in unique(eff.data.heatmap2$conc_rep)) {
#   pheatmap(as.matrix(eff.data.heatmap2[eff.data.heatmap2$conc_rep == i,] %>%
#                        column_to_rownames('drug') %>%
#                        dplyr::select(contains('IPR')) %>%
#                        t()),
#            border_color = "white",
#            cellwidth = 12,
#            cellheight = 12,
#            breaks = breaks,
#            cluster_rows = F,
#            main = paste('efficiency change, concentration =', i),
#            annotation_col = target,
#            color = colorRampPalette(rev(brewer.pal(n = 7, name = "RdBu")))(100),
#            annotation_colors = annotation_colors_scr, annotation_legend = T)
# }
# 

breaks <- seq(-0.1, 0.1, 0.001)

pheatmap(t(slope.prot.features.scale.dcast.100nm),
         border_color = "white",
         cellwidth = 12,
         cellheight = 12,
         breaks = breaks,
         annotation_col = target,
         main = "100 nM",
         color = colorRampPalette(rev(brewer.pal(n = 7, name = "RdBu")))(length(breaks)),
         annotation_colors = annotation_colors_scr, annotation_legend = T)

breaks <- seq(-0.1, 0.1, 0.001)

pheatmap(t(slope.prot.features.scale.dcast.1um),
         border_color = "white",
         cellwidth = 12,
         cellheight = 12,
         breaks = breaks,
         annotation_col = target,
         main = "1 uM",
         color = colorRampPalette(rev(brewer.pal(n = 7, name = "RdBu")))(length(breaks)),
         annotation_colors = annotation_colors_scr, annotation_legend = T)

breaks <- seq(-0.2, 0.2, 0.001)

pheatmap(t(slope.prot.features.scale.dcast.10um),
         border_color = "white",
         cellwidth = 12,
         cellheight = 12,
         breaks = breaks,
         annotation_col = target,
         main = "10 uM",
         color = colorRampPalette(rev(brewer.pal(n = 7, name = "RdBu")))(length(breaks)),
         annotation_colors = annotation_colors_scr, annotation_legend = T)

```





# Figure 4B: Matrix for hits
```{r heatmap}
#genes to select
tmp.hits <- slope.protein.features %>%
  # filter(p.value < 0.05 & term == "unlist(model.dt[j])") %>% 
  pull(drug_conc) %>%

# heatmap
ggplot(slope.protein.features.all %>% filter(gene %in% tmp.gene.hits & abs(slope) > 0.05)) + 
  geom_tile(aes(fct_relevel(gene,heatmap.gene.order.slope.diff.hits),fct_relevel(feature, heatmap.chromatin.order.slope.diff.hits), fill = slope)) +
  scale_fill_gradient2( low = "#8c510a",mid = "#f5f5f5", high = "#01665e")  + theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5), axis.title = element_blank(), legend.position = "top") + coord_fixed(expand = F,ratio = 1)

# Dendrogram
ggdendrogram(heatmap.slope.diff.hits$tree_row)

# Relative slope
tmp.rel.slope <- slope.protein.features.all %>% filter(gene %in% tmp.gene.hits) %>% mutate(rel.epistasis = slope*6/mean.fc)
```

# Figure 4C: Epistasis per library
```{r}
slope.library <- slope.protein.features.all %>% left_join(hits.both.screens %>% separate(ID_gene, into = "gene", remove = T) %>% select(gene,library,pathway) %>% distinct()) %>% filter(library %in% c("Chromatin","DNA_repair"))

# Plot per library(all)
ggplot(slope.library) + geom_quasirandom(aes(library,abs(epistasis), color = library)) + theme_bw() + scale_color_manual(values = levels(as_factor(library.colors$color))) + theme(legend.position = "top")

#test
chromatin.genes.slope <- slope.library %>% filter(library == "Chromatin") %>% pull(epistasis) %>% abs()
dna_repair.genes.slope <- slope.library %>% filter(library == "DNA_repair") %>% pull(epistasis) %>% abs()

ks.test.library <- ks.test(chromatin.genes.slope,dna_repair.genes.slope)
```

#Figure 4D: Epistasis per pathway
```{r}
# Plot per library 
ggplot(slope.library %>% filter(pathway != "NA")) + geom_quasirandom(aes(pathway, abs(epistasis), color = pathway)) + theme_bw() + scale_color_manual(values = c("#EB2030","#2E358F")) + theme(legend.position = "top")

#test
NHEJ.genes.slope <- slope.library %>% filter(pathway == "NHEJ") %>% pull(epistasis) %>% abs()
eff.genes.slope <- slope.library %>% filter(pathway == "eff") %>% pull(epistasis) %>% abs()

ks.test.pathway <- ks.test(NHEJ.genes.slope,eff.genes.slope)
```


