---
title: "FACS analysis"
author: "Max Trauernicht"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output: 
  html_document:
    theme: journal #cerulean
    highlight: monochrome
    toc: true
    toc_float: true
    code_folding: show
  editor_options:
    chunk_output_type: console
---

# knitr document van Steensel lab

# TRIP Clone identification

# Introduction
This script generates plots of FACS data. This way the effect of drugs on cell cycle will be visualized.
In this particular experiment, decitabine and DNA-PKi synergy was tested on several days after break induction (3,4,5,6 days).


## Description of Data
How to make a good rendering table: 
```{r table1, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
tabl <- "
| column1	|	column2	|	column3	|	
|----|----|----|
|1	|	2	|	3	|	
|a	|	b	|	c	|	
"
cat(tabl) # output the table in a format good for HTML/PDF/docx conversion
```

# Data processing
## Path, Libraries, Parameters and Useful Functions
```{r setup}
knitr::opts_chunk$set(echo = TRUE)
StartTime <-Sys.time()

# 8-digit Date tag:
Date <- substr(gsub("-","",Sys.time()),1,8) 
# libraries:
library(maditr)
library(ggplot2)
```

### Custom functions
Functions used thoughout this script.
```{r}
SetFileName <- function(filename, initials) {
  # Set filename with extension and initials to make filename with date integrated.
  filename <- substitute(filename)
  initials <- substitute(initials)
  filename <- paste0(initials, Date, filename)
  filename
}
```


## Data import
```{r}
setwd("/DATA/usr/m.trauernicht/projects/EpiScreen/FACS/Data/")
synergy.samples2 <- read.csv2("mt20190228_FlowJo_2.csv")
```

# Analysis

## The same for second df
```{r}
synergy.samples2 <- synergy.samples2[,-6]
synergy.samples2 <- setnames(synergy.samples2, "Drug", "drug")
synergy.samples2$day <- 3
synergy.samples2$rep <- 3
synergy.samples2$rep[c(13:18,25:30,49:54,61:66)] <- 4
synergy.samples2$rep[c(19:24,31:36,55:60,85:90)] <- 5
synergy.samples2$rep[c(67:72)] <- 6
synergy.samples2$rep[c(73:78)] <- 7
synergy.samples2$rep[c(79:84)] <- 8
synergy.samples2$rep[c(91:96)] <- 9
synergy.samples2$rep[c(97:102)] <- 10
synergy.samples2$rep[c(103:108)] <- 11
synergy.samples2$damage <- "10 Gy"
synergy.samples2$damage[c(109:114)] <- "GFP"
synergy.samples2$damage[c(115:120)] <- "LBR2"

# Remove not needed rows
synergy.samples2 <- synergy.samples2[synergy.samples2$Statistic != "",]

# Remove not needed column
synergy.samples2 <- synergy.samples2[,-1]

# Rename column
synergy.samples2$Name <- gsub("[A-Z][1-9].fcs/Cells(.*?)", "\\1", synergy.samples2$Name)
synergy.samples2$Name <- gsub("[A-Z][1-9][0-9].fcs/Cells(.*?)", "\\1", synergy.samples2$Name)
synergy.samples2$Name <- gsub("/Single Cells", "\\1", synergy.samples2$Name)
synergy.samples2$Name <- gsub("/", "\\1", synergy.samples2$Name)

# Remove not needed rows
synergy.samples2 <- synergy.samples2[synergy.samples2$Name != "",]
synergy.samples2 <- na.omit(synergy.samples2)

# Remove data with low cell numbers 
synergy.samples2$cell.count <- ave(synergy.samples2$X.Cells, synergy.samples2$drug, synergy.samples2$day, synergy.samples2$rep, FUN = function(x) sum(x))
synergy.samples2 <- synergy.samples2[synergy.samples2$cell.count > 1000,]
# synergy.samples2 <- synergy.samples2 [,c(-3,-8)]

## Rearrange to wide format
synergy.samples2 <- dcast(synergy.samples2, drug + day + rep + damage ~ Name, value.var = "Statistic", fun.aggregate = mean)


# Compute relative G2, G1 and S fraction
for (i in synergy.samples2$rep) {
synergy.samples2$sum.apo[synergy.samples2$rep == i] <- synergy.samples2$G1[synergy.samples2$rep == i] + 
  synergy.samples2$S[synergy.samples2$rep == i] + synergy.samples2$G2[synergy.samples2$rep == i] + 
  synergy.samples2$`Apoptotic cells`[synergy.samples2$rep == i]
}
for (i in synergy.samples2$rep) {
synergy.samples2$sum[synergy.samples2$rep == i] <- synergy.samples2$G1[synergy.samples2$rep == i] + 
  synergy.samples2$S[synergy.samples2$rep == i] + synergy.samples2$G2[synergy.samples2$rep == i] 
}

synergy.samples2$G2 <- synergy.samples2$G2 / synergy.samples2$sum
synergy.samples2$G1 <- synergy.samples2$G1 / synergy.samples2$sum
synergy.samples2$S <- synergy.samples2$S / synergy.samples2$sum
synergy.samples2 <- synergy.samples2[,c(-8:-9)]

# Merge the two dfs
synergy.samples2 <- melt(synergy.samples2, id.vars = c("drug", "day", "rep", "damage"))
synergy.samples2$drug <- as.character(synergy.samples2$drug)
synergy.samples2$drug[grep("Decitabine + NU7441", synergy.samples2$drug)] <- "NU7441 + Decitabine"

# Combine drug and day columns
# synergy.samples2$drug <- paste0(synergy.samples2$drug, "_", synergy.samples2$day)
# synergy.samples2 <- synergy.samples2[,-2]
```


## Visualisation
```{r out.width= "80%", fig.align= "center", echo = FALSE, warning = FALSE}
# Arrange everything in the right order
# synergy.samples2$conc <- factor(synergy.samples2$conc, levels=c("100nm", "1um", "10um"))
synergy.samples2$variable <- factor(synergy.samples2$variable, levels=c("G2", "S","G1"))
synergy.samples2$drug <- factor(synergy.samples2$drug, levels=c("DMSO", "NU7441", "Decitabine", "Decitabine + NU7441", "NU7441_GFP", "NU7441_LBR2"))

synergy.samples2 <- synergy.samples2[!duplicated(synergy.samples2),]
synergy.samples2 <- na.omit(synergy.samples2)

synergy.samples2$drug <- paste(synergy.samples2$drug, synergy.samples2$damage)

synergy.samples2$SD <- ave(synergy.samples2$value, synergy.samples2$drug, 
                                   synergy.samples2$variable, FUN = function(x) sd(x))
synergy.samples2$mean.value <- ave(synergy.samples2$value, synergy.samples2$drug, 
                                   synergy.samples2$variable, FUN = function(x) mean(x))
synergy.samples3 <- synergy.samples2
synergy.samples2 <- synergy.samples2[,c(-2:-4,-6)]
synergy.samples2 <- unique(synergy.samples2)


# Barplot of each drug next to each other
ggplot(data = synergy.samples2, aes(y = mean.value, x = drug, fill = variable)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1.0, vjust = 0)) +
  geom_bar(stat="identity") +
  geom_errorbar(aes(ymin = mean.value-SD, ymax = mean.value+SD, color = variable), width = 0.2) +
  scale_fill_manual(values = c("#DED89C","#BF7538","#749075"))
  labs(title = "cell cycle arrest upon blocking NHEJ (NU7441) and/or MMEJ (Decitabine)", 
       x = "days after damage induction", y = "G1, S, and G2 fractions") + 
  scale_fill_manual(values = c("#DED89C","#BF7538","#749075"))
  
  
# facet_grid(rows = vars(rep), cols = vars(drug))
# geom_errorbar(aes(ymin = mean.value-SD, ymax = mean.value+SD), data = synergy.samples2, width = 0.2)

  
```

# Statistics
```{r}
t.test.facs <- t.test(synergy.samples3$value[synergy.samples3$variable == "G2" & 
                                               synergy.samples3$drug == "DMSO 10 Gy"],
                      synergy.samples3$value[synergy.samples3$variable == "G2" & 
                                               synergy.samples3$drug == "Decitabine + NU7441 10 Gy"])

t.test.facs <- t.test(synergy.samples3$value[synergy.samples3$variable == "G2" & 
                                               synergy.samples3$drug == "DMSO 10 Gy"],
                      synergy.samples3$value[synergy.samples3$variable == "G2" & 
                                               synergy.samples3$drug == "NU7441 10 Gy"])
```

# Conclusion: We get a massive G2 arrest upon double drug treatment and dna damage induction


## Exporting potential data. 
```{r}
```

# Session Info
```{r}
paste("Run time: ",format(Sys.time()-StartTime))
getwd()
date()
sessionInfo()
```

