---
title: "FACS analysis"
author: "Max Trauernicht"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output: 
  html_document:
    theme: journal #cerulean
    highlight: monochrome
    toc: true
    toc_float: true
    code_folding: show
  editor_options:
    chunk_output_type: console
---

# knitr document van Steensel lab

# TRIP Clone identification

# Introduction
This script generates plots of FACS data. This way the effect of drugs on cell cycle will be visualized.
In this particular experiment, decitabine and DNA-PKi synergy was tested on several days after break induction (3,4,5,6 days).


## Description of Data
How to make a good rendering table: 
```{r table1, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
tabl <- "
| column1	|	column2	|	column3	|	
|----|----|----|
|1	|	2	|	3	|	
|a	|	b	|	c	|	
"
cat(tabl) # output the table in a format good for HTML/PDF/docx conversion
```

# Data processing
## Path, Libraries, Parameters and Useful Functions
```{r setup}
knitr::opts_chunk$set(echo = TRUE)
StartTime <-Sys.time()

# 8-digit Date tag:
Date <- substr(gsub("-","",Sys.time()),1,8) 
# libraries:
library(maditr)
library(ggplot2)
```

### Custom functions
Functions used thoughout this script.
```{r}
SetFileName <- function(filename, initials) {
  # Set filename with extension and initials to make filename with date integrated.
  filename <- substitute(filename)
  initials <- substitute(initials)
  filename <- paste0(initials, Date, filename)
  filename
}
```


## Data import
```{r}
setwd("/DATA/usr/m.trauernicht/projects/EpiScreen/FACS/Data/")
synergy.samples <- read.csv2("mt20190419_FlowJo_synergy.csv")
```

# Analysis

## Bring dataframe in right format
```{r out.width= "80%", fig.align= "center", echo = FALSE, warning = FALSE}
# Remove not needed rows
synergy.samples <- synergy.samples[synergy.samples$Statistic != "",]

# Remove not needed column
synergy.samples <- synergy.samples[,-1]

# Rename column
synergy.samples$Name <- gsub("[A-Z][1-9].fcs/Cells(.*?)", "\\1", synergy.samples$Name)
synergy.samples$Name <- gsub("[A-Z][1-9][0-9].fcs/Cells(.*?)", "\\1", synergy.samples$Name)
synergy.samples$Name <- gsub("/Single Cells", "\\1", synergy.samples$Name)
synergy.samples$Name <- gsub("/", "\\1", synergy.samples$Name)

# Remove not needed rows
synergy.samples <- synergy.samples[synergy.samples$Name != "",]

# Make apopototic cells df
synergy.samples$Statistic <- as.numeric(as.character(synergy.samples$Statistic))
apoptotic.samples <- synergy.samples[synergy.samples$Name == "Apoptotic cells",]
apoptotic.samples$drug <- paste0(apoptotic.samples$drug, "_", apoptotic.samples$day)
apoptotic.samples$drug <- paste0(apoptotic.samples$drug, "_", apoptotic.samples$damage)


# Look at which samples have highest apoptotic cell fractions
ggplot(data = apoptotic.samples[apoptotic.samples$Statistic >= 1,], aes(x = reorder(drug, -Statistic), y = Statistic)) +
  geom_bar(stat = "identity")+
   theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 10))+ 
  ylab("Apoptotic cell population (%)") + xlab("Drug and experiment day")+ labs(title = "Apoptotic cell population after drug treatment")


# Remove data with low cell numbers 
synergy.samples$cell.count <- ave(synergy.samples$X.Cells, synergy.samples$drug, synergy.samples$day, synergy.samples$rep, FUN = function(x) sum(x))
synergy.samples <- synergy.samples[synergy.samples$cell.count > 1000,]
synergy.samples <- synergy.samples [,c(-3,-8)]

## Rearrange to wide format
synergy.samples <- dcast(synergy.samples, drug + day + rep + damage ~ Name, value.var = "Statistic", fun.aggregate = mean)


# Compute relative G2, G1 and S fraction
for (i in synergy.samples$rep) {
synergy.samples$sum.apo[synergy.samples$rep == i] <- synergy.samples$G1[synergy.samples$rep == i] + 
  synergy.samples$S[synergy.samples$rep == i] + synergy.samples$G2[synergy.samples$rep == i] + 
  synergy.samples$`Apoptotic cells`[synergy.samples$rep == i]
}
for (i in synergy.samples$rep) {
synergy.samples$sum[synergy.samples$rep == i] <- synergy.samples$G1[synergy.samples$rep == i] + 
  synergy.samples$S[synergy.samples$rep == i] + synergy.samples$G2[synergy.samples$rep == i] 
}

synergy.samples$G2 <- synergy.samples$G2 / synergy.samples$sum
synergy.samples$G1 <- synergy.samples$G1 / synergy.samples$sum
synergy.samples$S <- synergy.samples$S / synergy.samples$sum
synergy.samples$`Apoptotic cells` <- synergy.samples$`Apoptotic cells` / synergy.samples$sum.apo
synergy.samples <- synergy.samples[,c(-9:-10)]

# Remove conditions with high apoptotic cell counts
synergy.samples <- synergy.samples[synergy.samples$`Apoptotic cells` < 0.10,]

# Merge the two dfs
synergy.samples <- melt(synergy.samples, id.vars = c("drug", "day", "rep", "damage"))

# Combine drug and day columns
# synergy.samples$drug <- paste0(synergy.samples$drug, "_", synergy.samples$day)
# synergy.samples <- synergy.samples[,-2]

```



## Visualisation
```{r out.width= "80%", fig.align= "center", echo = FALSE, warning = FALSE}
# Arrange everything in the right order
# synergy.samples$conc <- factor(synergy.samples$conc, levels=c("100nm", "1um", "10um"))
synergy.samples$variable <- factor(synergy.samples$variable, levels=c("Apoptotic cells", "G2", "S","G1"))
synergy.samples$drug <- factor(synergy.samples$drug, levels=c("DMSO", "NU7441", "Decitabine", "NU7441 + Decitabine"))

synergy.samples$SD <- ave(synergy.samples$value, synergy.samples$drug, 
                             synergy.samples$day, synergy.samples$variable, 
                             synergy.samples$damage,
                             FUN = function(x) sd(x))

synergy.samples$value <- ave(synergy.samples$value, synergy.samples$drug, 
                             synergy.samples$day, synergy.samples$variable, 
                             synergy.samples$damage,
                             FUN = function(x) mean(x))

synergy.samples <- synergy.samples[,-3]
synergy.samples <- synergy.samples[!duplicated(synergy.samples),]


synergy.samples <- synergy.samples[synergy.samples$day == 3, ]

# Barplot of each drug next to each other
ggplot(data = synergy.samples[synergy.samples$variable != "Apoptotic cells",], aes(y = value, x = day, fill = variable)) +
  geom_bar(stat="identity") + 
  geom_errorbar(aes(ymin = value-SD, ymax = value+SD, color = variable), width = 0.2) +
  labs(title = "cell cycle arrest upon blocking NHEJ (NU7441) and/or MMEJ (Decitabine)", 
       x = "days after damage induction", y = "G1, S, and G2 fractions") + 
  scale_fill_manual(values = c("#DED89C","#BF7538","#749075"))+
  facet_grid(rows = vars(damage), cols = vars(drug))


```


# Conclusion: We get a massive G2 arrest upon double drug treatment and dna damage induction

## Exporting potential data. 
```{r}
```

# Session Info
```{r}
paste("Run time: ",format(Sys.time()-StartTime))
getwd()
date()
sessionInfo()
```

