---
title: "FACS analysis"
author: "Max Trauernicht"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output: 
  html_document:
    theme: journal #cerulean
    highlight: monochrome
    toc: true
    toc_float: true
    code_folding: show
  editor_options:
    chunk_output_type: console
---

# knitr document van Steensel lab

# TRIP Clone identification

# Introduction
This script generates plots of FACS data. This way the effect of drugs on cell cycle will be visualized.
In this particular script the RSTP2#5 cells were treated with 5 candidate drugs from the epigenetic screening.


## Description of Data
How to make a good rendering table: 
```{r table1, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
tabl <- "
| column1	|	column2	|	column3	|	
|----|----|----|
|1	|	2	|	3	|	
|a	|	b	|	c	|	
"
cat(tabl) # output the table in a format good for HTML/PDF/docx conversion
```

# Data processing
## Path, Libraries, Parameters and Useful Functions
```{r setup}
knitr::opts_chunk$set(echo = TRUE)
StartTime <-Sys.time()

# 8-digit Date tag:
Date <- substr(gsub("-","",Sys.time()),1,8) 
# libraries:
library(maditr)
library(ggplot2)
```

### Custom functions
Functions used thoughout this script.
```{r}
SetFileName <- function(filename, initials) {
  # Set filename with extension and initials to make filename with date integrated.
  filename <- substitute(filename)
  initials <- substitute(initials)
  filename <- paste0(initials, Date, filename)
  filename
}
```


## Data import
```{r}
setwd("/DATA/usr/m.trauernicht/projects/EpiScreen/FACS/Data/")
gfp.samples <- read.csv2("mt20190402_GFP_FlowJo.csv")
lbr.samples <- read.csv2("mt20190402_LBR_FlowJo.csv")
```

# Analysis

## Bring dataframe in right format
```{r}
# Remove not needed column
gfp.samples <- gfp.samples[,c(-1,-4)]
lbr.samples <- lbr.samples[,c(-1,-4)]

# Add replicates
gfp.samples$rep <- "1"
gfp.samples[grep("[A-Z](5|7|9|11).*", gfp.samples$Name),]$rep <- "2"
gfp.samples[grep("[C](2).*", gfp.samples$Name),]$rep <- "2"
gfp.samples[grep("[D](2).*", gfp.samples$Name),]$rep <- "3"
lbr.samples$rep <- "1"
lbr.samples[grep("[A-Z](5|7|9|11).*", lbr.samples$Name),]$rep <- "2"
lbr.samples[grep("[C](2).*", lbr.samples$Name),]$rep <- "2"
lbr.samples[grep("[D](2).*", lbr.samples$Name),]$rep <- "3"

# Remove not needed rows
gfp.samples <- gfp.samples[gfp.samples$Statistic != "",]
lbr.samples <- lbr.samples[lbr.samples$Statistic != "",]

# Rename column
gfp.samples$Name <- gsub("[A-Z][1-9].fcs/Cells(.*?)", "\\1", gfp.samples$Name)
lbr.samples$Name <- gsub("[A-Z][1-9].fcs/Cells(.*?)", "\\1", lbr.samples$Name)
gfp.samples$Name <- gsub("[A-Z][1-9][0-9].fcs/Cells(.*?)", "\\1", gfp.samples$Name)
lbr.samples$Name <- gsub("[A-Z][1-9][0-9].fcs/Cells(.*?)", "\\1", lbr.samples$Name)
gfp.samples$Name <- gsub("/Single Cells", "\\1", gfp.samples$Name)
lbr.samples$Name <- gsub("/Single Cells", "\\1", lbr.samples$Name)
gfp.samples$Name <- gsub("/", "\\1", gfp.samples$Name)
lbr.samples$Name <- gsub("/", "\\1", lbr.samples$Name)


# Remove not needed rows
gfp.samples <- gfp.samples[gfp.samples$Name != "",]
lbr.samples <- lbr.samples[lbr.samples$Name != "",]

## Prepare data for visualization
gfp.samples$Statistic <- as.numeric(as.character(gfp.samples$Statistic))
gfp.samples <- dcast(gfp.samples, drug + conc + rep ~ Name, value.var = "Statistic",  fun.aggregate = mean)
lbr.samples$Statistic <- as.numeric(as.character(lbr.samples$Statistic))
lbr.samples <- dcast(lbr.samples, drug + conc + rep ~ Name , value.var = "Statistic", fun.aggregate = mean)

# Compute relative G2, G1 and S fraction
gfp.samples$sum <- gfp.samples$G1 + gfp.samples$S + gfp.samples$G2
gfp.samples$G2 <- gfp.samples$G2 / gfp.samples$sum
gfp.samples$G1 <- gfp.samples$G1 / gfp.samples$sum
gfp.samples$S <- gfp.samples$S / gfp.samples$sum
gfp.samples <- gfp.samples[-7]

lbr.samples$sum <- lbr.samples$G1 + lbr.samples$S + lbr.samples$G2
lbr.samples$G2 <- lbr.samples$G2 / lbr.samples$sum
lbr.samples$G1 <- lbr.samples$G1 / lbr.samples$sum
lbr.samples$S <- lbr.samples$S / lbr.samples$sum
lbr.samples <- lbr.samples[-7]

# Merge the two dfs
lbr.samples$drug <- paste0("LBR2_", lbr.samples$drug)
all.samples <- rbind(lbr.samples, gfp.samples)
all.samples <- na.omit(all.samples)
all.samples <- melt(all.samples, id.vars = c("drug", "conc", "rep"))

# Calculate mean and SD
all.samples2 <- all.samples
all.samples$SD <- ave(all.samples$value, all.samples$drug, all.samples$variable, all.samples$conc, FUN = function(x) sd(x))
all.samples$value <- ave(all.samples$value, all.samples$drug, all.samples$variable, all.samples$conc, FUN = function(x) mean(x))
all.samples <- all.samples[,-3]
all.samples[is.na(all.samples)] <- 0
```



## Visualisation
```{r}
# Arrange everything in the right order
all.samples <- all.samples[-grep("PAO", all.samples$drug),]
all.samples$conc <- factor(all.samples$conc, levels=c("0.1um", "0.75um", "1um", "1.25um", "5um", "10um", "15um", "20um"))
all.samples$variable <- factor(all.samples$variable, levels=c("G2", "S","G1"))
all.samples$drug <- factor(all.samples$drug, levels=c("DMSO", "LBR2_DMSO", "NU7441", "LBR2_NU7441", "Decitabine", "LBR2_Decitabine", "GSK-J4", "LBR2_GSK-J4", "AMG-900", "LBR2_AMG-900", "AZ-960", "LBR2_AZ-960"))

all.samples <- na.omit(all.samples)
all.samples <- unique(all.samples)

# Barplot of each drug next to each other
ggplot(data = all.samples, aes(y = value, x = conc, fill = variable)) +
  geom_bar(stat="identity") + 
  geom_errorbar(aes(ymin = value-SD, ymax = value+SD, color = variable),data = all.samples, width = 0.2) +
  labs(title = "cell cycle arrest upon treatment with drugs", 
       x = "cells", y = "G1, S, and G2 fractions") + 
  scale_fill_manual(values = c("#DED89C","#BF7538","#749075"))+
  facet_wrap(~drug, nrow = 4, ncol = 4, scales = "free_x")


```

# Statistical analysis
```{r}
# Perform t-test
t.test.samples <- t.test(all.samples2$value[all.samples2$variable == "G2" & 
                                               all.samples2$drug == "DMSO"],
                      all.samples2$value[all.samples2$variable == "G2" & 
                                           all.samples2$conc == "10um" &
                                               all.samples2$drug == 
                                           "LBR2_AMG-900"],
                      alternative = "less")
```

# Conclusion: Aurora kinases induce G2 arrest

## Exporting potential data. 
```{r}
setwd("/DATA/usr/m.trauernicht/projects/EpiScreen/FACS/Data/")
filename <- SetFileName("_five.drugs", "mt")
save(all.samples2, file = filename)
```

# Session Info
```{r}
paste("Run time: ",format(Sys.time()-StartTime))
getwd()
date()
sessionInfo()
```

