---
title: "FACS analysis"
author: "Max Trauernicht"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output: 
  html_document:
    theme: journal #cerulean
    highlight: monochrome
    toc: true
    toc_float: true
    code_folding: show
  editor_options:
    chunk_output_type: console
---

# knitr document van Steensel lab

# TRIP Clone identification

# Introduction
This script generates plots of FACS data. This way the effect of drugs on cell cycle will be visualized.


## Description of Data
How to make a good rendering table: 
```{r table1, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
tabl <- "
| column1	|	column2	|	column3	|	
|----|----|----|
|1	|	2	|	3	|	
|a	|	b	|	c	|	
"
cat(tabl) # output the table in a format good for HTML/PDF/docx conversion
```

# Data processing
## Path, Libraries, Parameters and Useful Functions
```{r setup}
knitr::opts_chunk$set(echo = TRUE)
StartTime <-Sys.time()

# 8-digit Date tag:
Date <- substr(gsub("-","",Sys.time()),1,8) 
# libraries:
library(maditr)
library(ggplot2)
```

### Custom functions
Functions used thoughout this script.
```{r}
SetFileName <- function(filename, initials) {
  # Set filename with extension and initials to make filename with date integrated.
  filename <- substitute(filename)
  initials <- substitute(initials)
  filename <- paste0(initials, Date, filename)
  filename
}
```


## Data import
```{r}
setwd("/DATA/usr/m.trauernicht/projects/EpiScreen/FACS/Data/")
gfp.samples <- read.csv2("mt20190415_GFP_FlowJo.csv")
lbr.samples <- read.csv2("mt20190415_LBR2_FlowJo.csv")
```

# Analysis

## Bring dataframe in right format
```{r}
# Remove not needed column
gfp.samples <- gfp.samples[,c(-1,-4)]
lbr.samples <- lbr.samples[,c(-1)]

# Remove not needed rows
gfp.samples <- gfp.samples[gfp.samples$Statistic != "",]
lbr.samples <- lbr.samples[lbr.samples$Statistic != "",]

# Rename column
gfp.samples$Name <- gsub("[A-Z][1-9].fcs/Cells(.*?)", "\\1", gfp.samples$Name)
lbr.samples$Name <- gsub("[A-Z][1-9].fcs/Cells(.*?)", "\\1", lbr.samples$Name)
gfp.samples$Name <- gsub("[A-Z][1-9][0-9].fcs/Cells(.*?)", "\\1", gfp.samples$Name)
lbr.samples$Name <- gsub("[A-Z][1-9][0-9].fcs/Cells(.*?)", "\\1", lbr.samples$Name)
gfp.samples$Name <- gsub("/Single Cells", "\\1", gfp.samples$Name)
lbr.samples$Name <- gsub("/Single Cells", "\\1", lbr.samples$Name)
gfp.samples$Name <- gsub("/", "\\1", gfp.samples$Name)
lbr.samples$Name <- gsub("/", "\\1", lbr.samples$Name)


# Remove not needed rows
gfp.samples <- gfp.samples[gfp.samples$Name != "",]
lbr.samples <- lbr.samples[lbr.samples$Name != "",]

## Prepare data for visualization
gfp.samples$Statistic <- as.numeric(as.character(gfp.samples$Statistic))
gfp.samples <- dcast(gfp.samples, drug + conc ~ Name, value.var = "Statistic", fun.aggregate = mean)
lbr.samples$Statistic <- as.numeric(as.character(lbr.samples$Statistic))
lbr.samples <- dcast(lbr.samples, drug + conc ~ Name, value.var = "Statistic", fun.aggregate = mean)

# Compute relative G2, G1 and S fraction
gfp.samples$sum <- gfp.samples$G1 + gfp.samples$S + gfp.samples$G2
gfp.samples$G2 <- gfp.samples$G2 / gfp.samples$sum
gfp.samples$G1 <- gfp.samples$G1 / gfp.samples$sum
gfp.samples$S <- gfp.samples$S / gfp.samples$sum
gfp.samples <- gfp.samples[-6]
gfp.samples$ID <- "GFP"

lbr.samples$sum <- lbr.samples$G1 + lbr.samples$S + lbr.samples$G2
lbr.samples$G2 <- lbr.samples$G2 / lbr.samples$sum
lbr.samples$G1 <- lbr.samples$G1 / lbr.samples$sum
lbr.samples$S <- lbr.samples$S / lbr.samples$sum
lbr.samples <- lbr.samples[-6]
lbr.samples$ID <- "LBR2"

# Merge the two dfs
all.samples <- rbind(lbr.samples, gfp.samples)
all.samples <- melt(all.samples, id.vars = c("drug", "conc", "ID"))

# Combine drug and conc columns
all.samples$drug <- paste0(all.samples$drug, "_", all.samples$conc)
all.samples <- all.samples[,-2]

```



## Visualisation
```{r}
# Arrange everything in the right order
all.samples <- all.samples[-grep("PAO", all.samples$drug),]
# all.samples$conc <- factor(all.samples$conc, levels=c("100nm", "1um", "10um"))
all.samples$variable <- factor(all.samples$variable, levels=c("G2", "S","G1"))
# all.samples$drug <- factor(all.samples$drug, levels=c("DMSO", "LBR2_DMSO","Alisertib", "LBR2_Alisertib","AMG-900", "LBR2_AMG-900","Apicidin", "AR-42", "LBR2_AR-42", "Aurora A Inhibitor I", "LBR2_Aurora A Inhibitor I", "Azacitidine","LBR2_Azacitidine", "Barasertib", "LBR2_Barasertib", "BML-210", "LBR2_BML-210", "C646", "LBR2_C646", "CI994", "LBR2_CI994", "CTPB", "LBR2_CTPB", "CX-6258", "LBR2_CX-6258", "CYC116", "LBR2_CYC116", "CYT387", "LBR2_CYT387", "Danusertib", "LBR2_Danusertib", "DZNeP", "LBR2_DZNeP", "Entinostat", "LBR2_Entinostat", "Givinostat", "LBR2_Givinostat", "I-BET-762", "LBR2_I-BET-762", "LY2784544", "LBR2_LY2784544", "MK-5108","LBR2_MK-5108", "NSC-3852", "LBR2_NSC-3852", "OTX015", "LBR2_OTX015", "PCI-24781", "LBR2_PCI-24781", "PJ34 HCl", "LBR2_PJ34 HCl", "Pracinostat", "LBR2_Pracinostat", "Quercetin", "LBR2_Quercetin", "Resveratrol", "LBR2_Resveratrol", "RG2833", "LBR2_RG2833", "RGFP966", "LBR2_RGFP966", "RVX-208", "LBR2_RVX-208", "Scriptaid", "LBR2_Scriptaid", "SGC 0946", "LBR2_SGC 0946", "SGC-CBP30", "LBR2_SGC-CBP30", "SNS-314 Mesylate", "LBR2_SNS-314 Mesylate", "Suramin", "LBR2_Suramin", "TG101209", "LBR2_TG101209", "Tubastatin A HCl", "LBR2_Tubastatin A HCl", "Vorinostat", "LBR2_Vorinostat"))
all.samples <- na.omit(all.samples)

# Barplot of each drug next to each other
ggplot() +
  geom_bar(data = all.samples, aes(y = value, x = ID, fill = variable), stat="identity") + 
  labs(title = "cell cycle arrest upon PD treatment", 
       x = "cells", y = "G1, S, and G2 fractions") + scale_fill_manual(values = c("#DED89C","#BF7538","#749075"))+
  facet_wrap(~drug, nrow = 10, ncol = 10, scales = "free_x")


```

# Results
```{r}
```

# Conclusions
```{r}
```

## Exporting potential data. 
```{r}
```

# Session Info
```{r}
paste("Run time: ",format(Sys.time()-StartTime))
getwd()
date()
sessionInfo()
```

