---
title: "Clone 5 annotation"
author: "Max Trauernicht"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  pdf_document:
    toc: yes
  html_document:
    code_folding: show
    highlight: monochrome
    theme: journal
    toc: yes
    toc_float: yes
  editor_options:
    chunk_output_type: console
---

# knitr document van Steensel lab

# TRIP Clone annotation

# Introduction
In this document, ChIP data from the location of the barcodes in RSTP2#5 will be imported in order to be able to characterize and annotate each barcode. In this way, we can identify which barcodes are in a rather active chromatin environment or barcodes that are in repressed environments. This will help us to understand the results we might get in future experiments (e.g. the epigenetic screening).

## Description of Data
How to make a good rendering table: 
```{r table1, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
tabl <- "
| column1	|	column2	|	column3	|	
|----|----|----|
|1	|	2	|	3	|	
|a	|	b	|	c	|	
"
cat(tabl) # output the table in a format good for HTML/PDF/docx conversion
```

# Data processing
## Path, Libraries, Parameters and Useful Functions
```{r setup}
knitr::opts_chunk$set(echo = TRUE)
StartTime <-Sys.time()

# 8-digit Date tag:
Date <- substr(gsub("-","",Sys.time()),1,8) 
# libraries:
library(dplyr)
library(data.table)
library(devtools)
library(RColorBrewer)
library(pheatmap)
library(rtracklayer)
library(GenomicRanges)
library(caret)
library(tidyverse)
library(ggbio)
```

### Custom functions
Functions used thoughout this script.
```{r}
SetFileName <- function(filename, initials) {
  # Set filename with extension and initials to make filename with date integrated.
  filename <- substitute(filename)
  initials <- substitute(initials)
  filename <- paste0(initials, Date, filename)
  filename
}
```


## Data import
```{r out.width= "80%", fig.align= "center", echo = FALSE, warning = FALSE}
# Import dam files
dam_files = list.files('/DATA/projects/DSBrepair/data/tracks_20190429/dam/',
                       full.names=T)

dam_meta.dt = data.table(pool=gsub('.*/(.*)-[0-9]+_.*', '\\1', dam_files),
                         binsize=as.numeric(gsub('.*-([0-9]+)_.*', '\\1', dam_files)),
                         dam_name=gsub('.*-[0-9]+_(.*).txt', '\\1', dam_files),
                         filename=dam_files)

damid.dt = dam_meta.dt[, fread(filename, stringsAsFactors=F,
                               col.names=c('ID', 'mean', 'gmean'))[,c('ID', 'mean')],
                       by=filename]

damid.dt = merge(damid.dt, dam_meta.dt, by='filename')

# Import coverage files (including WGBS, repliseq & DNAse data)
cov_files = list.files('/DATA/projects/DSBrepair/data/tracks_20190429/coverage',
                       full.names=T, patter='.tsv')

cov_meta.dt = data.table(pool=gsub('.*/(.*)-[0-9]+_.*', '\\1', cov_files),
                         binsize=as.numeric(gsub('.*-([0-9]+)_.*', '\\1', cov_files)),
                         filename=cov_files)

cov_data.dt = cov_meta.dt[,fread(filename, stringsAsFactors=F),by=c('pool', 'binsize')]


# Import ChIP files
chip_files = list.files('/DATA/projects/DSBrepair/data/tracks_20190429/chip/site_means/',
                        full.names=T)

chip_meta.dt = data.table(pool=gsub('.*/(.*)-[0-9]+_.*', '\\1', chip_files),
                          binsize=as.numeric(gsub('.*-([0-9]+)_.*', '\\1', chip_files)),
                          chip_name=gsub('.*-[0-9]+_(.*).txt', '\\1', chip_files),
                          filename=chip_files)

chip_meta.dt[,chip_source:=gsub('(.*)_.*_.*', '\\1', chip_name)]
chip_meta.dt[,chip_target:=gsub('.*_(.*)_.*', '\\1', chip_name)]
chip_meta.dt[,chip_id:=gsub('.*_.*_(.*)', '\\1', chip_name)]


chip.dt = chip_meta.dt[, fread(filename, stringsAsFactors=F)[,c('ID', 'Input', 'log2_mean')],
                  by=filename]

chip.dt = merge(chip.dt, chip_meta.dt, by='filename')




# Import damonly files
damonly_files = list.files('/DATA/projects/DSBrepair/data/features/',
                           full.names=T, pattern='pool_r[0-9]_Dam')



meta_damonly.dt = data.table(dam_name=gsub('.*_(r[0-9]+_.*)-.*', '\\1',
                                           damonly_files),
                             filename=damonly_files)

meta_damonly.dt[,rep:=gsub('(.*?)_.*', '\\1', dam_name)]
meta_damonly.dt[,dam_id:=gsub('.*?_(.*)', '\\1', dam_name)]
meta_damonly.dt[,dam_id:=gsub('_.*', '', dam_id)]


damonly.dt = meta_damonly.dt[, fread(filename, stringsAsFactors=F,
                                     col.names=c('ID', 'count')),
                             by=c('dam_name', 'rep', 'dam_id', 'filename')]
damonly.dt[,count:=ifelse(is.na(count),0,count)]

damonly_mean = damonly.dt[,list(sum=sum(count) + 1), by=ID]
setkey(damonly_mean, 'ID')


```

# Analysis

## Pre-processing of data
```{r out.width= "80%", fig.align= "center", echo = FALSE, warning = FALSE}
# Calculate pseudo count for DNAse data to calculate the mean of the data
dnase_pseudo = cov_data.dt[DNAse_SE>0, min(DNAse_SE)] / 2
cov_data.dt[,DNAse:=(log2(DNAse_PE) + log2(DNAse_SE + dnase_pseudo))/2]

# Calculate mean of WGBS and repliseq data
cov_data.dt[,WGBS:=(log2(WGBS_r1) + log2(WGBS_r2))/2]
cov_data.dt[,repliseq:=(repliseq_r1 + repliseq_r2)/2]

# Select only reliable ChiP datasets
meta_selection.dt = chip_meta.dt[chip_source%in%c('Pradeepa2016', 'Minoda2018',
                                             'Schmidl2015', 'Bernstein2012',
                                             'Snyder2016', 'Bernstein2011',
                                             'Stamatoyannopoulos2011',
                                             'Ratnakumar2012', 'Shah2018',
                                             'Tyler2017', 'Snyder2017',
                                             'Salzberg2017', 'Snyder2012',
                                             'Snyder2018')]
meta_selection.dt = meta_selection.dt[!chip_id%in%c('ENCSR031TFS',
                                                    'ENCSR000APE'),]


chip.dt = meta_selection.dt[,fread(filename, stringsAsFactors=F)[,c('ID', 'Input', 'log2_mean')],
                            by=filename]

chip.dt = merge(chip.dt, chip_meta.dt, by='filename')


# Bring chip dataframe and damid dataframe in right format
mean_chip.dt = chip.dt[,list(log2_mean=mean(log2_mean)),
                       by=c('binsize', 'ID', 'pool', 'chip_target') ]

chip_cast.dt = data.table::dcast(mean_chip.dt, pool + binsize + ID ~ chip_target,
                                 value.var='log2_mean')

damid_cast.dt = data.table::dcast(damid.dt, pool + binsize + ID ~ dam_name,
                                  value.var='mean')

features.dt = merge(chip_cast.dt, damid_cast.dt, by=c('ID', 'binsize', 'pool'))
features.dt = merge(features.dt, cov_data.dt[,c('barcode', 'binsize', 'pool',
                                                'DNAse', 'WGBS', 'repliseq')],
                    by.y=c('barcode','binsize', 'pool'),
                    by.x=c('ID', 'binsize', 'pool'))
  
features.dt[,c("ARID3A", "BRD4", "ELK1","GTF2F1","H3K122ac","H4K5acK8ac",
                         "LARP7","MAX","MAZ","MLLT1","MYC","POL2AS2","PU1","RCOR1",
                         "SMC3","Znf143","4xAP3","EZH2","H2AFZ","macroH2A", "CENPB")] <- list(NULL)

features.dt.long <- melt(features.dt[,c(1,4:20)])


pool_2000 = features.dt[binsize==2000 & !is.na(repliseq) &
                          pool=='20180918.RSTP2', -c('binsize', 'pool')]
clone5_2000 = features.dt[binsize==2000 & !is.na(repliseq) &
                            pool=='clone5', -c('binsize', 'pool')]
```


## Normalization mean values
```{r out.width= "80%", fig.align= "center", echo = FALSE, warning = FALSE}
# Calculate SD and mean of the log values of each ChIP dataset on the complete pool
p = preProcess(pool_2000, method=c('center', 'scale'))

# Now predict the means and SD to calculate the z-scores on the clone5 data only
normalized_clone5 = predict(p, clone5_2000)
```

## Create code for heatmaps: mapping_mean dataset
```{r out.width= "80%", fig.align= "center", echo = FALSE, warning = FALSE}
# Annotate chr. marks: here I create a df to annotate which chromatin marks are active/repressive, 
# this is being used in the heatmap later

# Add type of chr. mark (repressive or active)
features.dt.long$type <- "Active"
features.dt.long[grep("K9me|K27me|LMNB1|repliseq", 
                       features.dt.long$variable),]$type <- "Repressive"

chromatin.mark_type1 <- features.dt.long[,c(-1,-3)]
chromatin.mark_type1 <- chromatin.mark_type1[!duplicated(chromatin.mark_type1), ]
chromatin.mark_type <- chromatin.mark_type1[,2]
chromatin.mark_type1 <- chromatin.mark_type1[,-2]
chromatin.mark_type1 <- as.data.frame(chromatin.mark_type1)
chromatin.mark_type <- as.data.frame(chromatin.mark_type)
row.names(chromatin.mark_type) <- chromatin.mark_type1$variable
setnames(chromatin.mark_type, "type", "Type of Chromatin Mark")

# Annotate barcodes
barcode_annotation <- normalized_clone5[,1]
barcode_annotation <- barcode_annotation %>% 
  remove_rownames %>% column_to_rownames(var="ID")
barcode_annotation$ID[c(20,3,11)] <- "Active Promoter/Enhancer"
barcode_annotation$ID[c(14,9)] <- "H3K9me2/3 Domain"
barcode_annotation$ID[c(4,16,5,19)] <- "Active Gene Body"
barcode_annotation$ID[c(6,15,7,17)] <- "LAD"
barcode_annotation$ID[c(8,18)] <- "No Marks Present"
barcode_annotation$ID[c(2,12,1,10,13)] <- "H3K27me3 Domain"
setnames(barcode_annotation, "ID", "Integration Site")
 

# Change colors of annotations
colors <- brewer.pal(length(unique(barcode_annotation$`Integration Site`)), "Pastel2")
names(colors) <- unique(barcode_annotation$`Integration Site`)

annotation_colors = list(
  `Type of Chromatin Mark` = c(Active="#6AB163", Repressive="grey90"),
  `Integration Site` = colors)

# Order the marks for annotation
ordered_marks <- c("Active Promoter/Enhancer","H3K9me2/3 Domain", "Active Gene Body", "LAD", "No Marks Present", "H3K27me3 Domain")
annotation_colors$`Integration Site` <- annotation_colors$`Integration Site`[ordered_marks]

# Set ChIP column as rownames and remove ChIP column
normalized_clone5 <- normalized_clone5 %>% 
  remove_rownames %>% column_to_rownames(var="ID")

# Transpose df
normalized_clone5 <- t(normalized_clone5)
```


Next, heatmaps can be visualized, using either the dataframe containing only essential ChIP data or all ChIP data.

\newpage
# Results
```{r out.width= "80%", fig.align= "center", echo = FALSE, warning = FALSE}
## Generate the heatmaps

# The following two heatmaps represent in color the intensity -
# of the chromatin marks at the barcodes integrated in RSTP2 clone 5

# Each chromatin mark is annotated with either repressive or active

# Additionally, LAD & H3K27me3 integration calling states are represented -
# per barcode (meaning: do these barcodes reside in a LAD/H3K27me3 domain?)

breaksList <- c(seq(min(as.matrix(normalized_clone5)), 0, 
                   length.out=ceiling(100/2) + 1), 
              seq(max(as.matrix(normalized_clone5))/100, 
                  max(as.matrix(normalized_clone5)), length.out=floor(100/2)))
breaksSaturated <- seq(-1.5,1.5,0.03)

# Generate heatmap: chr. mark intensity of ChIP datasets vs. barcodes
pheatmap(as.matrix(normalized_clone5[,c(-1,-11)]),
         annotation_row = chromatin.mark_type, 
         annotation_col = barcode_annotation,
         fontsize_row = 12, annotation_names_row = F, 
         fontsize_col = 12, annotation_names_col = F,
         annotation_colors = annotation_colors,
         color = colorRampPalette(rev(brewer.pal(n = 7, name = "RdBu")))(100),
         breaks = breaksSaturated,
         main = "RSTP2#5: chromatin mark intensity per barcode")

# Only essential ChIP datasets (barcode dendogram is extracted from this heatmap for further use)
x <- pheatmap(as.matrix(normalized_clone5[c(-11:-12,-15,-17),c(-1,-11)]),
         annotation_row = chromatin.mark_type, 
         annotation_col = barcode_annotation,
         fontsize = 12,
         fontsize_row = 12, annotation_names_row = F, 
         fontsize_col = 12, annotation_names_col = F,
         annotation_colors = annotation_colors,
         color = colorRampPalette(rev(brewer.pal(n = 7, name = "RdBu")))(100),
         breaks = breaksSaturated,
         labels_col = c("Integration 15","Integration 1","Integration 5","Integration 6",
                        "Integration 10","Integration 9","Integration 13","Integration 3",
                        "Integration 17","Integration 16","Integration 18", "Integration 4",
                        "Integration 11","Integration 7","Integration 12", "Integration 14",
                        "Integration 8","Integration 2"),
         main = "RSTP2#5: chromatin mark intensity per barcode")
```

## Conclusions from this heatmap:
- LAD barcodes cluster together, but still have some intra-group variability
- There seems to be a high variability between barcodes (we have pure LADs (e.g. CATCC...), repressed barcodes (e.g. GAGC...), rather
boring barcodes (no marks, e.g. GCGCA...), and very active barcodes(e.g. TATG...))
- active marks tend to cluster together


# Conclusion
The heatmaps reveal many valuable information about the integrations of clone 5. 
We can now use this information to look at how individual barcodes responded in the drug screening.
We can also try to cluster barcodes based on the above heatmaps and use these clusters in the analysis of the epigenetic drug screening.



## Exporting potential data. 
```{r}
setwd("/DATA/usr/m.trauernicht/projects/EpiScreen/files_scripts/")
filename <- SetFileName("_mapping_mean", "mt")
save(normalized_clone5, file = filename)


filename <- SetFileName("_mapping_mean_long", "mt")
save(features.dt.long, file = filename)

filename <- SetFileName("_chip_dendogram", "mt")
save(x, file = filename)


setwd("/DATA/usr/m.trauernicht/projects/EpiScreen/files_scripts/")
filename <- SetFileName("_barcode_annotation.csv", "mt")
write.csv2(barcode_annotation, file = filename)
```

# Session Info
```{r}
paste("Run time: ",format(Sys.time()-StartTime))
getwd()
date()
sessionInfo()
```

