---
title: <font color="#ff8000">**GCF-7637 - Differential H3K27ac - K562 treated with HDAC inhibitors**</font>
author: <font color="gray">Sebastian Gregoricchio</font>
date: <font color="gray">`r format(Sys.time(), '%d %B, %Y')`</font>
output:
  html_document:
    toc: yes
    toc_float: yes
    number_sections: true
    df_print: paged
    theme: "united"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, rows.print=10, warning = F, message = F)
```

------------------------------

**Environment details**
```{r global_vars, message=FALSE, warning=FALSE}
# Working dir
wd = "/DATA/s.gregoricchio/GCF-7637_K562_ChIP/"
wd = gsub("/$","",wd)
setwd(wd)

diffBind_dir = "06_differential_analyses/H3K27ac"
diffBind_dir = gsub("/$","",diffBind_dir)
dir.create(path = paste0(wd,"/",diffBind_dir), showWarnings = F, recursive = T)

prefix_analyses = "diffbind_K562_H3K27ac"


## Analyses setup
constrast_list = list("PCI.vs.DMSO" = c("Condition", "PCI", "DMSO"),
                      "Tuba.vs.DMSO" = c("Condition", "Tuba", "DMSO"))

FDR_threshold = 0.05
FC_threshold = 2 #linear value
promoter_range = c(-2000, 1000)


# Attach packages
require(dplyr)
require(DiffBind)
require(ggplot2)


# Color palettes
conditions = c("DMSO", "PCI", "Tuba")
cell_colors = rainbow(length(conditions))
names(cell_colors) = conditions

rep_colors = colorRampPalette(c("#c4e6e6", "#1a6875"))(3)
names(rep_colors) = as.character(1:3)


n.colors = 101

Spectral = colorRampPalette(RColorBrewer::brewer.pal(n = 11, name = "Spectral"))(n.colors)
RdYlGn = colorRampPalette(RColorBrewer::brewer.pal(n = 11, name = "RdYlGn"))(n.colors)
RdYlBu = colorRampPalette(RColorBrewer::brewer.pal(n = 11, name = "RdYlBu"))(n.colors)
RdGy = colorRampPalette(RColorBrewer::brewer.pal(n = 11, name = "RdGy"))(n.colors)
RdBu = colorRampPalette(RColorBrewer::brewer.pal(n = 11, name = "RdBu"))(n.colors)
PuOr = colorRampPalette(RColorBrewer::brewer.pal(n = 11, name = "PuOr"))(n.colors)
PRGn = colorRampPalette(RColorBrewer::brewer.pal(n = 11, name = "PRGn"))(n.colors)
PiYG = colorRampPalette(RColorBrewer::brewer.pal(n = 11, name = "PiYG"))(n.colors)
BrBG = colorRampPalette(RColorBrewer::brewer.pal(n = 11, name = "BrBG"))(n.colors)

YlOrRd = colorRampPalette(RColorBrewer::brewer.pal(n = 9, name = "YlOrRd"))(n.colors)
YlOrBr = colorRampPalette(RColorBrewer::brewer.pal(n = 9, name = "YlOrBr"))(n.colors)
YlGnBu = colorRampPalette(RColorBrewer::brewer.pal(n = 9, name = "YlGnBu"))(n.colors)
YlGn = colorRampPalette(RColorBrewer::brewer.pal(n = 9, name = "YlGn"))(n.colors)
Reds = colorRampPalette(RColorBrewer::brewer.pal(n = 9, name = "Reds"))(n.colors)
RdPu = colorRampPalette(RColorBrewer::brewer.pal(n = 9, name = "RdPu"))(n.colors)
Purples = colorRampPalette(RColorBrewer::brewer.pal(n = 9, name = "Purples"))(n.colors)
PuRd = colorRampPalette(RColorBrewer::brewer.pal(n = 9, name = "PuRd"))(n.colors)
PuBuGn = colorRampPalette(RColorBrewer::brewer.pal(n = 9, name = "PuBuGn"))(n.colors)
PuBu = colorRampPalette(RColorBrewer::brewer.pal(n = 9, name = "PuBu"))(n.colors)
OrRd = colorRampPalette(RColorBrewer::brewer.pal(n = 9, name = "OrRd"))(n.colors)
Oranges = colorRampPalette(RColorBrewer::brewer.pal(n = 9, name = "Oranges"))(n.colors)
Greys = colorRampPalette(RColorBrewer::brewer.pal(n = 9, name = "Greys"))(n.colors)
Greens = colorRampPalette(RColorBrewer::brewer.pal(n = 9, name = "Greens"))(n.colors)
GnBu = colorRampPalette(RColorBrewer::brewer.pal(n = 9, name = "GnBu"))(n.colors)
BuPu = colorRampPalette(RColorBrewer::brewer.pal(n = 9, name = "BuPu"))(n.colors)
BuGn = colorRampPalette(RColorBrewer::brewer.pal(n = 9, name = "BuGn"))(n.colors)
Blues = colorRampPalette(RColorBrewer::brewer.pal(n = 9, name = "Blues"))(n.colors)
```

--------------------------------------

<br>

# <font color="#074fa5">**DBA object set-up**</font>
```{r sample_tb_config, warning=FALSE}
## Read sample table configuration
sample_tb =
  read.delim(paste0(wd, "/06_differential_analyses/diffBind_sample_sheet.txt"), h=T, sep="\t") %>%
  dplyr::filter(Factor == "H3K27ac")

sample_tb



## Generate DBA object
dba_setup = dba(sampleSheet = sample_tb,
                config = data.frame(AnalysisMethod = DBA_DESEQ2,
                                    th = FDR_threshold,
                                    design = TRUE,
                                    cores = 30,
                                    RunParallel = TRUE,
                                    doBlacklist = TRUE,
                                    doGreylist = FALSE))
```
```
K562_H3K27ac_DMSO_rep1 K562 H3K27ac DMSO  1 narrow
K562_H3K27ac_DMSO_rep2 K562 H3K27ac DMSO  2 narrow
K562_H3K27ac_DMSO_rep3 K562 H3K27ac DMSO  3 narrow
K562_H3K27ac_PCI_rep1 K562 H3K27ac PCI  1 narrow
K562_H3K27ac_PCI_rep2 K562 H3K27ac PCI  2 narrow
K562_H3K27ac_PCI_rep3 K562 H3K27ac PCI  3 narrow
K562_H3K27ac_Tuba_rep1 K562 H3K27ac Tuba  1 narrow
K562_H3K27ac_Tuba_rep2 K562 H3K27ac Tuba  2 narrow
K562_H3K27ac_Tuba_rep3 K562 H3K27ac Tuba  3 narrow
```

```{r warning=FALSE}
dba_setup
```

<br>



## Apply Blacklist
Pay attention that the blacklist cmust be compatibale with tour data (e.g., genome building and, presence or not of chr in the bams). <br> Typical problem here is that it removes almost everythng from your data. This would mean that the balck list is not the good one.
```{r deblacklisting, message=FALSE, warning=FALSE}
deblacklisted_dba = dba.blacklist(dba_setup, blacklist = DBA_BLACKLIST_GRCH38)
```
```
Genome detected: Hsapiens.NCBI.GRCh38
Applying blacklist...
Removed: 179 of 399201 intervals.
Removed: 18 merged (of 71271) and 11 (of 46448) consensus.
```
```{r show_deblacklisted, warning=FALSE}
deblacklisted_dba
```

<br>



# <font color="#074fa5">**Global occupancy analysis**</font>
## Global correlation at peaks
```{r global_corr_plot, warning=FALSE, fig.height=5, fig.width=5, fig.align='center'}
dba.plotHeatmap(deblacklisted_dba,
                attributes = DBA_CONDITION,
                correlations = T,
                colScheme = GnBu,
                ColAttributes = DBA_CONDITION,
                colSideCols = cell_colors,
                RowAttributes = DBA_REPLICATE,
                rowSideCols = rep_colors,
                minval = 0,
                maxval = 1)
```
```{r export_global_corr_plot, include=FALSE}  
dir.create(paste0(wd, "/",diffBind_dir, "/plots/global_analyses/"), showWarnings = F, recursive = T)
pdf(file = paste0(wd, "/",diffBind_dir, "/plots/global_analyses/heatmap_global_correlation.pdf"),
    width = 7, height = 7.5)
dba.plotHeatmap(deblacklisted_dba,
                attributes = DBA_CONDITION,
                correlations = T,
                colScheme = GnBu,
                ColAttributes = DBA_CONDITION,
                colSideCols = cell_colors,
                RowAttributes = DBA_REPLICATE,
                rowSideCols = rep_colors,
                minval = 0,
                maxval = 1)
dev.off()

### export tables
dir.create(paste0(wd, "/",diffBind_dir, "/tables_and_counts/"), showWarnings = F, recursive = T)
correlations = dba.plotHeatmap(deblacklisted_dba, correlations = T, olPlot = DBA_COR)
write.table(x = data.frame(correlations),
            file = paste0(wd, "/",diffBind_dir, "/tables_and_counts/heatmap_global_correlation.txt"),
            sep = "\t", col.names = T, row.names = T, quote = F)
#perc_overlap = dba.plotHeatmap(deblacklisted_dba, correlations = T, olPlot = DBA_OLAP)
#num_common_peaks = dba.plotHeatmap(deblacklisted_dba, correlations = T, olPlot = DBA_INALL)
```

<br>

## Global PCA at peaks
```{r global_PCA, warning=FALSE, fig.height=5, fig.width=5, fig.align='center'}
for (i in list(c(1,2), c(2,3), c(3,4), c(4,5))) {
  dba.plotPCA(deblacklisted_dba,
              attributes = DBA_CONDITION,
              vColors = cell_colors,
              label = DBA_REPLICATE,
              components = i)
}
```
```{r export_global_PCA, include=FALSE}
for (i in list(c(1,2), c(2,3), c(3,4), c(4,5), c(5,6), c(6,7))) {
  dir.create(paste0(wd, "/",diffBind_dir, "/plots/global_analyses/"), showWarnings = F, recursive = T)
  pdf(file = paste0(wd, "/",diffBind_dir, "/plots/global_analyses/global_PCA.",i[1],"-", i[2],".pdf"),
      width = 7, height = 7.5)
  dba.plotPCA(deblacklisted_dba,
              attributes = DBA_CONDITION,
              vColors = cell_colors,
              label = DBA_REPLICATE,
              components = i)
  dev.off()
}
```


<br>

# <font color="#074fa5">**Differential analysis**</font>
## Count reads in peaks *(it takes a while!)*
Usually I compute the counts with the following code and I save it in a Rdata file.
```{r compute_counts, warning=FALSE, eval=F}
counts = dba.count(deblacklisted_dba, bParallel = TRUE)

dir.create(paste0(wd, "/",diffBind_dir, "/tables_and_counts/"), showWarnings = F, recursive = F)

dba.save(counts,
         file = prefix_analyses,
         dir = paste0(wd, "/",diffBind_dir, "/tables_and_counts/"),
         pre = "dba_count_",
         ext = "RData")
```
```
## Computing summits...
## Re-centering peaks...
## Reads will be counted as Paired-end.
```

So then I can reload it if I need it
```{r read_counts, warning=FALSE}
counts = dba.load(file = prefix_analyses,
                  dir = paste0(wd, "/",diffBind_dir, "/tables_and_counts/"),
                  pre = "dba_count_",
                  ext = "RData")

counts
```


<font color="red">**Please note**</font> that `dba.count` default filters peaks with CPM (Count Per Million) < 1.

<br>

## Data normalization
Default method for normalization uses library sizes
```{r normalize_data, warning=FALSE}
## get library sizes
info = dba.show(counts)
lib.sizes = cbind(LibReads = info$Reads,
                  FRiP = info$FRiP,
                  PeakReads = round(info$Reads * info$FRiP))

rownames(lib.sizes) = info$ID
lib.sizes


## Normalize the data
norm_dba = dba.normalize(counts)
```

<br>

## Evaluate contrast
```{r evaluate_contrast, eval=FALSE}
constrast_results_list = list()
constrast_norms_used_list = list()

for (i in 1:length(constrast_list)) {
  cat(paste0("-----------------------------------------------------------------------------\n",
             names(constrast_list)[i],
             "\n-----------------------------------------------------------------------------\n"))
  
  constrast_results_list[[i]] =
    dba.contrast(norm_dba,
                 contrast = constrast_list[[i]]) 
  print(constrast_results_list[[i]])
  
  ## Get normalization methods used for your own information
  constrast_norms_used_list[[i]] = dba.normalize(constrast_results_list[[i]], bRetrieve = TRUE)
  print(constrast_norms_used_list[[i]])
}

names(constrast_results_list) = names(constrast_list)
names(constrast_norms_used_list) = names(constrast_list)


## Save analyses:
saveRDS(constrast_results_list,
        file=paste0(wd,"/",diffBind_dir, "/tables_and_counts/",prefix_analyses,"_contrast.analyses.list.RData"))
```

```{r contrast_import, echo=FALSE}
# import contrast_results_list
constrast_results_list = 
  readRDS(paste0(wd,"/",diffBind_dir, "/tables_and_counts/",prefix_analyses,"_contrast.analyses.list.RData"))
constrast_norms_used_list = list()

for (i in 1:length(constrast_list)) {
  cat(paste0("-----------------------------------------------------------------------------\n",
             names(constrast_list)[i],
             "\n-----------------------------------------------------------------------------\n"))
  
  print(constrast_results_list[[i]])
  
  ## Get normalization methods used for your own information
  constrast_norms_used_list[[i]] = dba.normalize(constrast_results_list[[i]], bRetrieve = TRUE)
  print(head(constrast_norms_used_list[[i]]))
}
```


<br>

## Differential binding analysis
```{r diff_analyses, warning=FALSE}
diffBind_analyses_results_list = list()

for (i in 1:length(constrast_list)) {
  cat(paste0("-----------------------------------------------------------------------------\n",
             names(constrast_list)[i],
             "\n-----------------------------------------------------------------------------\n"))
  diffBind_analyses_results_list[[i]] = dba.analyze(constrast_results_list[[i]])
  #print(head(diffBind_analyses_results_list[[i]]))
}


# ## Save analyses:
# saveRDS(diffBind_analyses_results_list,
#         file=paste0(wd,"/",diffBind_dir, "/tables_and_counts/",prefix_analyses,"_contrast.analyses.list_analyzed.RData"))
```

```{r diff_analyses_import, echo=FALSE}
# # import contrast_results_list
# diffBind_analyses_results_list = 
#   readRDS(paste0(wd,"/",diffBind_dir, "/tables_and_counts/",prefix_analyses,"_contrast.analyses.list_analyzed.RData"))
# 
# 
# for (i in 1:length(constrast_list)) {
#   cat(paste0("-----------------------------------------------------------------------------\n",
#              names(constrast_list)[i],
#              "\n-----------------------------------------------------------------------------\n"))
#   print(head(diffBind_analyses_results_list[[i]]))
# }
```


<!-- To see the overlaps of peaks among samples (available for max 4 samples): -->
```{r venn_diagram_overlaps, eval=FALSE, fig.align='center', fig.height=4.5, fig.width=5, warning=FALSE, include=FALSE}
#dba.plotVenn(diffBind_analyses_results_list[[i]], diffBind_analyses_results_list[[i]]$masks$Consensus)
```

<br>

## Correlation at *differential* peaks
### Corrlplot
Notice that you could perform the contrast with a list of contrasts, thus the 'contrast = 1' indicates that we want to plot the first contrast in your list.

```{r diff_corr_plot, warning=FALSE, fig.height=5, fig.width=5, fig.align='center'}
for (i in 1:length(constrast_list)) {
  dba.plotHeatmap(diffBind_analyses_results_list[[i]],
                  attributes = DBA_CONDITION,
                  correlations = T,
                  colScheme = PRGn,
                  ColAttributes = DBA_CONDITION,
                  colSideCols = cell_colors,
                  RowAttributes = DBA_REPLICATE,
                  rowSideCols = rep_colors,
                  minval = 0,
                  maxval = 1)
  
  
  # Export plot
  dir.create(paste0(wd, "/",diffBind_dir, "/plots/differential_analyses/"), showWarnings = F, recursive = T)
  pdf(file = paste0(wd, "/",diffBind_dir, "/plots/differential_analyses/heatmap_diffPeaks_correlation_", names(constrast_list)[i], ".pdf"),
      width = 7, height = 7.5)
  dba.plotHeatmap(diffBind_analyses_results_list[[i]],
                  attributes = DBA_CONDITION,
                  correlations = T,
                  colScheme = PRGn,
                  ColAttributes = DBA_CONDITION,
                  colSideCols = cell_colors,
                  RowAttributes = DBA_REPLICATE,
                  rowSideCols = rep_colors,
                  minval = 0,
                  maxval = 1)
  dev.off()
  
  
  ### export tables
  correlations_diffPeaks = dba.plotHeatmap(diffBind_analyses_results_list[[i]], correlations = T, olPlot = DBA_COR)
  write.table(x = data.frame(correlations_diffPeaks),
              file = paste0(wd, "/",diffBind_dir, "/tables_and_counts/heatmap_diffPeaks_correlation_", names(constrast_list)[i], ".txt"),
              sep = "\t", col.names = T, row.names = T, quote = F)
  #perc_overlap = dba.plotHeatmap(diffbind_MCF7.vs.Ishikawa, correlations = T, olPlot = DBA_OLAP)
  #num_common_peaks = dba.plotHeatmap(diffbind_MCF7.vs.Ishikawa, correlations = T, olPlot = DBA_INALL)
}
```


### Heatmap individual peaks
```{r diff_heatmaps, warning=FALSE, fig.height=5, fig.width=5, fig.align='center'}
for (i in 1:length(constrast_list)) {
  dba.plotHeatmap(diffBind_analyses_results_list[[i]],
                  attributes = DBA_CONDITION,
                  contrast = 1,
                  correlations = F,
                  colScheme = GnBu,
                  ColAttributes = c(DBA_CONDITION, DBA_REPLICATE),
                  colSideCols = list(rev(cell_colors[names(cell_colors) %in% constrast_list[[i]][2:3]]),
                                     rep_colors),
                  score = DBA_SCORE_NORMALIZED)
  
  
  # Export the plot
  pdf(file = paste0(wd, "/",diffBind_dir, "/plots/differential_analyses/heatmap_diffPeaks_scores_individual_peaks_", names(constrast_list)[i], ".pdf"),
      width = 7, height = 7.5)
  dba.plotHeatmap(diffBind_analyses_results_list[[i]],
                  attributes = DBA_CONDITION,
                  contrast = 1,
                  correlations = F,
                  colScheme = GnBu,
                  ColAttributes = c(DBA_CONDITION, DBA_REPLICATE),
                  colSideCols = list(rev(cell_colors[names(cell_colors) %in% constrast_list[[i]][2:3]]),
                                     rep_colors),
                  score = DBA_SCORE_NORMALIZED)
  dev.off()
}
```


### Plot MA
```{r DB_diff_MAplot, warning=FALSE, fig.height=5, fig.width=5, fig.align='center'}
for (i in 1:length(constrast_list)) {
  dba.plotMA(diffBind_analyses_results_list[[i]], th = FDR_threshold, contrast = 1)

  pdf(file = paste0(wd, "/",diffBind_dir, "/plots/differential_analyses/MAplot_diffPeaks_", names(constrast_list)[i], ".pdf"),
      width = 5, height = 4)
  dba.plotMA(diffBind_analyses_results_list[[i]], th = FDR_threshold, contrast = 1)
  dev.off()
}
```


### Plot volcano
```{r DB_diff_volacno, warning=FALSE, fig.height=5, fig.width=5, fig.align='center'}
for (i in 1:length(constrast_list)) {
  dba.plotVolcano(diffBind_analyses_results_list[[i]], th = FDR_threshold, contrast = 1)
  
  pdf(file = paste0(wd, "/",diffBind_dir, "/plots/differential_analyses/volcanoDiffBind_diffPeaks_", names(constrast_list)[i], ".pdf"),
      width = 5, height = 4)
  dba.plotVolcano(diffBind_analyses_results_list[[i]], th = FDR_threshold, contrast = 1)
  dev.off()
}
```



### Plot PCA
```{r PCA_diff, warning=FALSE, fig.height=5, fig.width=5, fig.align='center'}
for (i in 1:length(constrast_list)) {
  dba.plotPCA(diffBind_analyses_results_list[[i]],
              DBA_CONDITION,
              th = FDR_threshold,
              contrast = 1,
              components = c(1,2),
              label = DBA_REPLICATE,
              vColors = rev(cell_colors[names(cell_colors) %in% constrast_list[[i]][2:3]]))
  
  pdf(file = paste0(wd, "/",diffBind_dir, "/plots/differential_analyses/PCA_diffPeaks_", names(constrast_list)[i], "_PC1-2.pdf"),
      width = 7, height = 7)
  dba.plotPCA(diffBind_analyses_results_list[[i]],
              DBA_CONDITION,
              th = FDR_threshold,
              contrast = 1,
              components = c(1,2),
              label = DBA_REPLICATE,
              vColors = rev(cell_colors[names(cell_colors) %in% constrast_list[[i]][2:3]]))
  dev.off()
}


for (i in 1:length(constrast_list)) {
  dba.plotPCA(diffBind_analyses_results_list[[i]],
              DBA_CONDITION,
              th = FDR_threshold,
              contrast = 1,
              components = c(2,3),
              label = DBA_REPLICATE,
              vColors = rev(cell_colors[names(cell_colors) %in% constrast_list[[i]][2:3]]))
  
  pdf(file = paste0(wd, "/",diffBind_dir, "/plots/differential_analyses/PCA_diffPeaks_", names(constrast_list)[i], "_PC2-3.pdf"),
      width = 7, height = 7)
  dba.plotPCA(diffBind_analyses_results_list[[i]],
              DBA_CONDITION,
              th = FDR_threshold,
              contrast = 1,
              components = c(2,3),
              label = DBA_REPLICATE,
              vColors = rev(cell_colors[names(cell_colors) %in% constrast_list[[i]][2:3]]))
  dev.off()
}
```



## Get the raw data for custum analyses/plotting
### Get the normalized binding data
```{r export_norm_data_tb, warning=FALSE}
# the CHR names will be from 1 to 24 because X,Y, MT, etc are just indicated as indexes,
# so we need to get the names and put back in the table

count_tables_list = list()

for (i in 1:length(constrast_list)) {
  count_tables_list[[i]] = 
    dplyr::left_join(as.data.frame(diffBind_analyses_results_list[[i]]$binding),
                     data.frame(CHR = 1:length(unique(as.data.frame(diffBind_analyses_results_list[[i]]$binding)$CHR)),
                                chr.ID = diffBind_analyses_results_list[[i]]$chrmap),
                     by = "CHR") %>%
    dplyr::mutate(CHR = chr.ID) %>%
    dplyr::select(-chr.ID)
  
  
  cat(paste0("-----------------------------------------------------------------------------\n",
             names(constrast_list)[i],
             "\n-----------------------------------------------------------------------------\n"))
  #print(head(count_tables_list[[i]]))
  
  
  # Save the table
  write.table(x = count_tables_list[[i]],
              file = paste0(wd, "/",diffBind_dir, "/tables_and_counts/normalized_counts_at_peaks_",
                            names(constrast_list)[i], ".txt"),
              quote = F, sep = "\t", row.names = F, col.names = T)
}

names(count_tables_list) = names(constrast_list)
```

<br>

### Get differential binding data (no threholds)
```{r export_diff_binding_data__and_RePlotting, warning=FALSE}
report_list = list()

for (i in 1:length(constrast_list)) {
  report_list[[i]] = dba.report(diffBind_analyses_results_list[[i]],
                                               contrast = 1,
                                               DataType = DBA_DATA_FRAME,
                                               th = 1) # "fake" threshold FDR == 1
  
  colnames(report_list[[i]])[7] = paste0("log2FoldChange_",
                                         paste0(gsub("Conc_",
                                                     "",
                                                     colnames(report_list[[i]][5:6])),
                                                collapse = ".vs."))
  
  report_list[[i]] =
    report_list[[i]] %>%
    dplyr::mutate(diff.status = Rseb::DE.status(log2FC = .[[7]],
                                                p.value.adjusted = FDR,
                                                FC_threshold = FC_threshold,
                                                p.value_threshold = FDR_threshold,
                                                low.FC.status.label = paste("down", constrast_list[[i]][2]),
                                                high.FC.status.label = paste("up", constrast_list[[i]][2])))
  
  print(report_list[[i]] %>%
          dplyr::group_by(diff.status) %>%
          dplyr::summarise(n = n()))
  
  
  ## Save the table
  write.table(x = report_list[[i]],
              file = paste0(wd, "/",diffBind_dir, "/tables_and_counts/diffbind_differential_analyses_",
                            names(constrast_list)[i], ".txt"),
              quote = F,
              sep = "\t",
              row.names = F,
              col.names = T)
  
  
  ##############
  ## Re-plot the volcano
  # rename column for plot
  tb = report_list[[i]]
  names(tb)[7] = "log2FC"
  
  colors_plot = c("down" = cell_colors[grep(paste0("^",constrast_list[[i]][3],"$"), names(cell_colors))],
                  "up" = cell_colors[grep(paste0("^",constrast_list[[i]][2],"$"), names(cell_colors))],
                  "NoResp" = "lightskyblue",
                  "NULL" = "gray")
  
  names(colors_plot)[1:2] = c(paste("down", constrast_list[[i]][2]),
                              paste("up", constrast_list[[i]][2]))
  
  
  volcano_diff =
    ggplot(data = dplyr::mutate(tb, diff.status = factor(diff.status, levels = names(colors_plot))),
           aes(x = log2FC,
               y = -log10(FDR),
               color = diff.status)) +
    geom_point(alpha = 0.5,
               size = 2,
               stroke = NA) +
    geom_hline(yintercept = -log10(FDR_threshold), linetype = 3) +
    geom_vline(xintercept = c(-1,1)*log2(FC_threshold), linetype = 3) +
    scale_color_manual(values = colors_plot,
                       drop = F) +
    theme_classic() +
    ggtitle(paste("Differential peaks", names(constrast_list)[i])) +
    theme(axis.line = element_line(color = "black"),
          axis.text = element_text(color = "black"),
          axis.ticks = element_line(color = "black"))
  
  volcano_diff =
    volcano_diff +
    xlim(c(-1,1) * max(abs(ggplot_build(volcano_diff)$layout$panel_params[[1]]$x.range)))
  
  print(volcano_diff)
  
  pdf(file = paste0(wd, "/",diffBind_dir, "/plots/differential_analyses/volcano_reMade_diffPeaks_",
                            names(constrast_list)[i], ".pdf"),
      width = 7, height = 7)
  print(volcano_diff)
  dev.off()
}
```



# <font color="#074fa5">**Differential peaks annotations and genomic distribution**</font> 
## Differential peaks annotation
```{r diff_peaks_annotation}
require(ChIPseeker)
txdb = TxDb.Hsapiens.UCSC.hg38.knownGene::TxDb.Hsapiens.UCSC.hg38.knownGene

peak_anno_list = list()
diff_peak_anno_list = list()

for (i in 1:length(constrast_list)) {
  bed_formatted_tb = 
    GenomicRanges::makeGRangesFromDataFrame(dplyr::mutate(report_list[[i]][,1:3],
                                                          Chr = paste0("chr", Chr)))
  
  peak_anno_list[[i]] = ChIPseeker::annotatePeak(bed_formatted_tb,
                                                 tssRegion = promoter_range,
                                                 TxDb = txdb,
                                                 annoDb = "org.Hs.eg.db",
                                                 addFlankGeneInfo = TRUE,
                                                 flankDistance = 2000,
                                                 assignGenomicAnnotation = TRUE,
                                                 verbose = FALSE)
  
  
  # export the annotated bed
  write.table(x = as.data.frame(peak_anno_list[[i]]),
              file = paste0(wd, "/",diffBind_dir, "/tables_and_counts/annotated_diffbind_peaks_",
                            names(constrast_list)[i], ".txt"),
              col.names = TRUE,
              row.names = FALSE,
              sep = "\t", quote = FALSE)
  
  
  # Differential table with annotations
  diff_peak_anno_list[[i]] =
    Rseb::move.df.col(data.frame = dplyr::left_join(x = report_list[[i]] %>% dplyr::mutate(Chr = gsub("chr","",Chr)),
                                                    y = as.data.frame(peak_anno_list[[i]]) %>% dplyr::mutate(seqnames = gsub("chr","",seqnames)),
                                                    by = c("Chr" = "seqnames", "Start" = "start", "End" = "end")),
                      move.command = "width after End; strand after width")
  
  write.table(x = diff_peak_anno_list[[i]],
              file = paste0(wd, "/",diffBind_dir, "/tables_and_counts/annotated_diffBind_analyses_",
                            names(constrast_list)[i], ".txt"),
              col.names = TRUE,
              row.names = FALSE,
              sep = "\t", quote = FALSE)
}

names(peak_anno_list) = names(constrast_list)
names(diff_peak_anno_list) = names(constrast_list)


# Export lists
saveRDS(object = peak_anno_list,
        file = paste0(wd, "/",diffBind_dir, "/tables_and_counts/annotated_diffbind_peaks_list.RData"))

saveRDS(object = diff_peak_anno_list,
        file = paste0(wd, "/",diffBind_dir, "/tables_and_counts/annotated_diffBind_analyses_list.RData"))
```



<br>

--------------------------------------

# **Session info**
```{r session_info, echo=FALSE}
Rseb::pkg.version(print.versions = T)
```







